---
title: "n1 EB003E6: QC and monocle analysis"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(SparseArray)
library(Seurat)
#library(rCASC)
#library(deepToolsUtils)
#library(R.utils)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
#dir.create(file.path("Output","scratch"))


dir.create(file.path("Output","QC"))
dir.create(file.path("Output","QC", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "ribomito"))
dir.create(file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "umap"))

dir.create(file.path("Output","monocle"))#create forders MONOCLE
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap"))

fname_scratch <- file.path("scratch")

#set up folder path QC
fname_prefix_R_QC <- file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_csv_QC<-file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_ribomito_QC <- file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "ribomito", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_umap_QC <- file.path("Output","QC", format(Sys.Date(), "%y%m%d"), "umap", 
                               format(Sys.Date(), "%y%m%d"))

#set up forder path MONOCLE
fname_prefix_R <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))


okabe_tab <- read_excel("scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")


pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```

###load data
```{r}
#load(paste0(getwd(),"/scratch/241205_data_filtered_for_monocle.RData"))

sample_info=read.csv(paste0(getwd(),"/scratch/aggregation_2.csv"), header = TRUE, sep = ",", quote = "\"", 
                     dec = ".", fill = TRUE, comment.char = "",row.names=1)
load(paste0(getwd(),"/scratch/protein_coding_genes.RData"))

#load the RNA data
data <- Seurat::Read10X_h5(paste0(getwd(),"/scratch/filtered_feature_bc_matrix.h5"),unique.features = T)
counts=data$`Gene Expression`
counts=counts[row.names(counts) %in% protein_coding_genes,]#filter for protein coding genes straight away

```


```{r}
#load the RNA data
data <- Seurat::Read10X_h5(paste0(getwd(),"/scratch/filtered_feature_bc_matrix.h5"),unique.features = T)
counts=data$`Gene Expression`
counts=counts[row.names(counts) %in% protein_coding_genes,]#filter for protein coding genes straight away

# create a Seurat object containing the RNA data
data_seurat <- CreateSeuratObject(
  counts = data$`Gene Expression`,
  assay = "RNA"
)

#data_seurat[['ADT']] =CreateAssayObject(counts=data[["Antibody Capture"]]) #it doesn't seem to be working so I will add the ADT data to the RNA data in the metadata

adt=as.data.frame(data[["Antibody Capture"]])
colnames(adt)="Antibody_Capture"
adt$libID=rownames(adt)

sample_info$orig.ident=c("1","2","3")
meta_original=data_seurat[[]]
meta_original$libID=rownames(meta_original)
meta_original=meta_original%>%dplyr::select(libID)%>%dplyr::mutate(orig.ident=sapply(strsplit(as.character(libID), "-"), "[[", 2))
sample_info$orig.ident <- as.factor(sample_info$orig.ident)
meta_original=meta_original%>%left_join(sample_info, by="orig.ident")

meta_original$sample_id=paste0(meta_original$knockdown,"_",meta_original$replicate)
meta_original=meta_original%>%left_join(adt, by="libID")
rownames(meta_original)=meta_original$libID

pre_QCcounts=meta_original %>%dplyr::group_by(replicate,sample_id)%>%dplyr::summarise(count = n())
write.csv(pre_QCcounts, file= paste0(fname_prefix_csv_QC, "_", "pre_QCcounts.csv"), row.names=T)

#update metadata
data_seurat<- AddMetaData(data_seurat, meta_original, col.name = NULL)

```

#QC library

```{r}

#########calculate % of mitocondiral reads and ribosomal reads
DefaultAssay(data_seurat) <- 'RNA'
data_seurat[["percent.mt"]] <- PercentageFeatureSet(data_seurat, pattern = "^MT-")
data_seurat[["percent.ribo"]] <- PercentageFeatureSet(data_seurat, pattern = "RPS")

QC=data_seurat[[c("nFeature_RNA","percent.mt","nCount_RNA","percent.ribo","sample_id")]]

#make a QC dataframe to plot the QC metrics and then use it as meta table for monocle
QC <- QC %>%
        dplyr::rename(nUMI = nCount_RNA,
                      nGene = nFeature_RNA)
QC$info=rownames(QC)
QC$log10GenesPerUMI <- log10(QC$nGene) / log10(QC$nUMI)

plot1=VlnPlot(data_seurat, features = c("nFeature_RNA"),group.by = "sample_id",
     pt.size=0.1)+scale_fill_manual(values =c("#818589","#ff6db6","#56B4E9")) +geom_hline(yintercept = 400) #nGene
plot2=VlnPlot(data_seurat, features = c("nCount_RNA"),group.by = "sample_id",
     pt.size=0.1)+scale_fill_manual(values =c("#818589","#ff6db6","#56B4E9")) +geom_hline(yintercept = 1800) #nUMI
plot1 + plot2
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "feature_QC.pdf"),
     width = 10, height = 8)
###n UMI
QC %>% 
  	ggplot(aes(color=sample_id, x=nUMI, fill= sample_id)) + 
  	geom_density(alpha = 0.2) +scale_fill_manual(values =c("#818589","#ff6db6","#56B4E9")) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 100)
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "UMI_QC.pdf"),
     width = 10, height = 8)
###n genes detected
QC %>% 
  	ggplot(aes(color=sample_id, x=nGene, fill= sample_id)) + 
  	geom_density(alpha = 0.2) +scale_fill_manual(values =c("#818589","#ff6db6","#56B4E9")) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 400)
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "gene_det_QC.pdf"),
     width = 10, height = 8)
QC %>% 
  	ggplot(aes(x=sample_id, y=log10(nGene), fill=sample_id)) + 
  	geom_boxplot() +scale_fill_manual(values =c("#818589","#ff6db6","#56B4E9")) + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("NCells vs NGenes")
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "gene_QC.pdf"),
     width = 10, height = 8)
QC %>% 
  	ggplot(aes(x=nUMI, y=nGene, color=percent.mt)) + 
  	geom_point(size=1.5,alpha = 0.5) + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 100) +
  	geom_hline(yintercept = 400) 
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "gene_UMI_QC.pdf"),
     width = 10, height = 8)

QC %>% 
  	ggplot(aes(x=nUMI, y=nGene, color=sample_id)) + 
  	geom_point(size=1.5,alpha = 0.5) + scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9")) + 
	#scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 100) +
  	geom_hline(yintercept = 400) 
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "gene_UMI_QC_sample_id.pdf"),
     width = 10, height = 8)

QC %>% 
  	ggplot(aes(x=percent.ribo, y=percent.mt, color=sample_id)) + 
  	geom_point(size=1.5,alpha = 0.5) + scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9")) +
  	#stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 20,linetype="dotted")+
  geom_hline(yintercept = 20,linetype="dotted")
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "percent_MT_RIBO_QC.pdf"),
     width = 10, height = 8)

QC %>%
  	ggplot(aes(x=log10GenesPerUMI, color = sample_id, fill=sample_id)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = 0.85)
#rownames(QC)=QC$info
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "log10GenesPerUMI_QC.pdf"),
     width = 10, height = 8)

plot1 <- FeatureScatter(data_seurat, feature1 = "nCount_RNA", feature2 = "percent.mt",group.by = "sample_id")+geom_vline(xintercept = 100,linetype="dotted")+
  geom_hline(yintercept = 20,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot2 <- FeatureScatter(data_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA",group.by = "sample_id")+geom_vline(xintercept = 100,linetype="dotted")+
  geom_hline(yintercept = 400,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot3 <- FeatureScatter(data_seurat, feature1 = "percent.ribo", feature2 = "percent.mt",group.by = "sample_id")+geom_vline(xintercept = 20,linetype="dotted")+
  geom_hline(yintercept = 20,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot1 + plot2+ plot3
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "all_QC.pdf"),
     width = 20, height = 8)

######################################
#FILTER CELLS by QC metrics #NB. MAria cleaned quite well from the barcode QC analysis
filtered_seurat <- subset(data_seurat, subset = nFeature_RNA>400 & nCount_RNA>100 & percent.mt<20)

metadata=filtered_seurat[[]]
counts <- GetAssayData(object = filtered_seurat, slot = "counts")

write.csv(counts, paste0(fname_prefix_csv_QC,"_silencing_matrix_all_samples_EB003E6.csv",sep=""))
write.csv(metadata, paste0(fname_prefix_csv_QC,"_cell_metadata_filtered_EB003E6.csv",sep=""))

Rdata_filename = paste0(fname_prefix_R_QC, "_seurat_data_filtered.RData")
save(meta_original, counts, metadata,filtered_seurat,sample_info,
     file = Rdata_filename)


plot1 <- FeatureScatter(filtered_seurat, feature1 = "nCount_RNA", feature2 = "percent.mt",group.by = "sample_id")+geom_vline(xintercept = 100,linetype="dotted")+
  geom_hline(yintercept = 20,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot2 <- FeatureScatter(filtered_seurat, feature1 = "nCount_RNA", feature2 = "nFeature_RNA",group.by = "sample_id")+geom_vline(xintercept = 100,linetype="dotted")+
  geom_hline(yintercept = 400,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot3 <- FeatureScatter(filtered_seurat, feature1 = "percent.ribo", feature2 = "percent.mt",group.by = "sample_id")+geom_vline(xintercept = 20,linetype="dotted")+
  geom_hline(yintercept = 20,linetype="dotted")+ scale_colour_manual(values =c("#818589","#ff6db6","#56B4E9"))
plot1 + plot2+ plot3
ggsave(filename = paste0(fname_prefix_ribomito_QC, "_", "all_QC_after_FILTER.pdf"),
     width = 20, height = 8)

```


#add prepare data for Monocle

```{r}
####gather info from Seurat
cell_metadata=metadata
expression_matrix <- filtered_seurat@assays[["RNA"]]@layers[["counts"]]
counts <- GetAssayData(object = filtered_seurat, slot = "counts")

# get gene annotations
gene_annotation=data.frame(rownames(counts))
colnames(gene_annotation) <- "gene_short_name"
rownames(gene_annotation)=gene_annotation$gene_short_name

```

####preprocess monocle data

```{r}

###create cds file
cds <- new_cell_data_set(counts,cell_metadata = cell_metadata,
                                      gene_metadata = gene_annotation)
cds <- preprocess_cds(cds, num_dim = 50)
#cds <- align_cds(cds, num_dim = 100, alignment_group = "sample")
plot_pc_variance_explained(cds)

ggsave(filename = paste0(fname_prefix_dotplot, "_", "variance.png"),
     width = 12, height = 8)

#cds <- align_cds(cds, alignment_group = "experiment")
set.seed(678686)
cds = reduce_dimension(cds,reduction_method="tSNE")
set.seed(79876858)
cds = reduce_dimension(cds,umap.min_dist = 0.2,reduction_method="UMAP")
#cds = reduce_dimension(cds,umap.min_dist = 0.5,reduction_method="UMAP")

#monocle.object = order_cells(monocle.object, reduction_method = "UMAP")
set.seed(12345)
cds_1 = cluster_cells(cds, resolution = 1e-2)
colData(cds_1)$monocle_clusters = as.character(monocle3::clusters(cds_1))
set.seed(123479870)
cds_2 = cluster_cells(cds, resolution = 1e-4)
colData(cds_2)$monocle_clusters = as.character(monocle3::clusters(cds_2))
set.seed(123457768)
cds_3 = cluster_cells(cds, resolution = 8e-4)
colData(cds_3)$monocle_clusters = as.character(monocle3::clusters(cds_3))
set.seed(123458980)
cds_4 = cluster_cells(cds, resolution = 2e-4)
colData(cds_4)$monocle_clusters = as.character(monocle3::clusters(cds_4))
set.seed(1234597698)
cds_5 = cluster_cells(cds, resolution = 5e-4)
colData(cds_5)$monocle_clusters = as.character(monocle3::clusters(cds_5))


plot_cells(cds_1, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_2, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_3, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_4, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_5, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)

Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered.RData")
save(cell_metadata, counts,cds, cds_1,
     cds_2, cds_3, cds_4, cds_5,
     file = Rdata_filename)
```

###chose the cluster iteration with better results and clean environment

```{r}
plot_cells(cds_3, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP.pdf"),
     width = 18, height = 8)

#map by seurat clusters generated
cds_3 <- learn_graph(cds_3)
cds=cds_3#this is the iteration we decided to use

#cds_cardio= cds[,cds@colData@listData[["monocle_clusters"]]=="5"|cds@colData@listData[["monocle_clusters"]]=="4"|cds@colData@listData[["monocle_clusters"]]=="3"|cds@colData@listData[["monocle_clusters"]]=="2"]


Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,
     file = Rdata_filename)

#clean environment
rm(combined_exp,data_exp5,expression_data,gene_annotation,expression_matrix,cds_1,cds_2,cds_3,cds_4,cds_5)
```

###select cds with protein coding genes
#add umap data on cds meta
#add pseudotime info

```{r}
#add UMAP data to cell_metadata
pbuild <- plot_cells(cds, color_cells_by = "monocle_clusters", label_cell_groups=T, show_trajectory_graph = F, label_leaves=F, label_branch_points=F, graph_label_size=3, group_label_size = 4, cell_size = 1) 
UmapCDS <- data.frame(pbuild$data$sample_name,UMAP1_monocle = pbuild$data$data_dim_1, UMAP2_monocle = pbuild$data$data_dim_2, row.names = 1) 
colData(cds)$UMAP_1_monocle=pbuild$data$data_dim_1 #add umap data for further analysis outside monocle
colData(cds)$UMAP_2_monocle=pbuild$data$data_dim_2 #add umap data for further analysis outside monocle

original_counts=counts
#counts=as.data.frame(exprs(cds))
counts=as.data.frame(normalized_counts(cds))
meta=as.data.frame(cds@colData@listData)

#cds<- order_cells(cds,reduction_method = "UMAP")#selected around cluster 3

#plot_cells(cds,
 #          color_cells_by = "pseudotime",
    #       label_cell_groups=FALSE,
   #        label_leaves=FALSE,
  #         label_branch_points=FALSE,
 #          graph_label_size=1.5)
#last_plot()+facet_grid(~sample)
#ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_pseudotime.pdf"),
 #    width = 18, height = 8)

Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,meta,original_counts,
     file = Rdata_filename)


```

##make graphs to visualize genes

```{r}

#neuro_genes=c("PAX6","ZIC1","ZIC4","SIX3","GBX2","OLIG3","NEUROG3","NEUROD6","FOXC1","ZNF37A","BAHCC1")

plot_cells(cds,
           genes = c("SNAI1","NODAL","SOX4","BMP6","BMP4","BMP2","CDH1","EGFR","FOXD1","MYL6","GATA4","SNAI2","LEF1","ZEB1","ALK","DDR2","ACTA2","COL1A2","NPY","FN1","ATP1B3","B2M"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_general.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("TTN","MYH6","MYH7","TNNT2","MYL3","EOMES","MYL4","COL6A3","COL1A1","HAND2","CDH5","TOP2A","IRX4","ESRRG","RYR2","CDH1","TBX18","ITGB1BP2","WT1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cardio.pdf"),
     width = 10, height = 10)
 
  plot_cells(cds,
           genes = c("SOX4","EGFR","MYL7","CDH1","TBX18","WT1","BMP4","HAND1","FOXC1","FOXC2","ADAMTS18","LHX1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_epi.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("CDH3","CDH1","TBX18","WT1","KRT19","KRT24","DSG2","ITGA3","PDPN","IL1R1","HAND1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_epithelial.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("COL6A3","COL1A1","COL1A2","DDR2","ACTA2"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_fibroblasts.pdf"),
     width = 10, height = 10)
 
 
  plot_cells(cds,
           genes = c("PDPN","TBX18","WT1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_epicardial_final.pdf"),
     width = 10, height = 5)
 
 plot_cells(cds,
           genes = c("CDH1","PODXL","SFRP5"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_epicardial_final2.pdf"),
     width = 10, height = 5)
 
plot_cells(cds,
           genes = c("FOXC1","EGFR","BMP4","HAND2"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl11_defa.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("FOXC1","EGFR","BMP4","HAND2"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl11_defa.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("FCRABP1","SH3BGRL3",
"ANXA1","SSR4","SDF2L1","LHX1","DNAH8","KRT10",
"CAPN6",
"RTL4",
"LAMTOR5",
"PHLDA2",
"PDCD4",
"DIO2",
"ENSG00000275091",
"ENSG00000272736",
"CTSB",
"CYP26B1",
"TUBB3","ENSG00000228999","ENSG00000289112","ENSG00000269559","ENSG00000290412","FOXC1","EGFR","BMP4","HAND2"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl11_defa.pdf"),
     width = 10, height = 10)
 
 plot_cells(cds,
           genes = c("CRABP1",
"SH3BGRL3","ANXA1","SSR4","SDF2L1","LHX1","DNAH8",
"KRT10","CAPN6","RTL4","LAMTOR5","PHLDA2","PDCD4","DIO2","ENSG00000275091","ENSG00000272736","CTSB","CYP26B1","TUBB3","ENSG00000228999","ENSG00000289112","ENSG00000269559","ENSG00000290412"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl11_defb.pdf"),
     width = 10, height = 10)
 


plot_cells(cds,
           genes = c("FTL","TMSB4X","LGALS1","ITM2B","KRT8","KRT18","RPS9","HBE1","EMP2","FREM2","CDKN1A","VGF","FGF13","HSP90AA1","LRP2"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl11_defc.pdf"),
     width = 10, height = 10)
 
 
 plot_cells(cds,
           genes = c("TCF21","TCF4","UPK3B","ITGA4","PODXL","SFRP5","ALDH1A2","VIM","TWIST1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_cl13_epi.pdf"),
     width = 10, height = 10)
```

###make graphs for clusters and features

```{r}

ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = monocle_clusters),
               size=0.8)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clusters")+
  scale_color_manual(name = "monocle",
                        values=c(pal))
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP.pdf"),
     width = 6, height = 6)

ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = monocle_clusters),
               size=0.8)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clusters")+
  scale_color_manual(name = "monocle",
                        values=c(pal))+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_monocle_clusters.pdf"),
     width = 12, height = 6)


ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = knockdown),
               size=0.8)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "knockdown")+
  scale_color_manual(name = "monocle",
                        values=c(okabe_pal))+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_samp.pdf"),
     width = 12, height = 6)
  
plot_feature=c("percent.mt", "percent.ribo","nFeature_RNA","nCount_RNA","Antibody_Capture")

for (feature in plot_feature) {
  
  p <- ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes_string(color = feature), size = 0.8) +
    labs(
      x = "UMAP 1",
      y = "UMAP 2",
      title = feature
    ) +
    scale_color_viridis(option = "E", discrete = FALSE) +
    facet_grid(~sample_id)
  
  ggsave(
    filename = paste0(fname_prefix_UMAP, "_", "UMAP_", feature, ".pdf"),
    plot = p,
    width = 8,
    height = 4
  )
}


```
###get top genes in a loop 

```{r}
marker_test_res = top_markers(cds, group_cells_by="monocle_clusters", reference_cells=NULL, cores=16)
marker_test_res <- top_markers(cds, group_cells_by="monocle_clusters", 
                               reference_cells=10000, cores=8)

# List of top_n values to iterate through
top_n_values <- c(2, 5, 10, 25)

# Loop through each top_n value
for (n in top_n_values) {
  
  # Filter markers and select the top_n genes
  top_specific_markers <- marker_test_res %>%
    filter(fraction_expressing >= 0.10) %>%
    group_by(cell_group) %>%
    top_n(n, pseudo_R2)
  
  # Extract unique gene IDs
  top_specific_marker_ids <- unique(top_specific_markers %>% dplyr::pull(gene_id))
  
  # Create a data frame for the top genes
  t_data <- data.frame(
    ids = top_specific_markers$gene_id, 
    cluster = top_specific_markers$cell_group, 
    stringsAsFactors = FALSE
  )
  
  # Write the data to a CSV file
  write.csv(
    t_data, 
    file = paste0(fname_prefix_csv, "_", "top_", n, "genes_per_cluster.csv"), 
    row.names = FALSE
  )
  
  # Generate a heatmap plot
  p <- plot_genes_by_group(
    cds,
    top_specific_marker_ids,
    group_cells_by = "monocle_clusters",
    ordering_type = "maximal_on_diag",
    max.size = 3
  ) + scale_color_viridis(option = "E", discrete = FALSE)
  
  # Save the plot
  ggsave(
    filename = paste0(fname_prefix_mosaic, "_", "top", n, "_heatmap.pdf"),
    plot = p,
    width = 8, 
    height = ifelse(n == 5, 10, 25) # Adjust the height for larger plots
  )
}

```


##get top 10 genes per cluster

```{r}

marker_test_res = top_markers(cds, group_cells_by="monocle_clusters", reference_cells=NULL, cores=16)
marker_test_res <- top_markers(cds, group_cells_by="monocle_clusters", 
                               reference_cells=10000, cores=8)

top_specific_markers <- marker_test_res %>%
                            filter(fraction_expressing >= 0.10) %>%
                            group_by(cell_group) %>%
                            top_n(10, pseudo_R2)

top_specific_marker_ids <- unique(top_specific_markers %>% dplyr::pull(gene_id))
t10=data.frame(top_specific_markers$gene_id, top_specific_markers$cell_group, stringsAsFactors = F)
colnames(t10)=c("ids", "cluster")

write.csv(t10, file= paste0(fname_prefix_csv, "_", "top_10genes_per_cluster.CSV"), row.names=FALSE)

plot_genes_by_group(cds,
                    top_specific_marker_ids,
                    group_cells_by="monocle_clusters",
                    ordering_type="maximal_on_diag",
                    max.size=3)+ scale_color_viridis(option="E", discrete=F)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "top10_heatmap.pdf"),
     width = 8, height = 15)

```

###add modules

###get gene modules shanging along pseudotime

```{r}

#cds@colData$group=paste0(cds@colData$shRNA_gene,"_",cds@colData$knockdown)
cds@colData$group=cds@colData$sample_id
#cell_metadata$group=paste0(cell_metadata$shRNA_gene,"_",cell_metadata$knockdown)
cell_metadata$group=cell_metadata$sample_id

#########find gene modules along trajectory
set.seed(685875890)
cds_pr_test_res <- graph_test(cds, neighbor_graph="knn", cores=4)#score genes not based on trajectory
genes_mod <- row.names(subset(cds_pr_test_res, q_value < 0.05))

gene_module_df <- find_gene_modules(cds[genes_mod,], reduction_method = c("UMAP"),resolution=0.01)#organize genes into gene modules

cell_group_df <- tibble::tibble(cell=row.names(colData(cds)), 
                                cell_group=colData(cds)$monocle_clusters)###group modules by cluster
cell_group_df_test <- tibble::tibble(cell=row.names(colData(cds)), 
                                cell_group=colData(cds)$group)###group modules by shRNA gene and KD (no selection yet) #gruping based by gene KD

agg_mat <- aggregate_gene_expression(cds, gene_module_df)###every lib ID has a value per each module

modules=as.data.frame(t(agg_mat))
#modules <- modules[, order(names(modules))]
modules <- modules[, order(as.numeric(names(modules)))]
modules=as.data.frame(modules)

cds@colData[, paste0("module_", 1:93)] <- modules[, 1:93]###added 93 modules
cell_metadata[, paste0("module_", 1:93)] <- modules[, 1:93]

agg_mat_cluster <- aggregate_gene_expression(cds, gene_module_df, cell_group_df)
row.names(agg_mat_cluster) <- stringr::str_c("module ", row.names(agg_mat))
hp1=pheatmap::pheatmap(agg_mat_cluster,
                   scale="row", clustering_method="ward.D2")

Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_cl.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()

agg_mat_test <- aggregate_gene_expression(cds, gene_module_df, cell_group_df_test)
row.names(agg_mat_test) <- stringr::str_c("module ", row.names(agg_mat_test))
hp2=pheatmap::pheatmap(agg_mat_test,
                   scale="row", clustering_method="ward.D2",main="row normalized modules all",color = viridis(n = 256, alpha = 1,begin = 0, end = 1, option = "H"))

Rdata_filename = paste0(fname_prefix_mosaic, "_hp_all_modules.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp2
dev.off()

##B2M and ChD7 look different than the others. I believe this is because CHD7 is mainly side 1 and B2M cells all disappear
Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_with_modules_FINAL.RData")
save(cell_metadata, counts,cds,meta,original_counts,modules,gene_module_df,cds_pr_test_res,genes_mod,
     agg_mat,agg_mat_cluster,agg_mat_test,
     file = Rdata_filename)

#update monocle with clusters without all the modules needed for plotting
Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,meta,original_counts,
     file = Rdata_filename)

```

####plot modules

```{r}

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(1:93)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle top modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

#ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules.pdf"),
 #    width = 21, height = 30)

```
####filter from github

```{r}
# Source the function from GitHub

#devtools::install_github("https://github.com/ebalmas/sc_RNAseq")

#library(scRNAseq)

#df=modules
#colnames(df)=paste0("module_", 1:93)

# Get the filter modules with custom thresholds
#x <- get_filter_modules(df, percentage_threshold = 0.1, value_threshold = 65)

# Print the filter modules
#print(filter_modules)
#print(x)
```

####module selection

```{r}

df=modules
colnames(df)=paste0("module_", 1:93)

convert_to_percentage <- function(column) {
  rng <- range(column)
  if (rng[2] > rng[1]) {
    # Scale the values such that the min becomes 0 and the max becomes 100.
    (column - rng[1]) / (rng[2] - rng[1]) * 100
  } else {
    # If all values are the same, return 0 for every cell
    rep(0, length(column))
  }
}

a <- data.frame(row.names = rownames(df))
# Apply the function to each module (column) from Module_1 to Module_93
for (i in 1:93) {
  module_name <- paste0("module_", i)
  a[[module_name]] <- convert_to_percentage(df[[module_name]])
}

# Function to filter columns based on the percentage of cells with values over 60%
filter_columns_by_percentage <- function(a) {
  # Calculate number of cells (rows)
  num_cells <- nrow(a)
  # Define the threshold for 10% of cells
  threshold <- ceiling(0.10 * num_cells)
  
  # Initialize an empty dataframe to store the filtered columns
  filtered_df <- data.frame(row.names = rownames(a))
  
  # Loop through each column in the dataframe
  for (colname in colnames(a)) {
    column <- a[[colname]]
    
    # Calculate the number of cells with values over 60%
    num_high_values <- sum(column > 65)
    
    # Print diagnostic info (optional; remove when satisfied)
    #cat(sprintf("Module %s: %d cells have values over 60%%\n", colname, num_high_values))
    
    # If more than 10% of cells have values over 60%, keep this column
    if (num_high_values >= threshold) {
      filtered_df[[colname]] <- column
    }
  }
  
  return(filtered_df)
}

# Apply the function to the dataframe `a`
filtered_a <- filter_columns_by_percentage(a)

filter_modules=colnames(filtered_a)

gene_module_df$module_nb=paste0("module_",gene_module_df$module)
plot_cells(cds,
           genes=gene_module_df %>% filter(module_nb %in% c(filter_modules)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle top modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

#ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules.pdf"),
 #    width = 21, height = 30)

####eliminate the modules with expression in only few cells abd general low expression
#make sure names of the dataframe to filter and the module vector are the same
b=as.data.frame(agg_mat_test)
b$m=rownames(b)
b=b%>%dplyr::select(m)%>%mutate(mod=sapply(strsplit(as.character(m), " "), "[[", 2)) 
b$m_new=paste0("module_",b$mod)
module_names=b$m_new

#select modules based on the filtered modules vector
select_modules=as.data.frame(agg_mat_test)
rownames(select_modules)=module_names
select_modules <- select_modules[rownames(select_modules) %in% filter_modules, ]

hp1=pheatmap::pheatmap(select_modules,
                   scale="row", clustering_method="ward.D2",main="row normalized modules filtered")#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_filtered.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()
hp2=pheatmap::pheatmap(agg_mat_test,
                   scale="row", clustering_method="ward.D2")#scaled on row (modules) ###original data >>>to compare modules between perturbations
hp2
hp1=pheatmap::pheatmap(select_modules,
                   scale="row", clustering_method="ward.D2",main="row normalized modules filtered",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_filtered_TURBO.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()


#######get z score of the modules (the values of the heatmap hp1 and hp2)

cal_z_score <- function(x){
  (x - mean(x)) / sd(x)
}

data_subset_norm <- t(apply(select_modules, 1, cal_z_score)) ###get z scored data
hp3=pheatmap::pheatmap(data_subset_norm, clustering_method="ward.D2")###check hp4 is the same as hp2

data_subset_norm=as.data.frame(data_subset_norm)
#data_subset_norm <- data_subset_norm%>%select(-CHD7_CTR,-CHD7_TET) #these created a byas due to the side
data_subset_norm$min<-apply(data_subset_norm,1,FUN=min)
data_subset_norm$max<-apply(data_subset_norm,1,FUN=max)
data_subset_norm$delta=data_subset_norm$max-data_subset_norm$min

top=data_subset_norm[order(-data_subset_norm$delta),] ####select top 16 based on delta
top=head(top, n=16)

select_modules=select_modules%>%dplyr::filter(rownames(select_modules)%in%rownames(top))

rownames(top)
to_plot=rownames(top)

# "module_15" "module_37" "module_60" "module_2"  "module_27" "module_41" "module_25" "module_48" "module_65" "module_44" "module_59" "module_73" "module_56" "module_62" "module_39"
# "module_17"

plot_cells(cds,
           genes=gene_module_df %>% filter(module_nb %in% c(to_plot)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle top 16 modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_top_16_modules.pdf"),
     width = 10, height = 10)

##make the same hm as before but flipped
select_modules=t(select_modules)#flip data do have genes in rows and modules in columns

hp3=pheatmap::pheatmap(select_modules,
                   scale="column", clustering_method="ward.D2",main="Top 16 modules")#scaled on row (modules)
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_top16.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp3
dev.off()

hp3=pheatmap::pheatmap(select_modules,
                   scale="column", clustering_method="ward.D2",main="Top 16 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on row (modules)
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_top16_Turbo.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp3
dev.off()


Rdata_filename = paste0(fname_prefix_R, "_updated_cds_with_modules_SELECTION.RData")
save(cell_metadata,counts,cds,meta,original_counts,modules,gene_module_df,cds_pr_test_res,genes_mod,filter_modules,b,to_plot,
     agg_mat,agg_mat_cluster,agg_mat_test,modules,top,select_modules,cds_pr_test_res,
     file = Rdata_filename) 

Rdata_filename = paste0(fname_prefix_R, "modules.RData")
save(gene_module_df,agg_mat,agg_mat_cluster,agg_mat_test,modules,top,select_modules,genes_mod,filter_modules,
     file = Rdata_filename) 
```



###convert to Seurat

```{r}

seurat<-as.Seurat(cds, assay = NULL)
FeaturePlot(seurat, features = c("TTN","MYH6","MYH7","TNNT2","MYL3","EOMES","MYL4","COL6A3","HAND2","CDH5","TOP2A","IRX4","ESRRG","RYR2","APOB","TBX18","ITGB1BP2","POU5F1","Antibody_Capture"))
DimPlot(seurat)

Rdata_filename = paste0(fname_prefix_R, "_monocle_seurat_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,filtered_a,top,meta,original_counts,seurat,
     file = Rdata_filename)
#load data EB003E6 if you want to start again from here
load("scratch/250227_monocle_seurat_cds_clustered_FINAL.RData")
#load AB003 seurat object
load("scratch/AB003.RData")
#load AB002 seurat object
load("scratch/AB002_cds.RData")#data cardioids GATA4KD

seurat_AB002_d4<-as.Seurat(cds_day4, assay = NULL)
#seurat_AB002_d7<-as.Seurat(cds_day7, assay = NULL)
#rm (cds_day7,counts, original_counts,filtered_a)

####combine UMAPS
seurat_AB003 <- FindVariableFeatures(seurat_AB003, selection.method = "vst", nfeatures = 8000)
seurat_AB003 =ScaleData(seurat_AB003)
seurat_AB003 =RunPCA(seurat_AB003 , npcs = 30)
DimPlot(seurat_AB003, group.by = "monocle_clusters", label = F)

seurat_AB002_d4=FindVariableFeatures(seurat_AB002_d4, selection.method = "vst", nfeatures = 8000)
seurat_AB002_d4 =ScaleData(seurat_AB002_d4)
seurat_AB002_d4 =RunPCA(seurat_AB002_d4 , npcs = 30)
DimPlot(seurat_AB002_d4, group.by = "monocle3_clusters", label = F)

seurat <- FindVariableFeatures(seurat, selection.method = "vst", nfeatures = 8000)
seurat =ScaleData(seurat)
seurat =RunPCA(seurat , npcs = 30)
DimPlot(seurat, group.by = "monocle_clusters", label = F)
# Find transfer anchors AB003
anchors <- FindTransferAnchors(reference = seurat_AB003, 
                               query = seurat, 
                               normalization.method = "LogNormalize", 
                               #features = VariableFeatures(object = seurat_AB003),
    reference.assay = "originalexp", query.assay = "originalexp", reduction = "pcaproject",
                               #features="PCA",
                               dims = 1:30)

# Transfer labels from reference to query AB003
predictions <- TransferData(anchorset = anchors, 
                            refdata = seurat_AB003$monocle_clusters, 
                            dims = 1:30)
predictions$libID=rownames(predictions)

predictions=predictions%>%select(predicted.id,libID)%>%dplyr::rename(predicted_AB003=predicted.id)

# Find transfer anchors AB002
anchors_2 <- FindTransferAnchors(reference = seurat_AB002_d4, 
                               query = seurat, 
                               normalization.method = "LogNormalize", 
                               #features = VariableFeatures(object = seurat_AB003),
    reference.assay = "originalexp", query.assay = "originalexp", reduction = "pcaproject",
                               #features="PCA",
                               dims = 1:30)

# Transfer labels from reference to query AB002
predictions_2 <- TransferData(anchorset = anchors_2, 
                            refdata = seurat_AB002_d4$monocle3_clusters, 
                            dims = 1:30)
predictions_2$libID=rownames(predictions_2)

predictions_2=predictions_2%>%select(predicted.id,libID)%>%dplyr::rename(predicted_AB002=predicted.id)

predictions=predictions%>%right_join(predictions_2, by="libID")
#cell_metadata=cell_metadata%>%select(-predicted_AB002,-predicted_AB003,-assigned_cell_type,-assigned_cell_type_short,-assigned_cell_type_short)
#cell_metadata=cell_metadata%>%select(-predicted_AB002.x,-predicted_AB003.x,-predicted_AB002.y,-predicted_AB003.y)
#cell_metadata=cell_metadata%>%select(-assigned_cell_type,-assigned_cell_type_short,-assigned_cell_type_short)


cell_metadata=cell_metadata%>%right_join(predictions, by="libID")
cds@colData$predicted_AB003=predictions$predicted_AB003
cds@colData$predicted_AB002=predictions$predicted_AB002

m=as.data.frame(cds@colData@listData)

cell_metadata$assigned_cell_type_AB003=cell_metadata$predicted_AB003
cell_metadata$assigned_cell_type_short_AB003=cell_metadata$predicted_AB003
cell_metadata$assigned_cell_type_AB002=cell_metadata$predicted_AB002
cell_metadata$assigned_cell_type_short_AB002=cell_metadata$predicted_AB002

cell_metadata$assigned_cell_type_AB003 <- dplyr::recode(cell_metadata$assigned_cell_type_AB003,
                                                 "1"="Cardiac fibroblasts",
                                                 "2"="Late cardiomyocytes",
                                                 "3"="Early cardiomyocytes",
                                                 "4"="Cardiac progenitors",
                                                 "5"="Lateral plate mesoderm",
                                                 "6"="Proliferating cardiomyocytes",
                                                 "7"="Endothelial cells",
                                                 "8"="Mesendoderm",
                                                 "9"="Endoderm derivatives")

cell_metadata$assigned_cell_type_short_AB003 <- dplyr::recode(cell_metadata$assigned_cell_type_short_AB003,
                                                 "1"="CF",
                                                 "2"="L-CM",
                                                 "3"="E-CM",
                                                 "4"="CP",
                                                 "5"="LPM",
                                                 "6"="P-CM",
                                                 "7"="EC",
                                                 "8"="MES",
                                                 "9"="ENDO")

cell_metadata$assigned_cell_type_AB002 <- dplyr::recode(cell_metadata$assigned_cell_type_AB002,
                                                 "1"="Cardiac fibroblasts",
                                                 "2"="Late cardiomyocytes-1",
                                                 "3"="Proliferative cardiomyocites",
                                                 "4"="Early cardiomyocytes-1",
                                                 "5"="Late cardiomyocites CTCF-KD",
                                                 "6"="Late cardiomyocites-2",
                                                 "7"="Mesendoderm",
                                                 "8"="Early cardiomyocites-2",
                                                 "9"="Undifferentiated",
                                                 "10"="Endoderm derivatives",
                                                 "11"="Late cardiomyocites-3",
                                                 "12"="Endothelial cells")

cell_metadata$assigned_cell_type_short_AB002 <- dplyr::recode(cell_metadata$assigned_cell_type_short_AB002,
                                                 "1"="CF",
                                                 "2"="L-CM-1",
                                                 "3"="P-CM",
                                                 "4"="E-CM-1",
                                                 "5"="L-CM CTCF-KD",
                                                 "6"="L-CM-2",
                                                 "7"="Mes",
                                                 "8"="E-CM-2",
                                                 "9"="Undif",
                                                 "10"="ENDO",
                                                 "11"="L-CM-3",
                                                 "12"="EC")

cds@colData$assigned_cell_type_short_AB003=cell_metadata$assigned_cell_type_short_AB003
cds@colData$assigned_cell_type_short_AB002=cell_metadata$assigned_cell_type_short_AB002
cds@colData$assigned_cell_type_AB003=cell_metadata$assigned_cell_type_AB003
cds@colData$assigned_cell_type_AB002=cell_metadata$assigned_cell_type_AB002

m=as.data.frame(cds@colData@listData)

# Add the predicted cluster annotations to the query metadata
seurat <- AddMetaData(object = seurat, metadata = cell_metadata)

DimPlot(seurat, group.by = "monocle_clusters", label = T)
DimPlot(seurat, group.by = "predicted_AB003", label = F)
DimPlot(seurat, group.by = "assigned_cell_type_short_AB003", label = T)
DimPlot(seurat, group.by = "assigned_cell_type_AB003", label = F)
DimPlot(seurat, group.by = "predicted_AB002", label = T)#data cardioids GATA4KD
DimPlot(seurat, group.by = "assigned_cell_type_short_AB002", label = T)#data cardioids GATA4KD
DimPlot(seurat, group.by = "assigned_cell_type_AB002", label = F)#data cardioids GATA4KD
DimPlot(seurat_AB002_d4, group.by = "monocle3_clusters", label = F)


plot_feature=c("predicted_AB003","predicted_AB002","assigned_cell_type_short_AB003","assigned_cell_type_short_AB002","assigned_cell_type_AB002","monocle_clusters")

for (feature in plot_feature) {
  
  p <- ggplot(m, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes_string(color = feature), size = 0.8,alpha=1) +
    labs(
      x = "UMAP 1",
      y = "UMAP 2",
      title = feature
    ) +
    scale_color_manual(name = "monocle",
                        values=c(pal)) +
    facet_grid(~sample_id)
  
  ggsave(
    filename = paste0(fname_prefix_UMAP, "_", "UMAP_", feature, ".pdf"),
    plot = p,
    width = 8,
    height = 4
  )
}

# Merge the reference and query datasets
#combined <- merge(seurat_AB003, seurat, add.cell.ids = c("reference", "query"))

# Create a combined UMAP
#combined <- FindVariableFeatures(combined, selection.method = "vst", nfeatures = 8000)
#combined =ScaleData(combined)
#combined =RunPCA(combined , npcs = 30)
#combined <- RunUMAP(combined, reduction = "pca", dims = 1:30)

# Plot the combined UMAP
#DimPlot(combined, group.by = "sample_id", label = TRUE)

rm(AB003,seurat_AB003,cds_day4,top_marker_cluster_day7,top_marker_cluster_day4,seurat_AB002_d4,combined)
gc()

Rdata_filename = paste0(fname_prefix_R, "_monocle_seurat_cds_clustered_RENAMED.RData")
save(cell_metadata, counts,cds,filtered_a,top,meta,original_counts,seurat,
     file = Rdata_filename)

```
####run DEG

```{r}
Idents(seurat) <- "monocle_clusters"
seurat <- FindAllMarkers(seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```





