---
title: "n2 EB003E6: DEG analysis with monocle3 and Seurat"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(readxl)
library(rCASC)
library(deepToolsUtils)
library(R.utils)
library(SparseArray)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data

dir.create(file.path("Output","monocle"))#create forders MONOCLE
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap"))
dir.create(file.path("Output", format(Sys.Date(), "%y%m%d"), "DEG"))


#set up forder path MONOCLE
fname_prefix_R <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output", format(Sys.Date(), "%y%m%d"), "DEG", 
                               format(Sys.Date(), "%y%m%d"))

fname_prefix_scratch="scratch/"

okabe_tab <- read_excel("scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")


pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```

########load data

```{r}
load("scratch/250310_monocle_seurat_cds_clustered_RENAMED.RData")
load("scratch/modules.RData")
#shRNA_IDs <- read_excel("scratch/shRNA_IDs.xlsx")
ID_key=read_excel("scratch/shRNA_IDs.xlsx")
clones=read_excel("scratch/clones.xlsx")

```

##Identify if clusters are too small to be analyzed (mode than 100 cells in at least 1 condition)

```{r}

colData(cds)$shRNA_gene=ifelse(cds$replicate=="SMAD","SMAD2","B2M")
colData(cds)$gene_polish=colData(cds)$shRNA_gene
colData(cds)$KD=colData(cds)$knockdown
colData(cds)$gene_polish=dplyr::recode(colData(cds)$gene_polish,"B2M"="control")
colData(cds)$gene_KD=paste0(cds$gene_polish,"_",cds$KD)
colData(cds)$g_KD=paste0(cds$shRNA_gene,"_",cds$KD)

#$shRNA_gene is a list of genes of the screening "B2M" and "SMAD2" in this case, or "B2M","SMAD2", "GATA4","KMT2D", ecc in another screening
#$KD is a list of "KD" and "CTR" in this case, or "TET" and "no_TET"

meta=as.data.frame(cds@colData@listData)###11275

##################
############ calculate % per cluster of each gene CTR and KD 
################
meta_to_use=meta

tot=table(meta_to_use$g_KD) #total of the cluster
df=table(meta_to_use$g_KD,meta_to_use$monocle_clusters,meta_to_use$shRNA_gene,meta_to_use$KD)
df = as.data.frame(df)
colnames(df)=c("gene_KD","monocle_clusters","gene","KD","freq")

#mdf2 = melt(df)
mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("gene_KD","tot")
mdf2=mdf2%>%left_join(total, by="gene_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)
#m <- m %>%
 # mutate(samp = ifelse(grepl("n3_", sampleID), "n3", "n4"))
by_gene_KD=m

by_gene_KD$monocle_cluster=as.integer(as.character(by_gene_KD$monocle_clusters))

########calculate KD and CTR separate
selection=by_gene_KD%>%select(monocle_clusters,gene_KD,freq)
selection=by_gene_KD%>%select(monocle_cluster,gene_KD,freq)
selection=spread(selection,key=gene_KD,value=freq)
selection <- replace(selection, is.na(selection), 0)
selection$has_100_counts <- apply(selection[, -1] >= 100, 1, any)

stat_filename = paste0(fname_prefix_stat, "_Freq_cells_by_gene_and_KD.csv")
write.csv(selection, stat_filename, quote = F, row.names = T, sep=",")

clust_sel=selection%>%select(monocle_cluster,has_100_counts)
clust_sel=clust_sel%>%dplyr::rename(monocle_clusters=monocle_cluster)
clust_sel$monocle_clusters=as.character(clust_sel$monocle_clusters)
###add selection filter

meta=meta%>%left_join(clust_sel,by="monocle_clusters")
cds@colData$has_100_counts=meta$has_100_counts

meta_cl=subset(meta,meta$has_100_counts==T)#from 11275 to 11239 

cds= cds[,cds@colData@listData[["has_100_counts"]]==T]#from 11275 to 11239 
seurat <- AddMetaData(object = seurat, metadata = meta)
seurat <- subset(seurat, subset = has_100_counts==T)

```

####run DEG

```{r}
Idents(seurat) <- "monocle_clusters"
DEG_all_clusters <- FindAllMarkers(seurat,only.pos = FALSE)
DEG_all_clusters$genes=rownames(DEG_all_clusters)

##cardiac progenitors?
DEG_cl1=FindMarkers(seurat, ident.1 = "1", ident.2 = NULL)
DEG_cl1$genes=rownames(DEG_cl1)

DEG_cl2=FindMarkers(seurat, ident.1 = "2", ident.2 = NULL)
DEG_cl2$genes=rownames(DEG_cl2)

DEG_cl1vs2=FindMarkers(seurat, ident.1 = "1", ident.2 = "2")
DEG_cl1vs2$genes=rownames(DEG_cl1vs2)

DEG_cl11vs1_2=FindMarkers(seurat, ident.1 = "11", ident.2 = c("1","2"))
DEG_cl11vs1_2$genes=rownames(DEG_cl11vs1_2)

DEG_cl11=FindMarkers(seurat, ident.1 = "11", ident.2 = NULL)
DEG_cl11$genes=rownames(DEG_cl11)

DEG_cl11vs4_7=FindMarkers(seurat, ident.1 = "11", ident.2 = c("4","7"))
DEG_cl11vs4_7$genes=rownames(DEG_cl11vs4_7)

#DEG_cl14=FindMarkers(seurat, ident.1 = "14", ident.2 = NULL)
#DEG_cl14$genes=rownames(DEG_cl14)

DEG_cl3=FindMarkers(seurat, ident.1 = "3", ident.2 = NULL)
DEG_cl3$genes=rownames(DEG_cl3)

DEG_cl13=FindMarkers(seurat, ident.1 = "13", ident.2 = NULL)
DEG_cl13$genes=rownames(DEG_cl13)

DEG_cl13vs3vs7=FindMarkers(seurat, ident.1 = "13", ident.2 = c("3","7"))
DEG_cl13vs3vs7$genes=rownames(DEG_cl13vs3vs7)

DEG_cl3vscl7=FindMarkers(seurat, ident.1 = "3", ident.2 = "7")
DEG_cl3vscl7$genes=rownames(DEG_cl3vscl7)

DEG_cl8=FindMarkers(seurat, ident.1 = "8", ident.2 = NULL)
DEG_cl8$genes=rownames(DEG_cl8)

DEG_cl8vs10=FindMarkers(seurat, ident.1 = "8", ident.2 = "10")
DEG_cl8vs10$genes=rownames(DEG_cl8vs10)

DEG_cl3vs9=FindMarkers(seurat, ident.1 = "3", ident.2 = "9")
DEG_cl3vs9$genes=rownames(DEG_cl3vs9)

DEG_cl6vs10=FindMarkers(seurat, ident.1 = "6", ident.2 = "10")
DEG_cl6vs10$genes=rownames(DEG_cl6vs10)

DEG_cl3_9vs6_10=FindMarkers(seurat, ident.1 = c("3","9"), ident.2 = c("6","10"))
DEG_cl3_9vs6_10$genes=rownames(DEG_cl3_9vs6_10)

n=c("DEG_all_clusters","DEG_cl1","DEG_cl2","DEG_cl1vs2","DEG_cl3","DEG_cl13","DEG_cl11","DEG_cl11vs1_2","DEG_cl13vs3vs7","DEG_cl3vscl7","DEG_cl8vs10","DEG_cl8","DEG_cl3vs9","DEG_cl6vs10","DEG_cl3_9vs6_10","DEG_cl11vs4_7")

#for (name in n) {
#stat_filename = paste0(fname_prefix_csv, "_",name,".csv")
#write.csv(get(name), stat_filename, quote = F, row.names = T, sep=",")
#}

if (!require(openxlsx)) install.packages("openxlsx")
library(openxlsx)

for (name in n) {
stat_filename = paste0(fname_prefix_csv, "_",name,".xlsx")
write.xlsx(get(name), file=stat_filename, quote = F, rowNames = T)
}
```

###setup clusters names variables

```{r}

Idents(seurat) <- "sample_id"
seurat_downsample=subset(seurat, downsample = 1972)
m=seurat_downsample[[]]
libID_downsample=m$libID#use libID after downsampling to subset cds file just for plotting an equal nb of cells

colData(cds)$assigned_cell_type <- as.character(clusters(cds))
colData(cds)$assigned_cell_type_short <- as.character(clusters(cds))

colData(cds)$assigned_cell_type <- dplyr::recode(colData(cds)$assigned_cell_type,
                                                 "1"="Cardiac Progenitors-1",#EOMES MYL6 GATA4
                                                 "2"="Cardiac Progenitors-2",#SOX4 BMP4 MYL6 NODAL
                                                 "3"="Late Cardiomyocytes 1",
                                                 "4"="Fibroblasts-1",
                                                 "5"="Proliferative Cardiomyocites",
                                                 "6"="Early Cardiomyocytes-2",
                                                 "7"="Fibroblasts-2",
                                                 "8"="Conductive Cardiomyocites",#low exp of functio
                                                 "9"="Late Cardiomyocytes 2",
                                                 "10"="Early Cardiomyocytes-1",
                                                 "11"="Epicardial-EMT", #SNAI1 https://stemcellres.biomedcentral.com/articles/10.1186/s13287-024-03823-z
                                                 "12"="Endotelial cells",
                                                 "13"="Epicardium")#TBX18 WT1 and CHD1 and endothelial markers
                                                 

colData(cds)$assigned_cell_type_short <- dplyr::recode(colData(cds)$assigned_cell_type_short,
                                                 "1"="CP-1",#SOX4 EOMES MYL6
                                                 "2"="CP-2",#NODAL
                                                 "3"="L-CM-1",
                                                 "4"="CF-1",
                                                 "5"="P-CM",
                                                 "6"="E-CM-2",
                                                 "7"="CF-2",
                                                 "8"="C-CM",
                                                 "9"="L-CM-2",
                                                 "10"="E-CM-1",
                                                 "11"="Epi-EMT",#EGFR FOXD1
                                                 "12"="EC",
                                                 "13"="Epi")

###prepare color/time order vectors
cell_order=c("Cardiac Progenitors-1","Cardiac Progenitors-2","Proliferative Cardiomyocites","Early Cardiomyocytes-1","Early Cardiomyocytes-2","Late Cardiomyocytes 1","Late Cardiomyocytes 2","Conductive Cardiomyocites","Epicardial-EMT","Epicardium","Fibroblasts-1","Fibroblasts-2","Endotelial cells")
cell_order_short=c("CP-1","CP-2","P-CM","E-CM-1","E-CM-2","L-CM-1","L-CM-2","C-CM","Epi-EMT","Epi","CF-1","CF-2","EC")
gene_order=c("B2M","SMAD2")
gene_order_polished=c("control","SMAD2")

####prepare table with cluster assignment

ct=c("Cardiac Progenitors-1","Cardiac Progenitors-2","Late Cardiomyocytes 1","Fibroblasts-1","Proliferative Cardiomyocites","Early Cardiomyocytes-2","Fibroblasts-2","Conductive Cardiomyocites","Late Cardiomyocytes 2","Early Cardiomyocytes-1","Epicardial-EMT","Endotelial cells","Epicardium")

ct_sh=c("CP-1","CP-2","L-CM-1","CF-1","P-CM","E-CM-2","CF-2","C-CM","L-CM-2","E-CM-1","Epi-EMT","EC","Epi")

clust_df=data.frame(monocle_cluster=c(1:13),cell_types=ct,cell_types_abbreviations=ct_sh)

#clust_df$colors=c("#000000","grey","#CC79A7","#f0E442","#920000" ,"#0072B2","#b66dff","#D55E00","#56B4E9","#009E73","#490092","#ffb6db")
clust_df$colors=pal=c("#000000","#004949" ,"#009292" ,"#ff6db6", "#ffb6db", "#490092", "#006ddb", "#b66dff",  "#24ff24", "#b6dbff", "#920000", "#924900", "#db6d00")


###prepare color/time order vectors

monocle_long=data.frame(cell_order=c("Cardiac Progenitors-1","Cardiac Progenitors-2","Conductive Cardiomyocites","Early Cardiomyocytes-1","Early Cardiomyocytes-2","Endotelial cells","Epicardial-EMT","Epicardium","Fibroblasts-1","Fibroblasts-2","Late Cardiomyocytes 1","Late Cardiomyocytes 2","Proliferative Cardiomyocites"),order_LONG=c(1:13))

monocle_short=data.frame(cell_order_short=c("C-CM","CF-1","CF-2","CP-1","CP-2","E-CM-1","E-CM-2","EC","Epi","Epi-EMT","L-CM-1","L-CM-2","P-CM"),order_SHORT=c(1:13))

color_df=data.frame(cell_order_short=cell_order_short,cell_order=cell_order)
color_df$cell_types=color_df$cell_order
color_df$cell_types_abbreviations=color_df$cell_order_short
color_df=color_df%>%left_join(clust_df,by=c("cell_types","cell_types_abbreviations"))

color_df=color_df%>%left_join(monocle_short,by="cell_order_short")%>%left_join(monocle_long,by="cell_order")

color_df$cl_col=color_df$colors
color_df$order=rownames(color_df)

color_df_l=color_df[order(color_df$order_LONG),]
color_df_s=color_df[order(color_df$order_SHORT),]

cds_subset=cds[,cds$libID%in%libID_downsample]


plot_cells(cds_subset, color_cells_by="assigned_cell_type", group_cells_by="assigned_cell_type",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_l$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters.pdf"),
     width = 8, height = 5)

plot_cells(cds_subset, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type_short",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_sh.pdf"),
     width = 8, height = 5)

plot_cells(cds_subset, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type_short",show_trajectory_graph =F,label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")+
  facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_sh_samp.pdf"),
     width = 10, height = 5)

s=cds_subset[,cds_subset$shRNA_gene=="SMAD2"]
plot_cells(s, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type_short",show_trajectory_graph =F,label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")+
  facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_sh_samp_SMAD2.pdf"),
     width = 8, height = 5)

plot_cells(s, color_cells_by="assigned_cell_type", group_cells_by="assigned_cell_type",show_trajectory_graph =F,label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_l$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")+
  facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_samp_SMAD2.pdf"),
     width = 8, height = 5)

#by_gene_KD=by_gene_KD%>%dplyr::left_join(color_df, by="monocle_cluster")

```
######heatmaps with top genes

```{r}

marker_test_res = top_markers(cds, group_cells_by="assigned_cell_type", reference_cells=NULL, cores=16)
marker_test_res <- top_markers(cds, group_cells_by="assigned_cell_type", 
                               reference_cells=10000, cores=8)

top_specific_markers <- marker_test_res %>%
                            filter(fraction_expressing >= 0.10) %>%
                            group_by(cell_group) %>%
                            top_n(10, pseudo_R2)

top_specific_marker_ids <- unique(top_specific_markers %>% dplyr::pull(gene_id))
t10=data.frame(top_specific_markers$gene_id, top_specific_markers$cell_group, stringsAsFactors = F)
colnames(t10)=c("ids", "assigned_cell_type")

write.csv(t10, file= paste0(fname_prefix_csv, "_", "top_10genes_per_cluster_final.csv"), row.names=FALSE)

plot_genes_by_group(cds,
                    top_specific_marker_ids,
                    group_cells_by="assigned_cell_type",
                    ordering_type="maximal_on_diag",
                    max.size=3)+ scale_color_viridis(option="E", discrete=F)
ggsave(filename = paste0(fname_prefix_heatmap, "_", "top10_heatmap.pdf"),
     width = 8, height = 15)
```


###calculate % cells of one gene in each cluster compared to its control

```{r}

meta_to_use=meta_cl

tot=table(meta_to_use$g_KD) #total of the cluster
df=table(meta_to_use$g_KD,meta_to_use$monocle_clusters,meta_to_use$shRNA_gene,meta_to_use$KD)
df = as.data.frame(df)
colnames(df)=c("gene_KD","monocle_cluster","gene","KD","freq")

#mdf2 = melt(df)
mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("gene_KD","tot")
mdf2=mdf2%>%left_join(total, by="gene_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)
#m <- m %>%
 # mutate(samp = ifelse(grepl("n3_", sampleID), "n3", "n4"))
by_gene_KD=m

by_gene_KD$monocle_cluster=as.integer(as.character(by_gene_KD$monocle_cluster))
#color_df$monocle_cluster=as.character(by_gene_KD$monocle_cluster)

by_gene_KD=by_gene_KD%>%dplyr::left_join(color_df, by="monocle_cluster")
by_gene_KD$monocle_cluster=as.character(by_gene_KD$monocle_cluster)

by_gene_KD$sample_ID=paste0(by_gene_KD$KD,"_",by_gene_KD$gene)

#clust_df$monocle_cluster=as.character(clust_df$monocle_cluster)

#by_gene_KD=by_gene_KD%>%dplyr::left_join(clust_df,by="monocle_cluster")

#####graph overview

ggplot(by_gene_KD, aes(x = factor(cell_types_abbreviations,levels=color_df$cell_order_short), y = frxn, fill=cell_types_abbreviations)) + geom_bar(stat = "identity") +theme(legend.position='none')+
  scale_fill_manual(values =color_df_s$cl_col)+
  facet_grid(~KD~factor(gene, levels = gene_order))
  
ylab = "% gene by treatment"
xlab = "cluster"

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "percent_cells_by_gene_clust_polished.pdf"),
     width = 16, height = 15)

####graph by gene

ggplot(by_gene_KD, aes(x = sample_ID, y = frxn, fill=cell_types_abbreviations)) + geom_bar(stat = "identity") +
  scale_fill_manual(name="cell type",values =color_df_s$cl_col)
  
ylab = "% gene by treatment"
xlab = "treatment"
last_plot() + labs(x = xlab, y = ylab)
output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_gene_by treatment_100_percent_cell_type_bis_polished.pdf")
ggsave(filename = output_file, width = 5, height = 5)


####graph by gene

ggplot(data=(subset(by_gene_KD,by_gene_KD$gene=="SMAD2")), aes(x = sample_ID, y = frxn, fill=cell_types_abbreviations)) + geom_bar(stat = "identity") +
  scale_fill_manual(name="cell type",values =color_df_s$cl_col)
  
ylab = "% gene by treatment"
xlab = "treatment"
last_plot() + labs(x = xlab, y = ylab)
output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_gene_by treatment_100_percent_cell_type_bis_polished_SMAD2.pdf")
ggsave(filename = output_file, width = 4, height = 5)

######calculate ratio KD vs CTR
CTR=subset(by_gene_KD,by_gene_KD$KD=="CTR")%>%select(-KD,-sample_ID)
CTR_B2M=subset(CTR,CTR$gene=="B2M")%>%dplyr::rename(n_cells_gene_CTR_B2M=freq,tot_gene_CTR_B2M=tot,percent_gene_CTR_B2M=frxn)%>%select(-gene,-gene_KD)
CTR_SMAD2=subset(CTR,CTR$gene=="SMAD2")%>%dplyr::rename(n_cells_gene_CTR=freq,tot_gene_CTR=tot,percent_gene_CTR=frxn)%>%select(-gene,-gene_KD)

KD=subset(by_gene_KD,by_gene_KD$KD=="TET")%>%dplyr::rename(n_cells_gene_KD=freq,tot_gene_KD=tot,percent_gene_KD=frxn)%>%select(-KD,-sample_ID,-gene)### NB. one cluster missing

by_gene_KD_ratio=CTR_SMAD2%>%dplyr::left_join(KD,by=c(colnames(color_df)))
by_gene_KD_ratio=CTR_B2M%>%dplyr::left_join(by_gene_KD_ratio,by=c(colnames(color_df)))
by_gene_KD_ratio <- replace(by_gene_KD_ratio, is.na(by_gene_KD_ratio), 0.01)

by_gene_KD_ratio=by_gene_KD_ratio%>%mutate(FC=percent_gene_KD/percent_gene_CTR,
                                           FC_B2M=percent_gene_KD/percent_gene_CTR_B2M)

by_gene_KD_ratio$monocle_cluster=as.integer(as.character(by_gene_KD_ratio$monocle_cluster))


#####graph overview FC
ggplot(by_gene_KD_ratio, aes(x = factor(cell_types_abbreviations,levels=color_df$cell_order_short), y = log2(FC), fill=cell_types_abbreviations)) + geom_bar(stat = "identity") +theme(legend.position='none')+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  geom_hline(yintercept=-1, linetype="dashed", color = "red")+
  scale_fill_manual(values =color_df_s$cl_col)#+
  #facet_grid(~factor(gene, levels = gene_order))
  
ylab = "log2 (% gene KD vs CTR)"
xlab = ""

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "FC_cells_by_gene_clust_polished_vsSMAD2CT.pdf"),
     width = 5, height = 4)

ggplot(by_gene_KD_ratio, aes(x = factor(cell_types_abbreviations,levels=color_df$cell_order_short), y = log2(FC_B2M), fill=cell_types_abbreviations)) + geom_bar(stat = "identity") +theme(legend.position='none')+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  geom_hline(yintercept=-1, linetype="dashed", color = "red")+
  scale_fill_manual(values =color_df_s$cl_col)#+
  #facet_grid(~factor(gene, levels = gene_order))
  
ylab = "log2 (% gene KD vs CTR B2M)"
xlab = ""

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "FC_cells_by_gene_clust_polished_vsB2MCT.pdf"),
     width = 5, height = 4)


###save data
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster_polished.csv")
write.csv(by_gene_KD, stat_filename, quote = F, row.names = T, sep=",")
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster_FC_polished.csv")
write.csv(by_gene_KD_ratio, stat_filename, quote = F, row.names = T, sep=",")

Rdata_filename = paste0(fname_prefix_R, "_calculations_nb_percent_polished.RData")
save(by_gene_KD,by_gene_KD_ratio,file = Rdata_filename)

```


#####calculate fisher stats

```{r}
#make true false annotations per cluster to make the contingency tables

colData(cds)$clust1 = cds$monocle_clusters=="1"
colData(cds)$clust2 = cds$monocle_clusters=="2"
colData(cds)$clust3 = cds$monocle_clusters=="3"
colData(cds)$clust4 = cds$monocle_clusters=="4"
colData(cds)$clust5 = cds$monocle_clusters=="5"
colData(cds)$clust6 = cds$monocle_clusters=="6"
colData(cds)$clust7 = cds$monocle_clusters=="7"
colData(cds)$clust8 = cds$monocle_clusters=="8"
colData(cds)$clust9 = cds$monocle_clusters=="9"
colData(cds)$clust10 = cds$monocle_clusters=="10"
colData(cds)$clust11 = cds$monocle_clusters=="11"
colData(cds)$clust12 = cds$monocle_clusters=="12"
colData(cds)$clust13 = cds$monocle_clusters=="13"
#colData(cds)$clust14 = cds$monocle_clusters=="14"

meta=as.data.frame(cds@colData@listData)
meta <- replace(meta, is.na(meta), "empty")
meta$KD <- dplyr::recode(meta$KD,"CTR"="CTR","TET"="KD")

#meta=subset(meta,meta$has_70_counts==T)

#####clusters to consider are 1 to 6
#make a loop to create the statistics
################## start without subsetting for cell populations

gene_list <- c("B2M","SMAD2")
result_list <- list()

meta_sh=meta%>%select(shRNA_gene,KD,clust1,clust2,clust3,clust4,clust5,clust6,clust7,clust8,clust9,clust10,clust11,clust12,clust13)
sh_list=unique(meta_sh$shRNA_gene)

meta_fisher_CTR=data.frame(shRNA_gene=sh_list,KD="CTR",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T,clust10=T,clust11=T,clust12=T,clust13=T)

meta_fisher_KD=data.frame(shRNA_gene=sh_list,KD="KD",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T,clust10=T,clust11=T,clust12=T,clust13=T)

meta_fisher_CTR_F=data.frame(shRNA_gene=sh_list,KD="CTR",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F,clust10=F,clust11=F,clust12=F,clust13=F)

meta_fisher_KD_F=data.frame(shRNA_gene=sh_list,KD="KD",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F,clust10=F,clust11=F,clust12=F,clust13=F)

meta_sh=rbind(meta_sh,meta_fisher_CTR,meta_fisher_KD,meta_fisher_CTR_F,meta_fisher_KD_F)

meta_sh$gKD=paste0(meta_sh$shRNA_gene,"_",meta_sh$KD)
meta_sh=subset(meta_sh,meta_sh$gKD!="B2M_KD")

meta_sh_SMAD2=subset(meta_sh,meta_sh$gKD=="SMAD2_CTR"|meta_sh$gKD=="SMAD2_KD")
meta_sh_B2M=subset(meta_sh,meta_sh$gKD=="B2M_CTR"|meta_sh$gKD=="SMAD2_KD")

gene_list <- c("SMAD2")
result_list <- list()


for (cluster in paste0("clust", 1:13)) {
  p_values <- numeric()

  for (gene in gene_list) {
    subset_data <- meta_sh[meta_sh$shRNA_gene == gene, c("KD", cluster)]
    
    if (nrow(subset_data) > 0) {
      # Perform chi-squared test
      fisher_result <- fisher.test(subset_data$KD, subset_data[[cluster]], simulate.p.value = FALSE)
      
      # Extract p-value
      p_value <- fisher_result[["p.value"]]
      
      p_values <- c(p_values, p_value)
    } else {
      # If no data for the gene in the current cluster, set p-value to NA
      p_values <- c(p_values, NA)
    }
  }

  result_df <- data.frame(gene = gene_list, cluster = cluster, p_value = p_values, stringsAsFactors = FALSE)
  
  # Add a q-value column using p.adjust
  result_df$q_value <- p.adjust(result_df$p_value, method = "fdr")
  
  # Add a column indicating significance at a 5% level
  result_df$sig <- result_df$q_value <= 0.05

  result_list[[cluster]] <- result_df
}

combined_result_df <- do.call(rbind, result_list)

by_gene_KD_ratio$cluster=paste0("clust",by_gene_KD_ratio$monocle_cluster)
FC=by_gene_KD_ratio
combined_result_df=combined_result_df%>%select(-gene)
combined_result_df=combined_result_df%>%dplyr::rename(p_value_SMAD2=p_value,sig_SMAD2=sig,q_value_SMAD2=q_value)

combined_result_df=combined_result_df%>%left_join(FC, by=c("cluster"))

stat_filename = paste0(fname_prefix_stat, "_fisher_stat_SMAD2_clusters.csv")
write.csv(combined_result_df, stat_filename, quote = F, row.names = T, sep=",")


for (cluster in paste0("clust", 1:13)) {
  p_values <- numeric()

    subset_data <- meta_sh_B2M[, c("KD", cluster)]
    
    if (nrow(subset_data) > 0) {
      # Perform chi-squared test
      fisher_result <- fisher.test(subset_data$KD, subset_data[[cluster]], simulate.p.value = FALSE)
      
      # Extract p-value
      p_value <- fisher_result[["p.value"]]
      
      p_values <- c(p_values, p_value)
    } else {
      # If no data for the gene in the current cluster, set p-value to NA
      p_values <- c(p_values, NA)
    }

  result_df <- data.frame(cluster = cluster, p_value = p_values, stringsAsFactors = FALSE)
  
  # Add a q-value column using p.adjust
  result_df$q_value <- p.adjust(result_df$p_value, method = "fdr")
  
  # Add a column indicating significance at a 5% level
  result_df$sig <- result_df$q_value <= 0.05

  result_list[[cluster]] <- result_df
}


combined_result_df_B2M <- do.call(rbind, result_list)

by_gene_KD_ratio$cluster=paste0("clust",by_gene_KD_ratio$monocle_cluster)
combined_result_df_B2M=combined_result_df_B2M%>%dplyr::rename(p_value_B2M=p_value,sig_B2M=sig,q_value_B2M=q_value)

combined_result_df=combined_result_df_B2M%>%left_join(combined_result_df, by=c("cluster"))

stat_filename = paste0(fname_prefix_stat, "_fisher_stat_B2M_SMAD2_clusters.csv")
write.csv(combined_result_df, stat_filename, quote = F, row.names = T, sep=",")

#sub=subset(clust_sel,clust_sel$has_70_counts==T)
#color_df_select=color_df%>%filter(cell_order_short%in%sub$assigned_cell_type_short)

ggplot(data=combined_result_df, aes(x=log2(as.numeric(FC)), y=log10(q_value_SMAD2)*-1)) +
  geom_point(aes(colour = factor(cell_types_abbreviations,levels=color_df_s$cell_order_short)), size=1.5) +scale_colour_manual(values=color_df_s$colors)+
  ggrepel::geom_text_repel(
    data = subset(combined_result_df,sig_SMAD2==T&FC>2|sig_SMAD2==T&FC<0.5),
    aes(label = cell_types_abbreviations),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  geom_vline(xintercept = log2(2),linetype="dotted")+
  geom_vline(xintercept = -log2(2),linetype="dotted")+
  #geom_hline(yintercept = -log10(0.05),linetype="dotted")+
  ggtitle("fisher test statistics in each cluster compared to SMAD2 CT")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2FC compared to CTR",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_fisher_stats.pdf"),
       width = 5, height = 4)

ggplot(data=combined_result_df, aes(x=log2(as.numeric(FC_B2M)), y=log10(q_value_B2M)*-1)) +
  geom_point(aes(colour = factor(cell_types_abbreviations,levels=color_df_s$cell_order_short)), size=1.5) +scale_colour_manual(values=color_df_s$colors)+
  ggrepel::geom_text_repel(
    data = subset(combined_result_df,sig_B2M==T&FC>2|sig_B2M==T&FC<0.5),
    aes(label = cell_types_abbreviations),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  geom_vline(xintercept = log2(2),linetype="dotted")+
  geom_vline(xintercept = -log2(2),linetype="dotted")+
  #geom_hline(yintercept = -log10(0.05),linetype="dotted")+
  ggtitle("fisher test statistics in each cluster compared to B2M CT")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2FC compared to B2M CTR",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_fisher_stats_B2M.pdf"),
       width = 5, height = 4)


#cds_subset=cds[,cds$has_70_counts==T]

plot_cells(cds, color_cells_by="assigned_cell_type", group_cells_by="assigned_cell_type",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8, show_trajectory_graph =F)+
  scale_color_manual(name="monocle_cluster",values=color_df_l$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_long.pdf"),
     width = 8, height = 5)

plot_cells(cds, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type_short",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8, show_trajectory_graph =F)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_short.pdf"),
     width = 8, height = 5)

```
#####analyze modules
####selected top 16 in n1 #filtered_a

```{r}
top_16=top%>%select(-max,-delta,-min)

hp3=pheatmap::pheatmap(select_modules,
                   scale="column", clustering_method="ward.D2",main="Top 16 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on row (modules)

Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_top16_Turbo.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp3
dev.off()
```
##KS stats
```{r}
analyze=data.frame(t(select_modules))
analyze$module=rownames(analyze)
analyze$module_nb=sapply(strsplit(as.character(analyze$mod), "_"), "[[", 2)
analyze$mod=paste0("Module ",analyze$module_nb)
rownames(analyze)=analyze$module
analyze$pal=c("#AA0DFE","#3283FE","#85660D","#782AB6","#565656","#1C8356","#16FF32","#F7E1A0","#E2E2E2","#1CBE4F","#C4451C","#DEA0FD","#FE00FA","#325A9B","#FEAF16","black")

#row=data.frame(module=rownames(top15_norm))
#rownames(row)=row$module

my_colour = list(module = c("Module 15"="#AA0DFE","Module 59"="#3283FE","Module 27"="#85660D","Module 41"="#782AB6","Module 2"="#565656","Module 25"="#1C8356","Module 37"="#16FF32","Module 56"="#F7E1A0","Module 39"="#E2E2E2","Module 73"="#1CBE4F","Module 17"="#C4451C","Module 48"="#DEA0FD","Module 62"="#FE00FA","Module 44"="#325A9B","Module 60"="#FEAF16","Module 65"="black"))


plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(analyze$module_nb)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules_top_15.pdf"),
     width = 10, height = 8)

analysis=c(analyze$module)

TET = subset(meta, meta$knockdown == "TET")
CTR = subset(meta, meta$knockdown == "CTR")

ks_stat = data.frame()

for (c in unique(meta$shRNA_gene)) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_gene == "SMAD2", analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == c, analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    
    ks_stat = rbind(ks_stat, data.frame(gene = c, pval = ks_pval, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("gene", "pval", "analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.01

ks_stat_filtered_modules=ks_stat

write.csv(ks_stat_filtered_modules, file= paste0(fname_prefix_csv, "_", "ks_statistics_filtered_modules.csv"), row.names=FALSE)


#######see modules that stands out
analysis=c(analyze$module)

TET = subset(meta, meta$knockdown == "TET")
CTR = subset(meta, meta$knockdown == "CTR")

ks_stat = data.frame()

for (c in unique(meta$shRNA_gene)) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_gene == "SMAD2", analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == c, analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    ks_pval_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["p.value"]]
    ks_sig_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["statistic"]][["D^+"]]
    ks_pval_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["p.value"]]
    ks_sig_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["statistic"]][["D^-"]]
    ks_sig = ks.test(tmp_TET, tmp_CTR)[["statistic"]][["D"]]
    
    ks_stat = rbind(ks_stat, data.frame(gene = c, pval = ks_pval, pval_g = ks_pval_g,pval_l = ks_pval_l,ks_sign_g=ks_sig_g,ks_sign_l=ks_sig_l,ks_sign=ks_sig, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("gene", "pval", "pval_greater","pval_less","ks_greater","ks_less","ks_sign","analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.01
ks_stat$greater <- ks_stat$pval_greater <= 0.01
ks_stat$less <- ks_stat$pval_less <= 0.01

ks_stat$k_sign_summary <- ifelse(
  # "greater" is TRUE and "less" is FALSE
  ks_stat$greater & !ks_stat$less,
  as.numeric(ks_stat$ks_greater),  # Return ks_sign value
  ifelse(
    # "greater" is FALSE and "less" is TRUE
    !ks_stat$greater & ks_stat$less,
    -as.numeric(ks_stat$ks_less),  # Return negative of ks_sign
    ks_stat$ks_sign  # Otherwise, return ks sign
  )
)

ks_stat$k_colour <- ifelse(
  ks_stat$sig==T,ks_stat$analysis,"")
 
k_stat_all=ks_stat
write.csv(ks_stat, file= paste0(fname_prefix_csv, "_", "ks_statistics_sign_top15.csv"), row.names=FALSE)

k_stat_all=k_stat_all%>%dplyr::rename(module=analysis)
k_stat_all=k_stat_all%>%left_join(analyze, by="module")

k_stat_SMAD=subset(k_stat_all,k_stat_all$gene=="SMAD2")
k_stat_B2M=subset(k_stat_all,k_stat_all$gene=="B2M")
```

#####volcano plot statistics
###genes

```{r}
library(pals)

a=k_stat_SMAD%>%filter(k_stat_SMAD$sig==T)%>%dplyr::select(module,pal)%>%distinct(module,pal)
a=a[order(a$module),]

ggplot(data=k_stat_SMAD, aes(x=k_sign_summary, y=log10(padj)*-1)) +
  geom_point(aes(colour = k_colour), size=1.5) +scale_color_manual(values=c("grey",a$pal))+
  ggrepel::geom_text_repel(
    data = subset(k_stat_SMAD, sig==T&k_sign_summary>0.1|sig==T&k_sign_summary<=-0.1),
    aes(label = module_nb),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  ggtitle("ks test statistics per gene vs its control")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "KS statistic summary",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_ks_stats_gene.pdf"),
       width = 6, height = 4)

```
######get info about gene modules

```{r}

mod_sig = subset(k_stat_SMAD, sig==T&k_sign_summary>0.1|sig==T&k_sign_summary<=-0.1)


gene_modules=list()

for (c in unique(gene_module_df$module)) {
 
tmp=subset(gene_module_df, gene_module_df$module==c)
gene_list=tmp$id

gene_modules[[c]]=gene_list
}

library(topGO)

library(org.Hs.eg.db)

hs <- org.Hs.eg.db

modules=c(mod_sig$module_nb)

go.de_list <- list()
k_list <- list()

# Iterate over each module in 'modules'
for (module in modules) {
  # Convert gene symbols to ENSEMBL or ENTREZID
  gene.df <- clusterProfiler::bitr(gene_modules[[module]], fromType = "SYMBOL",
                                    toType = c("ENSEMBL", "ENTREZID"),
                                    OrgDb = org.Hs.eg.db)
  
  # Perform GO analysis
  go.de <- goana(list(DE1 = gene.df$ENTREZID), species = "Hs", convert = TRUE) %>%
    mutate(module = module)
  
  # Perform KEGG analysis
  k <- kegga(list(DE1 = gene.df$ENTREZID), species = "Hs", convert = TRUE) %>%
    mutate(module = module)
  
  # Store the results in lists
  go.de_list[[module]] <- go.de
  k_list[[module]] <- k
}

# Combine all data frames in the lists into single data frames
GO <- do.call(rbind, go.de_list)%>%filter(Ont!="CC",DE1>=20,P.DE1<=0.05)
k <- do.call(rbind, k_list)%>%filter(DE1>=6,P.DE1<=0.05)

write.csv(GO, file= paste0(fname_prefix_csv, "_", "GO_modules.csv"), row.names=FALSE)
write.csv(k, file= paste0(fname_prefix_csv, "_", "kegg_modules.csv"), row.names=FALSE)

```

####save data

```{r}
meta=as.data.frame(cds@colData@listData)

Rdata_filename = paste0(fname_prefix_R, "_downsampledata.RData")
save(seurat_downsample,libID_downsample,cds_subset,cds,seurat,
     file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_cds_updated.RData")
save(meta,cds,clust_df,color_df,color_df_l,color_df_s,select,seurat,meta_cl,meta_sh,
     file = Rdata_filename)
```



