---
title: "n1 EB003E1: QC and monocle analysis"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(SparseArray)
#library(rCASC)
#library(deepToolsUtils)
#library(R.utils)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
#dir.create(file.path("Output","scratch"))

dir.create(file.path("Output","monocle"))#create forders MONOCLE
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap"))

fname_scratch <- file.path("scratch")

#set up forder path MONOCLE
fname_prefix_R <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","monocle", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))


okabe_tab <- read_excel("scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")


pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```

###load data
```{r}
load(paste0(getwd(),"/scratch/241205_data_filtered_for_monocle.RData"))
libID_EB003E1_E2=read.csv(paste0(getwd(),"/scratch/241210_libID_EB003E1_2.csv"), header = TRUE, sep = ",", quote = "\"", 
                     dec = ".", fill = TRUE, comment.char = "",row.names=1)

# get side information
clone_side=read.csv(paste0(getwd(),"/scratch/clone_side.csv"), header = TRUE, sep = ",", quote = "\"", 
                     dec = ".", fill = TRUE, comment.char = "",row.names=1)
#get colour scale
shRNA_IDs <- read_excel("scratch/shRNA_IDs.xlsx")
```

#add side data and shRNA data

```{r}
##add side data and shRNAdata
cell_metadata=cell_metadata%>%left_join(clone_side, by="clone")#7603
cell_metadata <- cell_metadata %>%
  mutate(side = ifelse(is.na(side), "unknown", side))

cell_metadata=cell_metadata%>%left_join(shRNA_IDs, by=c("shRNA_BC","shRNA_gene"))#7603
rownames(cell_metadata)=cell_metadata$info

libID_EB003E1_E2=libID_EB003E1_E2%>%left_join(clone_side, by="clone")#7603
libID_EB003E1_E2 <- libID_EB003E1_E2 %>%
  mutate(side = ifelse(is.na(side), "unknown", side))

libID_EB003E1_E2=libID_EB003E1_E2%>%left_join(shRNA_IDs, by=c("shRNA_BC","shRNA_gene"))#8767

```

####preprocess monocle data

```{r}

###create cds file
cds <- new_cell_data_set(counts,cell_metadata = cell_metadata,
                                      gene_metadata = gene_annotation)
cds <- preprocess_cds(cds, num_dim = 50)
#cds <- align_cds(cds, num_dim = 100, alignment_group = "sample")
plot_pc_variance_explained(cds)

ggsave(filename = paste0(fname_prefix_dotplot, "_", "variance.png"),
     width = 12, height = 8)

#cds <- align_cds(cds, alignment_group = "experiment")
set.seed(678686)
cds = reduce_dimension(cds,reduction_method="tSNE")
set.seed(79876858)
cds = reduce_dimension(cds,umap.min_dist = 0.2,reduction_method="UMAP")
#cds = reduce_dimension(cds,umap.min_dist = 0.5,reduction_method="UMAP")

#monocle.object = order_cells(monocle.object, reduction_method = "UMAP")
set.seed(12345)
cds_1 = cluster_cells(cds, resolution = 1e-2)
colData(cds_1)$monocle_clusters = as.character(monocle3::clusters(cds_1))
set.seed(123479870)
cds_2 = cluster_cells(cds, resolution = 1e-4)
colData(cds_2)$monocle_clusters = as.character(monocle3::clusters(cds_2))
set.seed(123457768)
cds_3 = cluster_cells(cds, resolution = 8e-4)
colData(cds_3)$monocle_clusters = as.character(monocle3::clusters(cds_3))
set.seed(123458980)
cds_4 = cluster_cells(cds, resolution = 2e-4)
colData(cds_4)$monocle_clusters = as.character(monocle3::clusters(cds_4))
set.seed(1234597698)
cds_5 = cluster_cells(cds, resolution = 5e-4)
colData(cds_5)$monocle_clusters = as.character(monocle3::clusters(cds_5))


plot_cells(cds_1, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_2, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_3, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_4, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
plot_cells(cds_5, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)

Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered.RData")
save(cell_metadata, counts,cds, cds_1,
     cds_2, cds_3, cds_4, cds_5,
     file = Rdata_filename)
```

###chose the cluster iteration with better results and clean environment

```{r}
plot_cells(cds_5, color_cells_by="cluster", group_cells_by="cluster",group_label_size = 7, cell_size=0.8)
last_plot()+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP.pdf"),
     width = 18, height = 8)

#map by seurat clusters generated
cds_5 <- learn_graph(cds_5)
cds=cds_5#this is the iteration we decided to use

#cds_cardio= cds[,cds@colData@listData[["monocle_clusters"]]=="5"|cds@colData@listData[["monocle_clusters"]]=="4"|cds@colData@listData[["monocle_clusters"]]=="3"|cds@colData@listData[["monocle_clusters"]]=="2"]


Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,geneInfo,libID_EB003E1_E2,
     file = Rdata_filename)

#clean environment
rm(combined_exp,data_exp5,expression_data,gene_annotation,expression_matrix,cds_1,cds_2,cds_3,cds_4,cds_5)
```

###select cds with protein coding genes
#add umap data on cds meta
#add pseudotime info

```{r}
#add UMAP data to cell_metadata
pbuild <- plot_cells(cds, color_cells_by = "monocle_clusters", label_cell_groups=T, show_trajectory_graph = F, label_leaves=F, label_branch_points=F, graph_label_size=3, group_label_size = 4, cell_size = 1) 
UmapCDS <- data.frame(pbuild$data$sample_name,UMAP1_monocle = pbuild$data$data_dim_1, UMAP2_monocle = pbuild$data$data_dim_2, row.names = 1) 
colData(cds)$UMAP_1_monocle=pbuild$data$data_dim_1 #add umap data for further analysis outside monocle
colData(cds)$UMAP_2_monocle=pbuild$data$data_dim_2 #add umap data for further analysis outside monocle

original_counts=counts
#counts=as.data.frame(exprs(cds))
counts=as.data.frame(normalized_counts(cds))
meta=as.data.frame(cds@colData@listData)

#cds<- order_cells(cds,reduction_method = "UMAP")#selected around cluster 3

#plot_cells(cds,
 #          color_cells_by = "pseudotime",
    #       label_cell_groups=FALSE,
   #        label_leaves=FALSE,
  #         label_branch_points=FALSE,
 #          graph_label_size=1.5)
#last_plot()+facet_grid(~sample)
#ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_pseudotime.pdf"),
 #    width = 18, height = 8)

Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,geneInfo,libID_EB003E1_E2,meta,original_counts,
     file = Rdata_filename)


```

##make graphs to visualize genes

```{r}

neuro_genes=c("PAX6","ZIC1","ZIC4","SIX3","GBX2","OLIG3","NEUROG3","NEUROD6","FOXC1","ZNF37A","BAHCC1")

plot_cells(cds,
           genes = c("PAX6","ZIC1","ZIC4","SIX3","OLIG3","NEUROG3","FOXC1","ZNF37A","BAHCC1"),
            label_cell_groups = FALSE,
            show_trajectory_graph = FALSE)+ 
     labs(x = "UMAP 1", y = "UMAP 2", title = "")+
     scale_color_viridis(option="E", discrete=F)
 
 ggsave(filename = paste0(fname_prefix_dotplot, "_", "UMAP_neuro.pdf"),
     width = 10, height = 10)

```

###make graphs for clusters and features

```{r}
ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = monocle_clusters),
               size=0.8)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clusters")+
  scale_color_manual(name = "monocle",
                        values=c(pal))+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_monocle_clusters.pdf"),
     width = 12, height = 6)


ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = side),
               size=0.8)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Side")+
  scale_color_manual(name = "monocle",
                        values=c(okabe_pal))+facet_grid(~sample_id)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_side.pdf"),
     width = 12, height = 6)
  
plot_feature=c("percent.mt", "percent.ribo","nFeature_RNA","nCount_RNA")

for (feature in plot_feature) {
  
  p <- ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes_string(color = feature), size = 0.8) +
    labs(
      x = "UMAP 1",
      y = "UMAP 2",
      title = feature
    ) +
    scale_color_viridis(option = "E", discrete = FALSE) +
    facet_grid(~sample_id)
  
  ggsave(
    filename = paste0(fname_prefix_UMAP, "_", "UMAP_", feature, ".pdf"),
    plot = p,
    width = 8,
    height = 4
  )
}

plot_shRNA=unique(meta$shRNA_gene)
for (gene in plot_shRNA) {
  
  s <- subset(meta, meta$shRNA_gene == gene)
  
  p <- ggplot(data = s, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = monocle_clusters),
               size = 0.8) +
    labs(x = "UMAP 1", y = "UMAP 2", title = gene) +
    scale_color_manual(name = "monocle",
                       values = c(pal)) +
    facet_grid(~knockdown ~ side)
  
  ggsave(
    filename = paste0(fname_prefix_UMAP, "_", "UMAP_", gene, "_side.pdf"),
    plot = p,
    width = 12,
    height = 6
  )
}

plot_shRNA=unique(meta$shRNA_gene)
for (gene in plot_shRNA) {
  
  s <- subset(meta, meta$shRNA_gene == gene)
  
  p <- ggplot(data = s, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = monocle_clusters),
               size = 0.8) +
    labs(x = "UMAP 1", y = "UMAP 2", title = gene) +
    scale_color_manual(name = "monocle",
                       values = c(pal)) +
    facet_grid(~knockdown)
  
  ggsave(
    filename = paste0(fname_prefix_UMAP, "_", "UMAP_", gene, ".pdf"),
    plot = p,
    width = 8,
    height = 4
  )
}

```
###get top genes in a loop 

```{r}
marker_test_res = top_markers(cds, group_cells_by="monocle_clusters", reference_cells=NULL, cores=16)
marker_test_res <- top_markers(cds, group_cells_by="monocle_clusters", 
                               reference_cells=10000, cores=8)

# List of top_n values to iterate through
top_n_values <- c(2, 5, 10, 25)

# Loop through each top_n value
for (n in top_n_values) {
  
  # Filter markers and select the top_n genes
  top_specific_markers <- marker_test_res %>%
    filter(fraction_expressing >= 0.10) %>%
    group_by(cell_group) %>%
    top_n(n, pseudo_R2)
  
  # Extract unique gene IDs
  top_specific_marker_ids <- unique(top_specific_markers %>% dplyr::pull(gene_id))
  
  # Create a data frame for the top genes
  t_data <- data.frame(
    ids = top_specific_markers$gene_id, 
    cluster = top_specific_markers$cell_group, 
    stringsAsFactors = FALSE
  )
  
  # Write the data to a CSV file
  write.csv(
    t_data, 
    file = paste0(fname_prefix_csv, "_", "top_", n, "genes_per_cluster.csv"), 
    row.names = FALSE
  )
  
  # Generate a heatmap plot
  p <- plot_genes_by_group(
    cds,
    top_specific_marker_ids,
    group_cells_by = "monocle_clusters",
    ordering_type = "maximal_on_diag",
    max.size = 3
  ) + scale_color_viridis(option = "E", discrete = FALSE)
  
  # Save the plot
  ggsave(
    filename = paste0(fname_prefix_mosaic, "_", "top", n, "_heatmap.pdf"),
    plot = p,
    width = 8, 
    height = ifelse(n == 25, 15, 8) # Adjust the height for larger plots
  )
}
```


##get top 10 genes per cluster

```{r}

marker_test_res = top_markers(cds, group_cells_by="monocle_clusters", reference_cells=NULL, cores=16)
marker_test_res <- top_markers(cds, group_cells_by="monocle_clusters", 
                               reference_cells=10000, cores=8)

top_specific_markers <- marker_test_res %>%
                            filter(fraction_expressing >= 0.10) %>%
                            group_by(cell_group) %>%
                            top_n(10, pseudo_R2)

top_specific_marker_ids <- unique(top_specific_markers %>% dplyr::pull(gene_id))
t10=data.frame(top_specific_markers$gene_id, top_specific_markers$cell_group, stringsAsFactors = F)
colnames(t10)=c("ids", "cluster")

write.csv(t10, file= paste0(fname_prefix_csv, "_", "top_10genes_per_cluster.CSV"), row.names=FALSE)

plot_genes_by_group(cds,
                    top_specific_marker_ids,
                    group_cells_by="monocle_clusters",
                    ordering_type="maximal_on_diag",
                    max.size=3)+ scale_color_viridis(option="E", discrete=F)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "top10_heatmap.pdf"),
     width = 8, height = 15)

```

###add modules

###get gene modules shanging along pseudotime

```{r}

cds@colData$group=paste0(cds@colData$shRNA_gene,"_",cds@colData$knockdown)
cell_metadata$group=paste0(cell_metadata$shRNA_gene,"_",cell_metadata$knockdown)

#########find gene modules along trajectory
set.seed(685875890)
cds_pr_test_res <- graph_test(cds, neighbor_graph="knn", cores=4)#score genes not based on trajectory
genes_mod <- row.names(subset(cds_pr_test_res, q_value < 0.05))

gene_module_df <- find_gene_modules(cds[genes_mod,], reduction_method = c("UMAP"),resolution=0.01)#organize genes into gene modules

cell_group_df <- tibble::tibble(cell=row.names(colData(cds)), 
                                cell_group=colData(cds)$monocle_clusters)###group modules by cluster
cell_group_df_test <- tibble::tibble(cell=row.names(colData(cds)), 
                                cell_group=colData(cds)$group)###group modules by shRNA hene and KD (no selection yet)

agg_mat <- aggregate_gene_expression(cds, gene_module_df)###every lib ID has a value per each module

modules=as.data.frame(t(agg_mat))
#modules <- modules[, order(names(modules))]
modules <- modules[, order(as.numeric(names(modules)))]
modules=as.data.frame(modules)

cds@colData[, paste0("module_", 1:87)] <- modules[, 1:87]###added 85 modules
cell_metadata[, paste0("module_", 1:87)] <- modules[, 1:87]

agg_mat_cluster <- aggregate_gene_expression(cds, gene_module_df, cell_group_df)
row.names(agg_mat_cluster) <- stringr::str_c("Module ", row.names(agg_mat))
hp1=pheatmap::pheatmap(agg_mat_cluster,
                   scale="row", clustering_method="ward.D2")

Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_cl.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()

agg_mat_test <- aggregate_gene_expression(cds, gene_module_df, cell_group_df_test)
row.names(agg_mat_test) <- stringr::str_c("Module ", row.names(agg_mat_test))
hp2=pheatmap::pheatmap(agg_mat_test,
                   scale="row", clustering_method="ward.D2",main="row normalized modules all",color = viridis(n = 256, alpha = 1,begin = 0, end = 1, option = "H"))

Rdata_filename = paste0(fname_prefix_mosaic, "_hp_all_modules.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp2
dev.off()

##B2M and ChD7 look different than the others. I believe this is because CHD7 is mainly side 1 and B2M cells all disappear

Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_with_modules_FINAL.RData")
save(cell_metadata, counts,cds,geneInfo,libID_EB003E1_E2,meta,original_counts,modules,gene_module_df,cds_pr_test_res,genes_mod,
     agg_mat,agg_mat_cluster,agg_mat_test,
     file = Rdata_filename)

#update monocle with clusters without all the modules needed for plotting
Rdata_filename = paste0(fname_prefix_R, "_monocle_cds_clustered_FINAL.RData")
save(cell_metadata, counts,cds,geneInfo,libID_EB003E1_E2,meta,original_counts,
     file = Rdata_filename)
```

####plot modules

```{r}

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(1:87)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle top modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

#ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules.pdf"),
 #    width = 21, height = 30)

```

####module selection


```{r}
####eliminate the modules with expression in only few cells abd general low expression
select_modules=as.data.frame(agg_mat_test)
select_modules=select_modules%>%dplyr::select(-B2M_CTR,-B2M_TET)####eliminate B2M because is making a bit of a mess (too few cells)
select_modules=select_modules%>%dplyr::filter(rownames(select_modules)!="Module 11"&rownames(select_modules)!="Module 27"&rownames(select_modules)!="Module 45"&rownames(select_modules)!="Module 55"&rownames(select_modules)!="Module 57"&rownames(select_modules)!="Module 63"&rownames(select_modules)!="Module 67"&rownames(select_modules)!="Module 69"&rownames(select_modules)!="Module 72"&rownames(select_modules)!="Module 76"&rownames(select_modules)!="Module 78"&rownames(select_modules)!="Module 80"&rownames(select_modules)!="Module 81"&rownames(select_modules)!="Module 83"&rownames(select_modules)!="Module 84"&rownames(select_modules)!="Module 85"&rownames(select_modules)!="Module 86"&rownames(select_modules)!="Module 87")#####select modules with decent amount of cells expressing genes

hp1=pheatmap::pheatmap(select_modules,
                   scale="row", clustering_method="ward.D2",main="row normalized modules filtered")#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_filtered.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()
hp2=pheatmap::pheatmap(agg_mat_test,
                   scale="row", clustering_method="ward.D2")#scaled on row (modules) ###original data >>>to compare modules between perturbations

hp1=pheatmap::pheatmap(select_modules,
                   scale="row", clustering_method="ward.D2",main="row normalized modules filtered",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_filtered_TURBO.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()


#######get z score of the modules (the values of the heatmap hp1 and hp2)

cal_z_score <- function(x){
  (x - mean(x)) / sd(x)
}

data_subset_norm <- t(apply(select_modules, 1, cal_z_score)) ###get z scored data
hp3=pheatmap::pheatmap(data_subset_norm, clustering_method="ward.D2")###check hp4 is the same as hp2

data_subset_norm=as.data.frame(data_subset_norm)
data_subset_norm <- data_subset_norm%>%select(-CHD7_CTR,-CHD7_TET) #these created a byas due to the side
data_subset_norm$min<-apply(data_subset_norm,1,FUN=min)
data_subset_norm$max<-apply(data_subset_norm,1,FUN=max)
data_subset_norm$delta=data_subset_norm$max-data_subset_norm$min

top=data_subset_norm[order(-data_subset_norm$delta),] ####select top 16 based on delta
top=head(top, n=16)

select_modules=select_modules%>%dplyr::filter(rownames(select_modules)%in%rownames(top))

rownames(top)

#"Module 19" "Module 26" "Module 5"  "Module 53" "Module 56" "Module 49" "Module 61" "Module 1"  "Module 23" "Module 9"  "Module 13" "Module 15" "Module 47" "Module 33"
#"Module 58" "Module 66"

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(1,5,9,13,15,19,23,26,33,47,49,53,56,58,61,66)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle top 16 modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_top_16_modules.pdf"),
     width = 10, height = 10)

##make the same hm as before but flipped
select_modules=t(select_modules)#flip data do have genes in rows and modules in columns

hp3=pheatmap::pheatmap(select_modules,
                   scale="column", clustering_method="ward.D2",main="Top 16 modules")#scaled on row (modules)
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_top16.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp3
dev.off()

hp3=pheatmap::pheatmap(select_modules,
                   scale="column", clustering_method="ward.D2",main="Top 16 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on row (modules)
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_top16_Turbo.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp3
dev.off()

module_final_selection=c("Module 19","Module 26","Module 5","Module 53","Module 56","Module 49","Module 61","Module 1","Module 23","Module 9","Module 13", "Module 15", "Module 47","Module 33","Module 58","Module 66")

select_modules_FINAL=as.data.frame(agg_mat_test)
select_modules_FINAL=select_modules_FINAL%>%dplyr::select(-B2M_CTR,-B2M_TET)
select_modules_FINAL=select_modules_FINAL%>%dplyr::filter(rownames(select_modules_FINAL)%in%module_final_selection)
select_modules_FINAL=t(select_modules_FINAL)
hp2=pheatmap::pheatmap(select_modules_FINAL,
                   scale="column", clustering_method="ward.D",main="Top selected modules")#scaled on row (modules)

Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_selected_FINAL_norm_ROW.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp2
dev.off()

hp2=pheatmap::pheatmap(select_modules_FINAL,
                   scale="column", clustering_method="ward.D",main="Top selected modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on row (modules)

Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_selected_FINAL_norm_ROW_TURBO.pdf")
pdf(Rdata_filename, height = 4, width = 4)
hp2
dev.off()


Rdata_filename = paste0(fname_prefix_R, "_updated_cds_with_modules_SELECTION.RData")
save(cell_metadata, counts,cds,geneInfo,libID_EB003E1_E2,meta,original_counts,modules,gene_module_df,cds_pr_test_res,genes_mod,
     agg_mat,agg_mat_cluster,agg_mat_test,modules,top,select_modules,select_modules_FINAL,gene_module_df,cds_pr_test_res,
     file = Rdata_filename) 


```



