---
title: "#A3 EB003E1 neural organoids E/D"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(rCASC)
library(deepToolsUtils)
library(R.utils)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
#dir.create(file.path("Output","scratch"))

dir.create(file.path("Output","E_D"))#create forders E_D (enrichment and depletion)
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "heatmap"))

fname_prefix_scratch="/scratch"
#set up forder path E_D
fname_prefix_R <- file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","E_D", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))


okabe_tab=read_excel("scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")

pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```


########load data

```{r}

load("scratch/241210_monocle_seurat_cds_clustered_FINAL.RData")#libID_EB003E1_E2
load("scratch/241211_clone_order.RData")
#shRNA_IDs <- read_excel("scratch/shRNA_IDs.xlsx")
ID_key=read_excel("scratch/shRNA_IDs.xlsx")
#get clones info
clones=read.csv(paste0(getwd(),"/scratch/all_clones.csv"), header = TRUE, sep = ";")
rownames(clones)=clones$clone
clones=clones%>%select(clone,has_10_counts,shRNA_gene,cell_BC,shRNA_ID_unified,cl_ID,order.shRNA,order.genes)

```

####### fisher test enrichment depletion EB003E1 only
###ER between neural organoids with and without TET

```{r}
#####total Fisher test
m_cl=cell_metadata%>%dplyr::select(clone,KD)
clone_list=unique(m_cl$clone)#52


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$KD, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=cell_metadata %>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = KD, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=cell_metadata%>%
  dplyr::group_by(KD)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_CTR=totals[1,2]
counts$tot_KD=totals[2,2]
counts$TOT=counts$tot_CTR+counts$tot_KD
counts$TOT_clone=counts$CTR+counts$KD

counts$tot_other_CTR=counts$tot_CTR-counts$CTR
counts$tot_other_KD=counts$tot_KD-counts$KD
counts=counts%>%mutate(FC=(KD/tot_KD)/(CTR/tot_CTR),
                       percent_CTR=CTR/tot_CTR*100,
                       percent_KD=KD/tot_KD*100,
                       percent_cl=TOT_clone/TOT*100)
side=cell_metadata %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones,by=c("clone"))
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_AB03_all.csv"), row.names=FALSE)
ED_EB003E1_all_results=result_ED_TOTAL

ED_EB003E1_all_results=subset(ED_EB003E1_all_results,ED_EB003E1_all_results$clone!="empty_NA_empty")
ED_EB003E1_all_results$clone_short=paste0(ED_EB003E1_all_results$shRNA_ID_unified," ",ED_EB003E1_all_results$cl_ID)

cl=levels(c(clone_order))
filtered_ED_EB003E1=ED_EB003E1_all_results[ED_EB003E1_all_results$clone %in% cl, ]
```
###mosaic plot

```{r}
library(see)
###selected and ordered only the one seen in the 2D exp
ggplot(filtered_ED_EB003E1, aes(x = factor(clone,levels = clone_order), y = log2(FC), fill=shRNA_gene,colour=side))+ geom_bar(stat = "identity")+scale_color_manual(values=c("black", "#490092", "#db6d00"))+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in TET vs CTR",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion vs TET")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_TET.pdf"),
     width = 12, height = 8)


```


####volcano plot

```{r}
ggplot(data=ED_EB003E1_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_all_results,ED_EB003E1_all_results$qval<=0.05&log2(ED_EB003E1_all_results$FC)>fc_level|ED_EB003E1_all_results$qval<=0.05&log2(ED_EB003E1_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_EB003E1_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)

ggplot(data=ED_EB003E1_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = side, size=1.5)) +
  scale_colour_manual(values = okabe_pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_all_results,ED_EB003E1_all_results$qval<=0.05&log2(ED_EB003E1_all_results$FC)>fc_level|ED_EB003E1_all_results$qval<=0.05&log2(ED_EB003E1_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "side"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_EB003E1_clone_enrichment_depletion_side.pdf"),
       width = 8, height = 5)
```

####### fisher test enrichment depletion EB003E1 only in SIDE 2 only
###ER between Neural organoids with and without TET

```{r}
cell_metadata_s2=subset(cell_metadata,cell_metadata$side=="Side2")

#####total Fisher test
m_cl=cell_metadata_s2%>%dplyr::select(clone,KD)
clone_list=unique(m_cl$clone)#52


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$KD, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=cell_metadata_s2 %>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = KD, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=cell_metadata_s2%>%
  dplyr::group_by(KD)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_CTR=totals[1,2]
counts$tot_KD=totals[2,2]
counts$TOT=counts$tot_CTR+counts$tot_KD
counts$TOT_clone=counts$CTR+counts$KD

counts$tot_other_CTR=counts$tot_CTR-counts$CTR
counts$tot_other_KD=counts$tot_KD-counts$KD
counts=counts%>%mutate(FC=(KD/tot_KD)/(CTR/tot_CTR),
                       percent_CTR=CTR/tot_CTR*100,
                       percent_KD=KD/tot_KD*100,
                       percent_cl=TOT_clone/TOT*100)
side=cell_metadata_s2 %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_EB003E1_s2.csv"), row.names=FALSE)
ED_EB003E1_s2_results=result_ED_TOTAL
ED_EB003E1_s2_results=subset(ED_EB003E1_s2_results,ED_EB003E1_s2_results$clone!="empty_NA_empty")
ED_EB003E1_s2_results$clone_short=paste0(ED_EB003E1_s2_results$shRNA_ID_unified," ",ED_EB003E1_s2_results$cl_ID)

cl=levels(c(clone_order))
filtered_ED_EB003E1_s2=ED_EB003E1_s2_results[ED_EB003E1_s2_results$clone %in% cl, ]

```

###mosaic plot

```{r}
###selected and ordered only the one seen in the 2D exp
ggplot(filtered_ED_EB003E1_s2, aes(x = factor(clone,levels = clone_order), y = log2(FC), fill=shRNA_gene))+ geom_bar(stat = "identity")+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in TET vs CTR",
    color = "shRNA",
    title= "Enrichment Depletion vs TET in Side 2")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_TET_s2.pdf"),
     width = 12, height = 8)


```

####volcano plot

```{r}
ggplot(data=ED_EB003E1_s2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_s2_results,ED_EB003E1_s2_results$qval<=0.05&log2(ED_EB003E1_s2_results$FC)>fc_level|ED_EB003E1_s2_results$qval<=0.05&log2(ED_EB003E1_s2_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR in side 2")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_EB003E1_clone_enrichment_depletion_SIDE2.pdf"),
       width = 8, height = 5)


```

####### fisher test enrichment depletion EB03E1 and hPSC all side
###ER between organoids with/without TET and hPSC (controls)

```{r}

libID_EB003E1_E2$set_ED=ifelse(libID_EB003E1_E2$diff=="hPSC"&libID_EB003E1_E2$KD=="CTR","hiPSC",
                                          ifelse(libID_EB003E1_E2$diff=="neurons"&libID_EB003E1_E2$KD=="CTR","neuron_CT",
                                        ifelse(libID_EB003E1_E2$diff=="neurons"&libID_EB003E1_E2$KD=="KD","neuron","other")))
meta_CT=subset(libID_EB003E1_E2,libID_EB003E1_E2$KD=="CTR")
#####total Fisher test
m_cl=meta_CT%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#50


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_CT %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_CT%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[1,2]
counts$tot_neuron=totals[2,2]
counts$tot_neuron_CT=totals[3,2]
counts$TOT_neuron=counts$tot_neuron+counts$tot_neuron_CT

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$TOT_neuron
counts$TOT_clone_CT=counts$hiPSC+counts$neuron_CT#tot clone without perturbation

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
#counts$tot_other_neuron=counts$tot_neuron-counts$neuron_CT
counts=counts%>%mutate(FC_CT=(neuron_CT/tot_neuron_CT)/(hiPSC/tot_hiPSC),
                       FC_neuron=(neuron/tot_neuron)/(neuron_CT/tot_neuron_CT),
                       FC=(neuron/tot_neuron)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_neuron_CT=neuron_CT/tot_neuron_CT*100,
                       percent_neuron=neuron/tot_neuron*100,
                       percent_cl=TOT_clone_CT/TOT*100)
side=meta_CT %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_CT_all.csv"), row.names=FALSE)
ED_EB003E1_E2_results=result_ED_TOTAL
ED_EB003E1_E2_results=subset(ED_EB003E1_E2_results,ED_EB003E1_E2_results$clone!="empty_NA_empty")
ED_EB003E1_E2_results$clone_short=paste0(ED_EB003E1_E2_results$shRNA_ID_unified," ",ED_EB003E1_E2_results$cl_ID)


cl=levels(c(clone_order))
filtered_ED_EB003E1_E2=ED_EB003E1_E2_results[ED_EB003E1_E2_results$clone %in% cl, ]


```

###mosaic plot

```{r}
###selected and ordered only the one seen in the 2D exp
ggplot(filtered_ED_EB003E1_E2, aes(x = factor(clone,levels = clone_order), y = log2(FC_CT), fill=shRNA_gene,colour=side))+ geom_bar(stat = "identity")+scale_color_manual(values=c("black", "#490092", "#db6d00"))+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in neuron CTR vs hPSC",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion neuron CTR vs hPSC")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_CONTROLS_vs_hPSC.pdf"),
     width = 12, height = 8)


ggplot(filtered_ED_EB003E1_E2, aes(x = factor(clone,levels = clone_order), y = log2(FC_neuron), fill=shRNA_gene,colour=side))+ geom_bar(stat = "identity")+scale_color_manual(values=c("black", "#490092", "#db6d00"))+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in TET vs CTR neuron",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion vs TET in neuron")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_neuron_TETvsCT.pdf"),
     width = 12, height = 8)

ggplot(filtered_ED_EB003E1_E2, aes(x = factor(clone,levels = clone_order), y = log2(FC), fill=shRNA_gene,colour=side))+ geom_bar(stat = "identity")+scale_color_manual(values=c("black", "#490092", "#db6d00"))+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in neuron TET vs hPSC",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion in neuron TET vs hPSC")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_hPSC.pdf"),
     width = 12, height = 8)

```

####volcano plot

```{r}
ggplot(data=ED_EB003E1_E2_results, aes(x=log2(FC_CT), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_E2_results,ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC_CT)>fc_level|ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC_CT)<fc_level2),
    aes(label = clone_short),
    max.overlaps=50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Neurons CT compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Neurons CTR vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_clone_enrichment_depletion_hPSC.pdf"),
       width = 8, height = 5)

ggplot(data=ED_EB003E1_E2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_E2_results,ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC)>fc_level|ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC)<fc_level2),
    aes(label = clone_short),
    max.overlaps=50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Neurons TET compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Neurons TET vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_clone_enrichment_depletion_neuronsTET_hPSC.pdf"),
       width = 8, height = 5)

ggplot(data=ED_EB003E1_E2_results, aes(x=log2(FC_CT), y=(log2(percent_cl)))) +
  geom_point(aes(colour = side, size=1.5)) +
  scale_colour_manual(values = okabe_pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_E2_results,ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC_CT)>fc_level|ED_EB003E1_E2_results$qval<=0.05&log2(ED_EB003E1_E2_results$FC_CT)<fc_level2),
    aes(label = clone_short),
    max.overlaps=50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Neurons CTR compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Neurons CTR vs hiPSC",
    y = "Log2 of clone fraction",
    color = "side"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_clone_enrichment_depletion_side_hPSC.pdf"),
       width = 8, height = 5)
```

####### fisher test enrichment depletion EB03E1 and hPSC in side 2
###ER between organoids with/without TET and hPSC (controls)

```{r}

meta_CT_s2=subset(meta_CT,meta_CT$side=="Side2")

#####total Fisher test
m_cl=meta_CT_s2%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#22


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_CT %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_CT%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[1,2]
counts$tot_neuron=totals[2,2]
counts$tot_neuron_CT=totals[3,2]
counts$TOT_neuron=counts$tot_neuron+counts$tot_neuron_CT

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$TOT_neuron
counts$TOT_clone_CT=counts$hiPSC+counts$neuron_CT#tot clone without perturbation

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
#counts$tot_other_neuron=counts$tot_neuron-counts$neuron_CT
counts=counts%>%mutate(FC_CT=(neuron_CT/tot_neuron_CT)/(hiPSC/tot_hiPSC),
                       FC_neuron=(neuron/tot_neuron)/(neuron_CT/tot_neuron_CT),
                       FC=(neuron/tot_neuron)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_neuron_CT=neuron_CT/tot_neuron_CT*100,
                       percent_neuron=neuron/tot_neuron*100,
                       percent_cl=TOT_clone_CT/TOT*100)
side=meta_CT %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_CT_side2.csv"), row.names=FALSE)
ED_EB003E1_E2_s2_results=result_ED_TOTAL
ED_EB003E1_E2_s2_results=subset(ED_EB003E1_E2_s2_results,ED_EB003E1_E2_s2_results$clone!="empty_NA_empty")

ED_EB003E1_E2_s2_results$clone_short=paste0(ED_EB003E1_E2_s2_results$shRNA_ID_unified," ",ED_EB003E1_E2_s2_results$cl_ID)


cl=levels(c(clone_order))
filtered_ED_EB003E1_E2_s2=ED_EB003E1_E2_s2_results[ED_EB003E1_E2_s2_results$clone %in% cl, ]


```

###mosaic plot

```{r}
###selected and ordered only the one seen in the 2D exp
ggplot(filtered_ED_EB003E1_E2_s2, aes(x = factor(clone,levels = clone_order), y = log2(FC_CT), fill=shRNA_gene))+ geom_bar(stat = "identity")+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in neuron CTR vs hPSC",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion neuron CTR vs hPSC in side 2")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_CONTROLS_vs_hPSC_s2.pdf"),
     width = 12, height = 8)


ggplot(filtered_ED_EB003E1_E2_s2, aes(x = factor(clone,levels = clone_order), y = log2(FC_neuron), fill=shRNA_gene))+ geom_bar(stat = "identity")+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in TET vs CTR neuron",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion vs TET in neuron in side 2")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_neuron_TETvsCT_s2.pdf"),
     width = 12, height = 8)

ggplot(filtered_ED_EB003E1_E2_s2, aes(x = factor(clone,levels = clone_order), y = log2(FC), fill=shRNA_gene))+ geom_bar(stat = "identity")+
  scale_fill_okabeito()+
  geom_hline(yintercept = fc_level, linetype="dotted")+
  geom_hline(yintercept = fc_level2, linetype="dotted")+
  labs(
    x = "",
    y = "Log2 of clone fraction in neuron TET vs hPSC",
    color = "side",
    color = "shRNA",
    title= "Enrichment Depletion in neuron TET vs hPSC in side 2")
ggsave(filename = paste0(fname_prefix_mosaic, "_", "enrichment_depletion_vs_hPSC_s2.pdf"),
     width = 12, height = 8)

```

####volcano plot

```{r}
ggplot(data=ED_EB003E1_E2_s2_results, aes(x=log2(FC_CT), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_E2_s2_results,ED_EB003E1_E2_s2_results$qval<=0.05&log2(ED_EB003E1_E2_s2_results$FC_CT)>fc_level|ED_EB003E1_E2_s2_results$qval<=0.05&log2(ED_EB003E1_E2_s2_results$FC_CT)<fc_level2),
    aes(label = clone_short),
    max.overlaps=50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion in Side 2 Neurons CT compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Neurons CTR vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_clone_enrichment_depletion_hPSC_s2.pdf"),
       width = 8, height = 5)

ggplot(data=ED_EB003E1_E2_s2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_EB003E1_E2_s2_results,ED_EB003E1_E2_s2_results$qval<=0.05&log2(ED_EB003E1_E2_s2_results$FC)>fc_level|ED_EB003E1_E2_s2_results$qval<=0.05&log2(ED_EB003E1_E2_s2_results$FC)<fc_level2),
    aes(label = clone_short),
    max.overlaps=50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Side 2 Neurons TET compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Neurons TET vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_clone_enrichment_depletion_neuronsTET_hPSC_s2.pdf"),
       width = 8, height = 5)


```


