---
title: "n1 2D analysis exp5 exp7 combined: QC and monocle analysis"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(rCASC)
library(deepToolsUtils)
library(R.utils)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
dir.create(file.path("Output","scratch"))
dir.create(file.path("Output","QC"))
dir.create(file.path("Output","QC","SCR_noTET"))
dir.create(file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "ribomito"))

dir.create(file.path("Output","monocle"))#create forders MONOCLE
dir.create(file.path("Output","monocle","SCR_noTET"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "heatmap"))

dir.create(file.path("Output","Seurat"))#create forders Seurat
dir.create(file.path("Output","Seurat","SCR_noTET"))
dir.create(file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "PCA"))
dir.create(file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "UMAP"))


fname_scratch <- file.path("Output","scratch")
#set up folder path QC
fname_prefix_R_QC <- file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_csv_QC<-file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_ribomito_QC <- file.path("Output","QC","SCR_noTET", format(Sys.Date(), "%y%m%d"), "ribomito", 
                               format(Sys.Date(), "%y%m%d"))

#set up forder path MONOCLE
fname_prefix_R <- file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","monocle","SCR_noTET", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))


#set up forder path Seurat
fname_prefix_R_Seurat <- file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_Seurat <- file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_PCA_Seurat<-file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "PCA", 
                               format(Sys.Date(), "%y%m%d"))

fname_prefix_UMAP_Seurat<-file.path("Output","Seurat","SCR_noTET", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))

okabe_tab=read_excel("Output/scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")

pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```

########load data

```{r}
load("Output/monocle/240407/R_objects/240407_updated_cds_with_modules_SELECTION_side.RData")
load("Output/monocle/240407/R_objects/240407_gene_modules_pseudotime_cardio.RData")
load("Output/monocle/240411/R_objects/240411_cds_updated.RData")
ID_key=read_excel("shRNA_IDs.xlsx")
cds@colData[, paste0("module_", 1:74)] <- modules[, 1:74]
meta[, paste0("module_", 1:74)] <- modules[, 1:74]

####prepare table with cluster assignment
ct=c("Cardiac fibroblasts","Late cardiomyocytes","Early cardiomyocytes","Cardiac progenitors","Lateral plate mesoderm","Proliferating cardiomyocytes","Endothelial cells","Mesendoderm","Endoderm derivatives")
ct_sh=c("CF","L-CM","E-CM","CP","LPM","P-CM","EC","MES","ENDO")

clust_df=data.frame(monocle_cluster=c(1:9),cell_types=ct,cell_types_abbreviations=ct_sh)

###prepare color/time order vectors
cell_order=c("Mesendoderm","Cardiac progenitors","Early cardiomyocytes","Late cardiomyocytes","Proliferating cardiomyocytes","Lateral plate mesoderm","Cardiac fibroblasts","Endothelial cells","Endoderm derivatives")
cell_order_short=c("MES","CP","E-CM","L-CM","P-CM","LPM","CF","EC","ENDO")
gene_order=c("NA","SCR","B2M","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")
gene_polish_order=c("control","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")

```

###get pseudotime data
```{r}

Ps_all=as.data.frame(pseudotime(cds, reduction_method = "UMAP"))
colnames(Ps_all)="pseudotime"
cds@colData$pseudotime=Ps_all$pseudotime

Ps_cardio=as.data.frame(pseudotime(cds_cardio, reduction_method = "UMAP"))
colnames(Ps_cardio)="pseudotime"
cds_cardio@colData$pseudotime=Ps_cardio$pseudotime

colData(cds_cardio)$gene_polish=ifelse(colData(cds_cardio)$shRNA_gene == "CHD7", "CHD7", 
                                ifelse(colData(cds_cardio)$shRNA_gene == "GATA4", "GATA4",
                                       ifelse(colData(cds_cardio)$shRNA_gene == "KMT2D", "KMT2D",
                                              ifelse(colData(cds_cardio)$shRNA_gene == "NKX2.5", "NKX2.5",
                                                     ifelse(colData(cds_cardio)$shRNA_gene == "SMAD2", "SMAD2","control")))))

meta=as.data.frame(cds@colData@listData)
meta_cardio=as.data.frame(cds_cardio@colData@listData)
sel=meta%>%select(has_70_counts,clone_20_counts_clone,shRNA_40_counts,libID)

meta_cardio=meta_cardio%>%left_join(sel, by="libID")

```

####select top 10 modules

```{r}
########plot all modules 
plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(1:74)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules.pdf"),
     width = 21, height = 30)

####eliminate the modules with expression in only few cells abd general low expression
select_modules=as.data.frame(agg_mat_test)
select_modules=select_modules%>%dplyr::filter(rownames(select_modules)!="Module 73"&rownames(select_modules)!="Module 74"&rownames(select_modules)!="Module 58")#####select modules with decent amount of cells expressing genes

hp1=pheatmap::pheatmap(select_modules,
                   scale="row", clustering_method="ward.D2",main="row normalized modules filtered",color = viridis(n = 256, alpha =1,begin = 0, end = 1, option = "H"))#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_modules_filtered.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp1
dev.off()

####get FC for each couple
select_modules=select_modules%>%dplyr::rename(empty_CTR=NA_CTR,empty_TET=NA_TET)
genes=c("empty","SCR","B2M","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")

B2M=select_modules%>%dplyr::select(B2M_CTR,B2M_TET)%>%mutate(B2M_delta=B2M_TET-B2M_CTR)%>%dplyr::select(-B2M_CTR,-B2M_TET)
SCR=select_modules%>%dplyr::select(SCR_CTR,SCR_TET)%>%mutate(SCR_delta=SCR_TET-SCR_CTR)%>%dplyr::select(-SCR_CTR,-SCR_TET)
empty=select_modules%>%dplyr::select(empty_CTR,empty_TET)%>%mutate(empty_delta=empty_TET-empty_CTR)%>%dplyr::select(-empty_CTR,-empty_TET)
CHD7=select_modules%>%dplyr::select(CHD7_CTR,CHD7_TET)%>%mutate(CHD7_delta=CHD7_TET-CHD7_CTR)%>%dplyr::select(-CHD7_CTR,-CHD7_TET)
GATA4=select_modules%>%dplyr::select(GATA4_CTR,GATA4_TET)%>%mutate(GATA4_delta=GATA4_TET-GATA4_CTR)%>%dplyr::select(-GATA4_CTR,-GATA4_TET)
KMT2D=select_modules%>%dplyr::select(KMT2D_CTR,KMT2D_TET)%>%mutate(KMT2D_delta=KMT2D_TET-KMT2D_CTR)%>%dplyr::select(-KMT2D_CTR,-KMT2D_TET)
NKX2.5=select_modules%>%dplyr::select(NKX2.5_CTR,NKX2.5_TET)%>%mutate(NKX2.5_delta=NKX2.5_TET-NKX2.5_CTR)%>%dplyr::select(-NKX2.5_CTR,-NKX2.5_TET)
SMAD2=select_modules%>%dplyr::select(SMAD2_CTR,SMAD2_TET)%>%mutate(SMAD2_delta=SMAD2_TET-SMAD2_CTR)%>%dplyr::select(-SMAD2_CTR,-SMAD2_TET)

delta=cbind(B2M,SCR,empty,CHD7,GATA4,KMT2D,NKX2.5,SMAD2)

hp2=pheatmap::pheatmap(delta, clustering_method="ward.D2",main="delta modules filtered",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))

hp3=pheatmap::pheatmap(delta,
                   scale="row", clustering_method="ward.D2",main="Delta row normalized modules filtered",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))#scaled on rows (modules) >>>to compare modules between perturbations
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_delta_modules_filtered.pdf")
pdf(Rdata_filename, height = 10, width = 8)
hp3
dev.off()

cal_z_score <- function(x){
  (x - mean(x)) / sd(x)
}

data_subset_norm <- t(apply(delta, 1, cal_z_score))

abs_delta_norm=abs(as.data.frame(data_subset_norm))
abs_delta_norm$max=apply(abs_delta_norm,1,FUN=max)
top=abs_delta_norm[order(-abs_delta_norm$max),] ####select top 10 based on delta
t10_norm=head(top, n=10)
t10_norm=rownames(t10_norm)
t15_norm=head(top, n=15)
t15_norm=rownames(t15_norm)
t20_norm=head(top, n=20)
t20_norm=rownames(t20_norm)

top10_norm=delta%>%filter(rownames(delta)%in%t10_norm)
top15_norm=delta%>%filter(rownames(delta)%in%t15_norm)
top20_norm=delta%>%filter(rownames(delta)%in%t20_norm)


abs_delta=abs(as.data.frame(delta))
abs_delta$max=apply(abs_delta,1,FUN=max)
top=abs_delta[order(-abs_delta$max),] ####select top 10 based on delta
t10=head(top, n=10)
t10=rownames(t10)
t15=head(top, n=15)
t15=rownames(t15)
t20=head(top, n=20)
t20=rownames(t20)
top10=delta%>%filter(rownames(delta)%in%t10)
top15=delta%>%filter(rownames(delta)%in%t15)
top20=delta%>%filter(rownames(delta)%in%t20)

pheatmap::pheatmap(top10_norm,
                   scale="row", clustering_method="ward.D2",main="top 10 modules from norm",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))

pheatmap::pheatmap(top10, scale="row",clustering_method="ward.D2",main="top 10 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))

hm_t15=pheatmap::pheatmap(top15_norm, scale="row",clustering_method="ward.D2",main="top 15 modules from norm",cutree_rows = 4,color = viridis(n = 256, alpha = 1,begin = 0, end = 1, option = "H"))
hm_t15
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_top15_modules.pdf")
pdf(Rdata_filename, height = 5, width = 8)
hm_t15
dev.off()

pheatmap::pheatmap(top15, scale="row",clustering_method="ward.D2",main="top 15 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))

pheatmap::pheatmap(top20_norm, scale="row",clustering_method="ward.D2",main="top 20 modules from norm",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))

pheatmap::pheatmap(top20, scale="row",clustering_method="ward.D2",main="top 20 modules",color = viridis(n = 256, alpha = 1, 
                                   begin = 0, end = 1, option = "H"))


```


###ks statistics for genes along pseudotime trajectory and modules expression

```{r}

analyze=data.frame(t15_norm=t15_norm)
analyze$module_nb=sapply(strsplit(as.character(t15_norm), " "), "[[", 2)
analyze$module=paste0("module_",analyze$module_nb)
analyze$pal=c("#AA0DFE","#3283FE","#85660D","#782AB6","#565656","#1C8356","#16FF32","#F7E1A0","#E2E2E2","#1CBE4F","#C4451C","#DEA0FD","#FE00FA","#325A9B","#FEAF16")
rownames(analyze)=analyze$t15_norm

row=data.frame(module=rownames(top15_norm))
rownames(row)=row$module

my_colour = list(module = c("Module 41"="#AA0DFE","Module 13"="#3283FE","Module 30"="#85660D","Module 2"="#782AB6","Module 64"="#565656","Module 1"="#1C8356","Module 14"="#16FF32","Module 16"="#F7E1A0","Module 72"="#E2E2E2","Module 62"="#1CBE4F","Module 22"="#C4451C","Module 26"="#DEA0FD","Module 59"="#FE00FA","Module 33"="#325A9B","Module 20"="#FEAF16"))

hm_t15=pheatmap::pheatmap(top15_norm, scale="row",clustering_method="ward.D2",annotation_colors = my_colour,annotation_row =row,main="top 15 modules from norm",cutree_rows = 4,color = viridis(n = 256, alpha = 1,begin = 0, end = 1, option = "H"))
hm_t15
Rdata_filename = paste0(fname_prefix_heatmap, "_hp_top15_modules.pdf")
pdf(Rdata_filename, height = 5, width = 8)
hm_t15
dev.off()

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(analyze$module_nb)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules_top_15.pdf"),
     width = 10, height = 8)

analysis=c(analyze$module)


TET = subset(meta, meta$shRNA_gene != "SCR"&meta$knockdown == "TET")
CTR = subset(meta, meta$shRNA_gene == "SCR"&meta$knockdown == "TET")

genes=c("NKX2.5","B2M","KMT2D","CHD7","GATA4","SMAD2","NA")

ks_stat = data.frame()

for (c in genes) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_gene == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == "SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    
    ks_stat = rbind(ks_stat, data.frame(gene = c, pval = ks_pval, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("gene", "pval", "analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05

ks_stat_filtered_modules=ks_stat

write.csv(ks_stat_filtered_modules, file= paste0(fname_prefix_csv, "_", "ks_statistics_filtered_modules_SCR.csv"), row.names=FALSE)


#######see modules that stands out
analysis=c(analyze$module)

TET = subset(meta, meta$shRNA_gene != "SCR"&meta$knockdown == "TET")
CTR = subset(meta, meta$shRNA_gene == "SCR"&meta$knockdown == "TET")

genes=c("NKX2.5","B2M","KMT2D","CHD7","GATA4","SMAD2","NA")

ks_stat = data.frame()

for (c in genes) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_gene == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == "SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    ks_pval_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["p.value"]]
    ks_sig_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["statistic"]][["D^+"]]
    ks_pval_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["p.value"]]
    ks_sig_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["statistic"]][["D^-"]]
    ks_sig = ks.test(tmp_TET, tmp_CTR)[["statistic"]][["D"]]
    
    ks_stat = rbind(ks_stat, data.frame(gene = c, pval = ks_pval, pval_g = ks_pval_g,pval_l = ks_pval_l,ks_sign_g=ks_sig_g,ks_sign_l=ks_sig_l,ks_sign=ks_sig, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("gene", "pval", "pval_greater","pval_less","ks_greater","ks_less","ks_sign","analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05
ks_stat$greater <- ks_stat$pval_greater <= 0.05
ks_stat$less <- ks_stat$pval_less <= 0.05

ks_stat$k_sign_summary <- ifelse(
  # "greater" is TRUE and "less" is FALSE
  ks_stat$greater & !ks_stat$less,
  as.numeric(ks_stat$ks_greater),  # Return ks_sign value
  ifelse(
    # "greater" is FALSE and "less" is TRUE
    !ks_stat$greater & ks_stat$less,
    -as.numeric(ks_stat$ks_less),  # Return negative of ks_sign
    ks_stat$ks_sign  # Otherwise, return ks sign
  )
)

ks_stat$k_colour <- ifelse(
  ks_stat$sig==T,ks_stat$analysis,"")
 
k_stat_all=ks_stat
write.csv(ks_stat, file= paste0(fname_prefix_csv, "_", "ks_statistics_sign_top15_SCR.csv"), row.names=FALSE)

k_stat_all=k_stat_all%>%dplyr::rename(module=analysis)
k_stat_all=k_stat_all%>%left_join(analyze, by="module")
```

#####volcano plot statistics
###genes

```{r}
library(pals)

a=k_stat_all%>%filter(k_stat_all$sig==T)%>%dplyr::select(module,pal)%>%distinct(module,pal)
a=a[order(a$module),]

ggplot(data=k_stat_all, aes(x=k_sign_summary, y=log10(padj)*-1)) +
  geom_point(aes(colour = k_colour), size=1.5) +scale_color_manual(values=c("grey",a$pal))+
  ggrepel::geom_text_repel(
    data = subset(k_stat_all, sig==T),
    aes(label = gene),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  ggtitle("ks test statistics per gene vs its control")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "KS statistic summary",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_ks_stats_gene_SCR.pdf"),
       width = 6, height = 4)

```
#######get statistics for shRNA

```{r}

meta_sh=subset(meta,meta$shRNA_40_counts==T)
analysis=c(analyze$module)

TET = subset(meta_sh, meta_sh$shRNA_gene != "SCR"&meta_sh$knockdown == "TET")
CTR = subset(meta_sh, meta_sh$shRNA_gene == "SCR"&meta_sh$knockdown == "TET")

ks_stat = data.frame()

a=unique(meta_sh$shRNA_label)
a=a [! a %in% "SCR"]

for (c in a) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_label == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_label =="SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    
    ks_stat = rbind(ks_stat, data.frame(shRNA_label = c, pval = ks_pval, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("shRNA_label", "pval", "analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05

ks_stat_filtered_modules_shRNA=ks_stat

write.csv(ks_stat_filtered_modules_shRNA, file= paste0(fname_prefix_csv, "_", "ks_statistics_shRNA_filtered_modules.csv"), row.names=FALSE)


#######see modules that stands out
analysis=c(analyze$module)

TET = subset(meta_sh, meta_sh$shRNA_gene != "SCR"&meta_sh$knockdown == "TET")
CTR = subset(meta_sh, meta_sh$shRNA_gene == "SCR"&meta_sh$knockdown == "TET")

ks_stat = data.frame()

a=unique(meta_sh$shRNA_label)
a=a [! a %in% "SCR"]

ks_stat = data.frame()

for (c in a) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$shRNA_label == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_label == "SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    ks_pval_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["p.value"]]
    ks_sig_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["statistic"]][["D^+"]]
    ks_pval_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["p.value"]]
    ks_sig_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["statistic"]][["D^-"]]
    ks_sig = ks.test(tmp_TET, tmp_CTR)[["statistic"]][["D"]]
    
    ks_stat = rbind(ks_stat, data.frame(shRNA_label = c, pval = ks_pval, pval_g = ks_pval_g,pval_l = ks_pval_l,ks_sign_g=ks_sig_g,ks_sign_l=ks_sig_l,ks_sign=ks_sig, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("shRNA_label", "pval", "pval_greater","pval_less","ks_greater","ks_less","ks_sign","analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05
ks_stat$greater <- ks_stat$pval_greater <= 0.05
ks_stat$less <- ks_stat$pval_less <= 0.05

ks_stat$k_sign_summary <- ifelse(
  # "greater" is TRUE and "less" is FALSE
  ks_stat$greater & !ks_stat$less,
  as.numeric(ks_stat$ks_greater),  # Return ks_sign value
  ifelse(
    # "greater" is FALSE and "less" is TRUE
    !ks_stat$greater & ks_stat$less,
    -as.numeric(ks_stat$ks_less),  # Return negative of ks_sign
    ks_stat$ks_sign  # Otherwise, return ks sign
  )
)

ks_stat$k_colour <- ifelse(
  ks_stat$sig==T,ks_stat$analysis,"")
 
k_stat_shRNA=ks_stat
write.csv(ks_stat, file= paste0(fname_prefix_csv, "_", "ks_statistics_shRNA_sign_top15.csv"), row.names=FALSE)

k_stat_shRNA=k_stat_shRNA%>%dplyr::rename(module=analysis)
k_stat_shRNA=k_stat_shRNA%>%left_join(analyze, by="module")
```

#####volcano plot statistics
###shRNA

```{r}
a=k_stat_shRNA%>%filter(k_stat_shRNA$sig==T)%>%dplyr::select(module,pal)%>%distinct(module,pal)
a=a[order(a$module),]

ggplot(data=k_stat_shRNA, aes(x=k_sign_summary, y=log10(padj)*-1)) +
  geom_point(aes(colour = k_colour), size=1.5)+
  scale_color_manual(values=c("grey",a$pal))+
  ggrepel::geom_text_repel(
    data = subset(k_stat_shRNA, sig==T),
    aes(label =shRNA_label),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  ggtitle("ks test statistics per shRNA vs its control")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "KS statistic summary",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_ks_stats_shRNA.pdf"),
       width = 6, height = 4)

```

#######get statistics for clones

```{r}

meta_cl=subset(meta,meta$clone_20_counts==T)
analysis=c(analyze$module)

TET = subset(meta_cl, meta_cl$shRNA_gene != "SCR"&meta_sh$knockdown == "TET")
CTR = subset(meta_cl, meta_cl$shRNA_gene == "SCR"&meta_sh$knockdown == "TET")

ks_stat = data.frame()

a=unique(meta_cl$clone)
a=a [! a %in% "GCCTGTGT_SCR_ACGGTC"]

ks_stat = data.frame()

for (c in a) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$clone == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == "SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    
    ks_stat = rbind(ks_stat, data.frame(clone = c, pval = ks_pval, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("clone", "pval", "analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05

ks_stat_filtered_modules_clone=ks_stat

write.csv(ks_stat_filtered_modules_clone, file= paste0(fname_prefix_csv, "_", "ks_statistics_clone_filtered_modules.csv"), row.names=FALSE)


#######see modules that stands out
analysis=c(analyze$module)

TET = subset(meta_cl, meta_cl$shRNA_gene != "SCR"&meta_sh$knockdown == "TET")
CTR = subset(meta_cl, meta_cl$shRNA_gene == "SCR"&meta_sh$knockdown == "TET")

ks_stat = data.frame()

a=unique(meta_cl$clone)
a=a [! a %in% "GCCTGTGT_SCR_ACGGTC"]

ks_stat = data.frame()

for (c in a) {
  for (analysis_type in analysis) {
    tmp_TET = TET[TET$clone == c, analysis_type]
    tmp_CTR = CTR[CTR$shRNA_gene == "SCR", analysis_type]
    
    ks_pval = ks.test(tmp_TET, tmp_CTR)[["p.value"]]
    ks_pval_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["p.value"]]
    ks_sig_g = ks.test(tmp_TET, tmp_CTR,alternative="greater")[["statistic"]][["D^+"]]
    ks_pval_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["p.value"]]
    ks_sig_l = ks.test(tmp_TET, tmp_CTR,alternative="less")[["statistic"]][["D^-"]]
    ks_sig = ks.test(tmp_TET, tmp_CTR)[["statistic"]][["D"]]
    
    ks_stat = rbind(ks_stat, data.frame(clone = c, pval = ks_pval, pval_g = ks_pval_g,pval_l = ks_pval_l,ks_sign_g=ks_sig_g,ks_sign_l=ks_sig_l,ks_sign=ks_sig, analysis = analysis_type))
  }
}

colnames(ks_stat) <- c("clone", "pval", "pval_greater","pval_less","ks_greater","ks_less","ks_sign","analysis")

ks_stat$padj <- p.adjust(ks_stat$pval, method = "BH")
ks_stat$sig <- ks_stat$padj <= 0.05
ks_stat$greater <- ks_stat$pval_greater <= 0.05
ks_stat$less <- ks_stat$pval_less <= 0.05

ks_stat$k_sign_summary <- ifelse(
  # "greater" is TRUE and "less" is FALSE
  ks_stat$greater & !ks_stat$less,
  as.numeric(ks_stat$ks_greater),  # Return ks_sign value
  ifelse(
    # "greater" is FALSE and "less" is TRUE
    !ks_stat$greater & ks_stat$less,
    -as.numeric(ks_stat$ks_less),  # Return negative of ks_sign
    ks_stat$ks_sign  # Otherwise, return ks sign
  )
)

ks_stat$k_colour <- ifelse(
  ks_stat$sig==T,ks_stat$analysis,"")
 
k_stat_clone=ks_stat
k_stat_clone=k_stat_clone%>%left_join(clones_AB03,by="clone")
write.csv(k_stat_clone, file= paste0(fname_prefix_csv, "_", "ks_statistics_clone_sign_top15.csv"), row.names=FALSE)

k_stat_clone=k_stat_clone%>%dplyr::rename(module=analysis)
k_stat_clone=k_stat_clone%>%left_join(analyze, by="module")

k_stat_clone$sig_FDR001=k_stat_clone$padj <= 0.01
```

#####volcano plot statistics
###clones

```{r}
a=k_stat_clone%>%filter(k_stat_clone$sig==T)%>%dplyr::select(module,pal)%>%distinct(module,pal)
a=a[order(a$module),]

ggplot(data=k_stat_clone, aes(x=k_sign_summary, y=log10(padj)*-1)) +
  geom_point(aes(colour = k_colour), size=1.5)+
  scale_color_manual(values=c("grey",a$pal))+
  ggrepel::geom_text_repel(
    data = subset(k_stat_clone, sig==T),
    aes(label =clone_short),
    max.overlaps = 50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  ggtitle("ks test statistics per clone vs SCR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "KS statistic summary",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_ks_stats_clone_SCR.pdf"),
       width = 6, height = 4)

ggplot(data=k_stat_clone, aes(x=k_sign_summary, y=log10(padj)*-1)) +
  geom_point(aes(colour = k_colour), size=1.5)+
  scale_color_manual(values=c("grey",a$pal))+
  ggrepel::geom_text_repel(
    data = subset(k_stat_clone, sig_FDR001==T),
    aes(label =clone_short),
    max.overlaps = 50,
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  ggtitle("ks test statistics per clone vs SCR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "KS statistic summary",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_ks_stats_clone_sig_FDR001_SCR.pdf"),
       width = 6, height = 4)

SIG=k_stat_shRNA%>%filter(k_stat_shRNA$sig==T)%>%dplyr::select(module_nb)%>%distinct(module_nb)

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(SIG$module_nb)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules_top_15_SIG.pdf"),
     width = 9, height = 6)


SIG=k_stat_shRNA%>%filter(k_stat_shRNA$sig==T)%>%dplyr::select(module_nb)%>%distinct(module_nb)

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(2,41,62,64)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules_MAIN.pdf"),
     width = 7, height = 5)

plot_cells(cds,
           genes=gene_module_df %>% filter(module %in% c(2,41,62,64, 1,13,16,20,22,30,59,72)),
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE,cell_size=1)+
labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle modules")
last_plot() + scale_color_viridis(option="E", discrete=F)

ggsave(filename = paste0(fname_prefix_UMAP, "_", "Monocle_modules_SUPP.pdf"),
     width = 13, height = 6)



```

######get info about gene modules

```{r}

gene_modules=list()

for (c in unique(gene_module_df$module)) {
 
tmp=subset(gene_module_df, gene_module_df$module==c)
gene_list=tmp$id

gene_modules[[c]]=gene_list
}

library(topGO)

library(org.Hs.eg.db)

hs <- org.Hs.eg.db

modules=c("41" ,"13", "30", "2","64" ,"1","14", "16", "72", "62", "22", "26", "59", "33", "20")

go.de_list <- list()
k_list <- list()

# Iterate over each module in 'modules'
for (module in modules) {
  # Convert gene symbols to ENSEMBL or ENTREZID
  gene.df <- clusterProfiler::bitr(gene_modules[[module]], fromType = "SYMBOL",
                                    toType = c("ENSEMBL", "ENTREZID"),
                                    OrgDb = org.Hs.eg.db)
  
  # Perform GO analysis
  go.de <- goana(list(DE1 = gene.df$ENTREZID), species = "Hs", convert = TRUE) %>%
    mutate(module = module)
  
  # Perform KEGG analysis
  k <- kegga(list(DE1 = gene.df$ENTREZID), species = "Hs", convert = TRUE) %>%
    mutate(module = module)
  
  # Store the results in lists
  go.de_list[[module]] <- go.de
  k_list[[module]] <- k
}

# Combine all data frames in the lists into single data frames
GO <- do.call(rbind, go.de_list)%>%filter(Ont!="CC",DE1>=20,P.DE1<=0.05)
k <- do.call(rbind, k_list)%>%filter(DE1>=6,P.DE1<=0.05)

write.csv(GO, file= paste0(fname_prefix_csv, "_", "GO_modules.csv"), row.names=FALSE)
write.csv(k, file= paste0(fname_prefix_csv, "_", "kegg_modules.csv"), row.names=FALSE)

```

#####cluster profiler

```{r}

modules=c("41" ,"13", "30", "2","64" ,"1","14", "16", "72", "62", "26", "59", "33", "20")

eGO_list <- list()

# Iterate over each module in 'modules'
for (module in modules) {
  # Perform enrichment analysis for the current module
  ego <- clusterProfiler::enrichGO(gene = gene_modules[[module]],
                                    OrgDb = org.Hs.eg.db,
                                    keyType = 'SYMBOL',
                                    ont = "ALL",
                                    pAdjustMethod = "BH")
  
  # Store the result in the list
  eGO_result <- ego@result %>% mutate(module = module)
  eGO_list[[module]] <- eGO_result
}

# Combine all data frames in the list into a single data frame
eGO <- do.call(rbind, eGO_list)

write.csv(eGO, file= paste0(fname_prefix_csv, "_", "eGO_modules.csv"), row.names=FALSE)
```


####cumulative frequency plots

```{r}
ggplot(meta, aes(pseudotime, colour = KD)) + stat_ecdf(geom = "step")+ 
  #scale_color_oi(order = c(9, 7))+
  theme_classic(base_size = 14)+
  ylab("Cumulative frequency")+
  xlab("monocle pseudotime")+
  ggtitle("Cumulative frequency cardio on pseudotime")+
  facet_grid(~factor(shRNA_gene,levels=gene_order))

ggsave(filename = paste0(fname_prefix_dotplot, "_", "cumulative_frequency_pseudotime.png"),
     width = 14, height = 5)

for (module_nb in analyze$module) {
  # Create the plot for the current module
  p <- ggplot(meta, aes_string(x = module_nb, colour = "KD")) + 
    stat_ecdf(geom = "step") + 
    theme_classic(base_size = 14) + 
    ylab("Cumulative frequency") + 
    xlab("mmodule aggregate gene expression") + 
    ggtitle(paste0("Cumulative frequency on ", module_nb)) + 
    facet_grid(~factor(shRNA_gene, levels = gene_order))
  
  # Save the plot as an image file
  ggsave(filename = paste0(fname_prefix_dotplot, "_", "cumulative_frequency_", gsub("module_", "", module_nb), ".png"),
         plot = p, width = 14, height = 5)
}

```


####save data

```{r}
meta=as.data.frame(cds@colData@listData)
meta_cardio=as.data.frame(cds_cardio@colData@listData)

#Rdata_filename = paste0(fname_prefix_R, "_calculations_nb_percent_clones_shRNA.RData")
#save(by_gene_KD,by_gene_KD_ratio,by_shRNA_KD,by_clone_KD,clone_count,shRNA_count,
 #    file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_GO.RData")
save(GO,k,eGO,
     file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_cds_updated.RData")
save(meta,meta_cardio,cds,cds_cardio,
     file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_cds_modules.RData")
save(meta,cds,analyze,agg_mat_test,gene_module_df, modules,top15_norm,delta, data_subset_norm,abs_FC,
     file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_gene_modules_pseudotime.RData")
save(cardio_cds_pr_test_res,genes_mod_pseudotime, gene_module_df, modules,agg_mat,agg_mat_test,
     file = Rdata_filename)

```


