---
title: "n1 2D analysis exp5 exp7 combined: QC and monocle analysis"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(deepToolsUtils)
library(R.utils)
library(see)
library(pals)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
dir.create(file.path("Output","scratch"))

dir.create(file.path("Output","SIDE2"))
dir.create(file.path("Output","SIDE2","monocle"))#create forders MONOCLE
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "mosaic"))


fname_scratch <- file.path("Output","scratch")

#set up forder path MONOCLE
fname_prefix_R <- file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","SIDE2","monocle", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))

okabe_tab=read_excel("Output/scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")


pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```

########load data

```{r}
####load modules
load("Output/monocle/240407/R_objects/240407_gene_modules_pseudotime_cardio.RData")
load("Output/monocle/240407/R_objects/240407_gene_modules_pseudotime.RData")
###load cds latest and modules info
load("Output/monocle/240407/R_objects/240407_updated_cds_with_modules_SELECTION_side.RData")
ID_key=read_excel(paste0(fname_scratch,"/shRNA_IDs.xlsx"))
BC=meta%>%select(shRNA_BC)%>%dplyr::distinct(shRNA_BC)
ID=ID_key%>%dplyr::select(shRNA_label,shRNA_copy_label,shRNA_BC,hex,okabe_hex,shRNA_ID_unified)%>%distinct(shRNA_label,shRNA_copy_label,shRNA_BC,hex,okabe_hex,shRNA_ID_unified)%>%filter(shRNA_BC%in%BC$shRNA_BC)

meta=meta%>%select(-shRNA_ID_unified)
meta=meta%>%left_join(ID, by="shRNA_BC")

cds@colData$shRNA_ID_unified=meta$shRNA_ID_unified
cds@colData$shRNA_label=meta$shRNA_label
cds@colData$shRNA_copy_label=meta$shRNA_copy_label
cds@colData$hex=meta$hex
cds@colData$okabe_hex=meta$okabe_hex

clones_AB03=read_excel(paste0(fname_scratch,"/clones_AB03.xlsx"))%>%select(-count)
meta=meta%>%left_join(clones_AB03, by=c("cell_BC","clone","shRNA_gene"))

colData(cds)$cl_ID=meta$cl_ID
colData(cds)$clone_short=meta$clone_short

okabe_tab=read_excel("Output/scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")

###correct empty that was placed in unknown category
meta <- meta %>%
    mutate(side = if_else(clone == "empty_NA_empty" & side == "unknown", "Side2", side))
colData(cds)$side=meta$side###correct empty taht was placed in unknown category

cds=cds[,cds$side=="Side2"]
meta=as.data.frame(cds@colData@listData)###4704
meta <- replace(meta, is.na(meta),"empty")
```


##### check cluster composition

```{r}

cell_order=c("Mesendoderm","Cardiac progenitors","Early cardiomyocytes","Late cardiomyocytes","Proliferating cardiomyocytes","Lateral plate mesoderm","Cardiac fibroblasts","Endothelial cells","Endoderm derivatives")
cell_order_short=c("MES","CP","E-CM","L-CM","P-CM","LPM","CF","EC","ENDO")

ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = factor(assigned_cell_type,levels=cell_order)),
               size=2)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Cluster composition")+
  labs(colour="")+
  scale_color_okabeito(reverse = T)+
  facet_grid(~KD)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_monocle_clusters.pdf"),
     width = 10, height = 6)

gene_order=c("NA","SCR","B2M","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")

######create df for gene colors
color_g=meta%>%select(shRNA_gene,okabe_hex)%>%distinct(shRNA_gene,okabe_hex)%>%mutate(order=shRNA_gene)
color_g$order=color_g$order%>%recode("NA"=1,"SCR"=2,"B2M"=3,"CHD7"=4,"GATA4"=5,"KMT2D"=6,"NKX2.5"=7,"SMAD2"=8)
color_g=color_g[order(color_g$order),]

ggplot(meta, aes(x = UMAP_1_monocle, y = UMAP_2_monocle)) +
    geom_point(aes(color = factor(assigned_cell_type_short,levels=cell_order_short)),
               size=2)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Cluster composition by gene",colour="")+
  scale_color_okabeito(reverse = T)+
  facet_grid(~factor(shRNA_gene,levels = gene_order )~KD)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_monocle_clusters_genes.pdf"),
     width = 6, height = 15)


####prepare table with cluster assignment

ct=c("Cardiac fibroblasts","Late cardiomyocytes","Early cardiomyocytes","Cardiac progenitors","Lateral plate mesoderm","Proliferating cardiomyocytes","Endothelial cells","Mesendoderm","Endoderm derivatives")

ct_sh=c("CF","L-CM","E-CM","CP","LPM","P-CM","EC","MES","ENDO")

clust_df=data.frame(monocle_cluster=c(1:9),cell_types=ct,cell_types_abbreviations=ct_sh)

###prepare color/time order vectors

cell_order=c("Mesendoderm","Cardiac progenitors","Early cardiomyocytes","Late cardiomyocytes","Proliferating cardiomyocytes","Lateral plate mesoderm","Cardiac fibroblasts","Endothelial cells","Endoderm derivatives")
cell_order_short=c("MES","CP","E-CM","L-CM","P-CM","LPM","CF","EC","ENDO")
gene_order=c("NA","SCR","B2M","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")
gene_order_polished=c("control","CHD7","GATA4","KMT2D","NKX2.5","SMAD2")

monocle_long=data.frame(cell_order=c("Cardiac fibroblasts","Cardiac progenitors","Early cardiomyocytes","Endoderm derivatives","Endothelial cells","Late cardiomyocytes","Lateral plate mesoderm","Mesendoderm","Proliferating cardiomyocytes"),order_LONG=c(1:9))

monocle_short=data.frame(cell_order_short=c("CF","CP","E-CM","EC","ENDO","L-CM","LPM","MES","P-CM"),order_SHORT=c(1:9))

color_df=data.frame(cell_order_short=cell_order_short,cell_order=cell_order,cl_col=c("#000000","grey","#CC79A7","#D55E00","#0072B2","#f0E442","#009E73","#56B4E9","#E69F00"))
color_df=color_df%>%left_join(monocle_short,by="cell_order_short")%>%left_join(monocle_long,by="cell_order")

color_df_l=color_df[order(color_df$order_LONG),]
color_df_s=color_df[order(color_df$order_SHORT),]

plot_cells(cds, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters.pdf"),
     width = 8, height = 5)

#####add last info to cds file and divide into CDS files per side

colData(cds)$gene_polish=colData(cds)$shRNA_gene
colData(cds)$gene_polish=dplyr::recode(colData(cds)$gene_polish,"B2M"="control","SCR"="control","NA"="control")
colData(cds)$gene_KD=paste0(cds$gene_polish,"_",cds$KD)
colData(cds)$g_KD=paste0(cds$shRNA_gene,"_",cds$KD)

meta=as.data.frame(cds@colData@listData)###5858
meta <- replace(meta, is.na(meta),"empty")


###########
clone_counts=meta %>%
  dplyr::group_by(clone,side) %>%
  dplyr::summarise(count = n())###tot clones 305
clone_counts=spread(clone_counts, key = side, value = count)
clone_counts_side <- clone_counts %>% replace(is.na(.), 0)

plot_cells(cds, color_cells_by="side", group_cells_by="assigned_cell_type",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_s$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering side")+
  facet_wrap(~batch~side, ncol=4)
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_batch.pdf"),
     width = 10, height = 6)


```

###calculate % cells of one gene in each cluster compared to its control

```{r}

###make files with % calculations

##################
############ calculate % per cluster of each gene CTR and KD 
################
meta_to_use=meta

tot=table(meta_to_use$g_KD) #total of the cluster
df=table(meta_to_use$g_KD,meta_to_use$monocle_clusters,meta_to_use$shRNA_gene,meta_to_use$KD)
df = as.data.frame(df)
colnames(df)=c("gene_KD","monocle_cluster","gene","KD","freq")

#mdf2 = melt(df)
mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("gene_KD","tot")
mdf2=mdf2%>%left_join(total, by="gene_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)
#m <- m %>%
 # mutate(samp = ifelse(grepl("n3_", sampleID), "n3", "n4"))
by_gene_KD=m

clust_df$monocle_cluster=as.character(clust_df$monocle_cluster)

by_gene_KD=by_gene_KD%>%dplyr::left_join(clust_df,by="monocle_cluster")

#####graph overview
ggplot(by_gene_KD, aes(x = factor(cell_types_abbreviations, levels = cell_order_short), y = frxn, fill=factor(cell_types_abbreviations, levels = cell_order_short))) + geom_bar(stat = "identity") +theme(legend.position='none')+
  scale_fill_okabeito(reverse=T)+
  facet_grid(~KD~factor(gene, levels = gene_order))
  
ylab = "% gene by treatment"
xlab = "cluster"

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "percent_cells_by_gene_clust.pdf"),
     width = 16, height = 15)

####graph by gene
ggplot(by_gene_KD, aes(x = KD, y = frxn, fill=c(factor(cell_types_abbreviations, levels = cell_order_short)))) + geom_bar(stat = "identity") +
  scale_fill_okabeito(reverse=T)+
  facet_wrap(~factor(gene, levels = gene_order), ncol=4)
  
ylab = "% gene by treatment"
xlab = "treatment"
last_plot() + labs(x = xlab, y = ylab)
output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_gene_by treatment_100_percent_cell_type_bis.pdf")
ggsave(filename = output_file, width = 10, height = 8)

######calculate ratio KD vs CTR
CTR=subset(by_gene_KD,by_gene_KD$KD=="CTR")%>%dplyr::rename(n_cells_gene_CTR=freq,tot_gene_CTR=tot,percent_gene_CTR=frxn)%>%select(-KD,-gene_KD)

KD=subset(by_gene_KD,by_gene_KD$KD=="KD")%>%dplyr::rename(n_cells_gene_KD=freq,tot_gene_KD=tot,percent_gene_KD=frxn)%>%select(-KD,-gene_KD)### NB. one cluster missing

by_gene_KD_ratio=CTR%>%dplyr::left_join(KD,by=c("gene","monocle_cluster","cell_types","cell_types_abbreviations"))
by_gene_KD_ratio <- replace(by_gene_KD_ratio, is.na(by_gene_KD_ratio), 0.01)

by_gene_KD_ratio=by_gene_KD_ratio%>%mutate(FC=percent_gene_KD/percent_gene_CTR)

#####graph overview FC
ggplot(by_gene_KD_ratio, aes(x = factor(cell_types_abbreviations, levels = cell_order_short), y = FC, fill=factor(cell_types_abbreviations, levels = cell_order_short))) + geom_bar(stat = "identity") +theme(legend.position='none')+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  scale_fill_okabeito(reverse=T)+
  facet_grid(~factor(gene, levels = gene_order))
  
ylab = "% gene KD vs CTR"
xlab = "cluster"

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "FC_cells_by_gene_clust.pdf"),
     width = 16, height = 4)


###save data
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster.csv")
write.csv(by_gene_KD, stat_filename, quote = F, row.names = T, sep=",")
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster_FC.csv")
write.csv(by_gene_KD_ratio, stat_filename, quote = F, row.names = T, sep=",")

Rdata_filename = paste0(fname_prefix_R, "_calculations_nb_percent.RData")
save(by_gene_KD,by_gene_KD_ratio,file = Rdata_filename)


```
##Assess frequency in clusters by gene
##Identify if clusters are too small to be analyzed (mode than 70 cells in at least 1 condition)
```{r}
########calculate KD and CTR separate
selection=by_gene_KD%>%select(cell_types_abbreviations,gene_KD,freq)
selection=spread(selection,key=gene_KD,value=freq)
selection <- replace(selection, is.na(selection), 0)
selection$has_70_counts <- apply(selection[, -1] >= 70, 1, any)

stat_filename = paste0(fname_prefix_stat, "_Freq_cells_by_gene_and_KD.csv")
write.csv(selection, stat_filename, quote = F, row.names = T, sep=",")

########calculate KD and CTR together and use this one
clust_count=meta %>%
  dplyr::group_by(assigned_cell_type_short,shRNA_gene) %>%
  dplyr::summarise(count = n())
clust_count=spread(clust_count,key=shRNA_gene,value=count)
clust_count <- replace(clust_count, is.na(clust_count), 0)
clust_count$has_70_counts <- apply(clust_count[, -1] >= 70, 1, any)

stat_filename = paste0(fname_prefix_stat, "_Freq_cells_by_gene.csv")
write.csv(clust_count, stat_filename, quote = F, row.names = T, sep=",")

clust_sel=clust_count%>%select(assigned_cell_type_short,has_70_counts)

meta=meta%>%left_join(clust_sel,by="assigned_cell_type_short")
cds@colData$has_70_counts=meta$has_70_counts

meta_cl=subset(meta,meta$has_70_counts==T)#from 5858 to 5445

```

######emake fisher stats per gene by selectiong only meaningful clusters
#####remake previous graphs

```{r}
##################
############ calculate % per cluster of each gene CTR and KD 
################
meta_to_use=meta_cl

tot=table(meta_to_use$g_KD) #total of the cluster
df=table(meta_to_use$g_KD,meta_to_use$monocle_clusters,meta_to_use$shRNA_gene,meta_to_use$KD)
df = as.data.frame(df)
colnames(df)=c("gene_KD","monocle_cluster","gene","KD","freq")

#mdf2 = melt(df)
mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("gene_KD","tot")
mdf2=mdf2%>%left_join(total, by="gene_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)
#m <- m %>%
 # mutate(samp = ifelse(grepl("n3_", sampleID), "n3", "n4"))
by_gene_KD=m

clust_df$monocle_cluster=as.character(clust_df$monocle_cluster)

by_gene_KD=by_gene_KD%>%dplyr::left_join(clust_df,by="monocle_cluster")

clust_select=clust_count%>%filter(has_70_counts==T)
color_df_select=color_df%>%filter(cell_order_short%in%clust_select$assigned_cell_type_short)
#####graph overview
ggplot(by_gene_KD, aes(x = factor(cell_types_abbreviations, levels = cell_order_short), y = frxn, fill=factor(cell_types_abbreviations, levels = cell_order_short))) + geom_bar(stat = "identity") +theme(legend.position='none')+
  scale_fill_manual(values=color_df_select$cl_col)+
  facet_grid(~KD~factor(gene, levels = gene_order))
  
ylab = "% gene by treatment"
xlab = "cluster"

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "percent_cells_by_gene_clust_polished.pdf"),
     width = 16, height = 15)

####graph by gene
ggplot(by_gene_KD, aes(x = KD, y = frxn, fill=c(factor(cell_types_abbreviations, levels = cell_order_short)))) + geom_bar(stat = "identity") +
  scale_fill_manual(values=color_df_select$cl_col)+
  facet_wrap(~factor(gene, levels = gene_order), ncol=4)
  
ylab = "% gene by treatment"
xlab = "treatment"
last_plot() + labs(x = xlab, y = ylab)
output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_gene_by treatment_100_percent_cell_type_bis_polished.pdf")
ggsave(filename = output_file, width = 10, height = 8)

######calculate ratio KD vs CTR
CTR=subset(by_gene_KD,by_gene_KD$KD=="CTR")%>%dplyr::rename(n_cells_gene_CTR=freq,tot_gene_CTR=tot,percent_gene_CTR=frxn)%>%select(-KD,-gene_KD)

KD=subset(by_gene_KD,by_gene_KD$KD=="KD")%>%dplyr::rename(n_cells_gene_KD=freq,tot_gene_KD=tot,percent_gene_KD=frxn)%>%select(-KD,-gene_KD)### NB. one cluster missing

by_gene_KD_ratio=CTR%>%dplyr::left_join(KD,by=c("gene","monocle_cluster","cell_types","cell_types_abbreviations"))
by_gene_KD_ratio <- replace(by_gene_KD_ratio, is.na(by_gene_KD_ratio), 0.01)

by_gene_KD_ratio=by_gene_KD_ratio%>%mutate(FC=percent_gene_KD/percent_gene_CTR)

#####graph overview FC
ggplot(by_gene_KD_ratio, aes(x = factor(cell_types_abbreviations, levels = cell_order_short), y = FC, fill=factor(cell_types_abbreviations, levels = cell_order_short))) + geom_bar(stat = "identity") +theme(legend.position='none')+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  scale_fill_manual(values=color_df_select$cl_col)+
  facet_grid(~factor(gene, levels = gene_order))
  
ylab = "% gene KD vs CTR"
xlab = "cluster"

last_plot() + labs(x = xlab, y = ylab)
ggsave(filename = paste0(fname_prefix_mosaic, "_", "FC_cells_by_gene_clust_polished.pdf"),
     width = 16, height = 4)


###save data
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster_polished.csv")
write.csv(by_gene_KD, stat_filename, quote = F, row.names = T, sep=",")
stat_filename = paste0(fname_prefix_stat, "_nb_percent_by_gene_in_each_cluster_FC_polished.csv")
write.csv(by_gene_KD_ratio, stat_filename, quote = F, row.names = T, sep=",")

Rdata_filename = paste0(fname_prefix_R, "_calculations_nb_percent_polished.RData")
save(by_gene_KD,by_gene_KD_ratio,file = Rdata_filename)
```



#####calculate fisher stats

```{r}
#make true false annotations per cluster to make the contingency tables

colData(cds)$clust1 = cds$monocle_clusters=="1"
colData(cds)$clust2 = cds$monocle_clusters=="2"
colData(cds)$clust3 = cds$monocle_clusters=="3"
colData(cds)$clust4 = cds$monocle_clusters=="4"
colData(cds)$clust5 = cds$monocle_clusters=="5"
colData(cds)$clust6 = cds$monocle_clusters=="6"
colData(cds)$clust7 = cds$monocle_clusters=="7"
colData(cds)$clust8 = cds$monocle_clusters=="8"
colData(cds)$clust9 = cds$monocle_clusters=="9"

meta=as.data.frame(cds@colData@listData)
meta <- replace(meta, is.na(meta), "empty")
#meta=subset(meta,meta$has_70_counts==T)

#####clusters to consider are 1 to 6
#make a loop to create the statistics
################## start without subsetting for cell populations

gene_list <- c("NKX2.5", "B2M", "KMT2D", "CHD7", "GATA4", "SMAD2", "NA", "SCR")
result_list <- list()

for (cluster in paste0("clust", 1:6)) {
  p_values <- numeric()

  for (gene in gene_list) {
    subset_data <- meta[meta$shRNA_gene == gene, c("KD", cluster)]
    
    if (nrow(subset_data) > 0) {
      # Perform chi-squared test
      fisher_result <- fisher.test(subset_data$KD, subset_data[[cluster]], simulate.p.value = FALSE)
      
      # Extract p-value
      p_value <- fisher_result[["p.value"]]
      
      p_values <- c(p_values, p_value)
    } else {
      # If no data for the gene in the current cluster, set p-value to NA
      p_values <- c(p_values, NA)
    }
  }

  result_df <- data.frame(gene = gene_list, cluster = cluster, p_value = p_values, stringsAsFactors = FALSE)
  
  # Add a q-value column using p.adjust
  result_df$q_value <- p.adjust(result_df$p_value, method = "fdr")
  
  # Add a column indicating significance at a 5% level
  result_df$sig <- result_df$q_value <= 0.05

  result_list[[cluster]] <- result_df
}

combined_result_df <- do.call(rbind, result_list)

######calculate FC (as before but divide controls)

tot=table(cds$gene_KD) #total of the cluster
df=table(cds$gene_KD,cds$monocle_clusters, cds$shRNA_gene,cds$KD)
df = as.data.frame(df)
colnames(df)=c("gene_KD","monocle_cluster","gene","KD","freq")

#mdf2 = melt(df)
mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("gene_KD","tot")
mdf2=mdf2%>%left_join(total, by="gene_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)
####add info
m=m%>%dplyr::left_join(clust_df,by="monocle_cluster")
m$cluster=paste0("clust",m$monocle_cluster)

CTR=subset(m,m$KD=="CTR")%>%dplyr::rename(n_cells_gene_CTR=freq,tot_gene_CTR=tot,percent_gene_CTR=frxn)%>%select(-KD,-gene_KD)
KD=subset(m,m$KD=="KD")%>%dplyr::rename(n_cells_gene_KD=freq,tot_gene_KD=tot,percent_gene_KD=frxn)%>%select(-KD,-gene_KD)

FC=CTR%>%dplyr::left_join(KD,by=c("gene","monocle_cluster","cluster","cell_types","cell_types_abbreviations"))
FC <- replace(FC, is.na(FC), 0.2)

FC=FC%>%mutate(FC=percent_gene_KD/percent_gene_CTR)
combined_result_df=combined_result_df%>%left_join(FC, by=c("gene","cluster"))

combined_result_df=subset(combined_result_df,!is.na(combined_result_df$monocle_cluster))

stat_filename = paste0(fname_prefix_stat, "_fisher_stat_genes_clusters.csv")
write.csv(combined_result_df, stat_filename, quote = F, row.names = T, sep=",")

sub=subset(clust_sel,clust_sel$has_70_counts==T)
color_df_select=color_df%>%filter(cell_order_short%in%sub$assigned_cell_type_short)

ggplot(data=combined_result_df, aes(x=log2(as.numeric(FC)), y=log10(q_value)*-1)) +
  geom_point(aes(colour = factor(cell_types_abbreviations, levels = cell_order_short)), size=1.5) +scale_colour_manual(values=color_df_select$cl_col)+
  ggrepel::geom_text_repel(
    data = subset(combined_result_df, sig==T),
    aes(label = gene),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #geom_vline(xintercept = log2(1.5),linetype="dotted")+
  #geom_vline(xintercept = -log2(1.5),linetype="dotted")+
  geom_hline(yintercept = -log10(0.05),linetype="dotted")+
  ggtitle("fisher test statistics per gene in each cluster")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2FC compared to CTR",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_fisher_stats.pdf"),
       width = 5, height = 4)


ggplot(data=combined_result_df, aes(x=as.numeric(percent_gene_KD), y=as.numeric(percent_gene_CTR))) +
  geom_point(aes(colour = factor(cell_types_abbreviations, levels = cell_order_short)), size=1.5) +scale_colour_manual(values=color_df_select$cl_col)+
  ggrepel::geom_text_repel(
    data = subset(combined_result_df, sig==T),
    aes(label = gene),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  geom_smooth(method ="lm",se = F,linetype=2,linewidth = 0.5)+
  ggtitle("fisher test statistics per gene in each cluster")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "% gene KD",
    y = "% gene CTR",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_percent_fisher_stats.pdf"),
       width = 5, height = 4)

cds_subset=cds[,cds$has_70_counts==T]

plot_cells(cds_subset, color_cells_by="assigned_cell_type_short", group_cells_by="assigned_cell_type",label_cell_groups = F,group_label_size = 7,label_roots=F,label_leaves=F,label_branch_points=F,cell_size=0.8)+
  scale_color_manual(name="cell types",values=color_df_select$cl_col)+
  labs(x = "UMAP 1", y = "UMAP 2", title = "Monocle clustering")
ggsave(filename = paste0(fname_prefix_UMAP, "_", "UMAP_clusters_POLISHED.pdf"),
     width = 8, height = 5)

```

######fisher stats by shRNA

```{r}
meta=as.data.frame(cds@colData@listData)

#####count shRNA to select the most meaninful ones (>40)
shRNA_count=meta%>%
  dplyr::group_by(shRNA_label,KD) %>%
  dplyr::summarise(count = n())###tot clones 305
shRNA_count=spread(shRNA_count, key = KD, value = count)%>%dplyr::rename(freq_KD=KD,freq_CTR=CTR)
shRNA_count <- replace(shRNA_count, is.na(shRNA_count), 0)

shRNA_count$shRNA_40_counts <- apply(shRNA_count[, -1] >= 40, 1, any)
sub=shRNA_count%>%dplyr::select(shRNA_label,shRNA_40_counts)
shRNA_selected=shRNA_count%>%filter(shRNA_40_counts==T)

stat_filename = paste0(fname_prefix_stat, "_shRNA_count.csv")
write.csv(shRNA_count, stat_filename, quote = F, row.names = T, sep=",")


#####add info to meta and then on cds
meta=meta%>%left_join(sub,by="shRNA_label")
colData(cds)$shRNA_40_counts=meta$shRNA_40_counts

colData(cds)$shRNA_KD=paste0(cds$shRNA_label,"_",cds$KD)###make a new column to calculate %
meta=as.data.frame(cds@colData@listData)
meta <- replace(meta, is.na(meta), "empty")

########select shRNA with more than 40 cells
#meta=subset(meta,meta$shRNA_40_counts==T)

####calculate % composition of each shRNA in clusters
tot=table(meta$shRNA_KD) #total of the cluster
df=table(meta$shRNA_KD,meta$monocle_clusters, meta$shRNA_label ,meta$shRNA_gene,meta$KD)
df = as.data.frame(df)
colnames(df)=c("shRNA_KD","monocle_cluster","shRNA","gene","KD","freq")

mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("shRNA_KD","tot")
mdf2=mdf2%>%left_join(total, by="shRNA_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)

by_shRNA_KD=m
#add cluster info
by_shRNA_KD=by_shRNA_KD%>%left_join(clust_df, by="monocle_cluster")

# Get unique genes
genes <- unique(by_shRNA_KD$gene)

# Loop through each gene
for (g in genes) {
  # Create a plot for the current gene
  p <- ggplot(subset(by_shRNA_KD, by_shRNA_KD$gene == g), 
              aes(x = KD, y = frxn, fill = factor(cell_types, levels = cell_order))) + 
    geom_bar(stat = "identity") +
    scale_fill_manual(values=color_df$cl_col) +
    labs(fill = "Cell type") +
    facet_wrap(~shRNA, ncol = 4) +
    labs(x = "treatment", y = "% clone by treatment")

  # Save the plot
  output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_shRNA_", g, "_100_percent_cell_type.pdf")
  ggsave(filename = output_file, plot = p, width = 8, height = 4)
}

#####polished graphs

polished=by_shRNA_KD%>%filter(shRNA%in%shRNA_selected$shRNA_label)

# Get unique genes
genes <- unique(polished$gene)

# Loop through each gene
for (g in genes) {
  # Create a plot for the current gene
  p <- ggplot(subset(polished, polished$gene == g), 
              aes(x = KD, y = frxn, fill = factor(cell_types, levels = cell_order))) + 
    geom_bar(stat = "identity") +
    scale_fill_manual(values=color_df$cl_col) +
    labs(fill = "Cell type") +
    facet_wrap(~shRNA, ncol = 4) +
    labs(x = "treatment", y = "% clone by treatment")

  # Save the plot
  output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_shRNA_polished_", g, "_100_percent_cell_type.pdf")
  ggsave(filename = output_file, plot = p, width = 8, height = 4)
}


########select shRNA with more than 40 cells
meta=subset(meta,meta$shRNA_40_counts==T)

####calculate fisher test

meta_sh=meta%>%select(shRNA_label,KD,clust1,clust2,clust3,clust4,clust5,clust6,clust7,clust8,clust9)
sh_list=unique(meta_sh$shRNA_label)

meta_fisher_CTR=data.frame(shRNA_label=sh_list,KD="CTR",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T)

meta_fisher_KD=data.frame(shRNA_label=sh_list,KD="KD",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T)

meta_fisher_CTR_F=data.frame(shRNA_label=sh_list,KD="CTR",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F)

meta_fisher_KD_F=data.frame(shRNA_label=sh_list,KD="KD",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F)

meta_sh=rbind(meta_sh,meta_fisher_CTR,meta_fisher_KD,meta_fisher_CTR_F,meta_fisher_KD_F)

result_list <- list()

for (cluster in paste0("clust", 1:6)) {
  p_values <- numeric()

  for (shRNA in sh_list) {
    subset_data <- meta_sh[meta_sh$shRNA_label == shRNA, c("KD", cluster)]
    
    if (nrow(subset_data) > 0) {
      # Perform Fisher's Exact Test
      fisher_result <- fisher.test(subset_data$KD, subset_data[[cluster]], simulate.p.value = FALSE)
      
      # Extract p-value
      p_value <- fisher_result$p.value
      
      p_values <- c(p_values, p_value)
    } else {
      # If no data for the clone in the current cluster, set p-value to NA
      p_values <- c(p_values, NA)
    }
  }

  result_df <- data.frame(shRNA = sh_list, cluster = cluster, p_value = p_values, stringsAsFactors = FALSE)
  
  # Add a q-value column using p.adjust
  result_df$q_value <- p.adjust(result_df$p_value, method = "BH")
  
  # Add a column indicating significance at a 5% level
  result_df$sig <- result_df$q_value <= 0.05

  result_list[[cluster]] <- result_df
}

combined_result_sh_df <- do.call(rbind, result_list)

m=by_shRNA_KD
CTR=subset(m,m$KD=="CTR")%>%dplyr::rename(n_cells_sh_CTR=freq,tot_sh_CTR=tot,percent_sh_CTR=frxn)%>%select(-KD,-shRNA_KD)
KD=subset(m,m$KD=="KD")%>%dplyr::rename(n_cells_sh_KD=freq,tot_sh_KD=tot,percent_sh_KD=frxn)%>%select(-KD,-shRNA_KD)

FC=CTR%>%dplyr::left_join(KD,by=c("gene","shRNA","monocle_cluster","cell_types","cell_types_abbreviations"))
FC <- replace(FC, is.na(FC), 0.3)

FC=FC%>%mutate(FC=percent_sh_KD/percent_sh_CTR,
               cluster=paste0("clust",monocle_cluster))
combined_result_sh_df=combined_result_sh_df%>%inner_join(FC, by=c("shRNA","cluster"))

ID=ID_key%>%dplyr::select(shRNA_ID_unified,shRNA_gene,shRNA_label,hex)%>%distinct(shRNA_ID_unified,shRNA_gene,shRNA_label,hex)%>%dplyr::rename(shRNA=shRNA_label)
combined_result_sh_df=combined_result_sh_df%>%left_join(ID, by=c("shRNA"))

Rdata_filename = paste0(fname_prefix_stat, "_fisher_stats.RData")
save(combined_result_sh_df,combined_result_df,meta,meta_sh,cds,
     file = Rdata_filename)

ggplot(data=subset(combined_result_sh_df,!is.na(combined_result_sh_df$gene)), aes(x=log2(as.numeric(FC)), y=log10(q_value)*-1)) +
  geom_point(aes(colour = factor(cell_types_abbreviations, levels = cell_order_short)), size=1.5) +scale_colour_manual(values=c(color_df_select$cl_col))+
  ggrepel::geom_text_repel(
    data = subset(combined_result_sh_df, sig==T),
    aes(label = shRNA),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #geom_vline(xintercept = log2(1.5),linetype="dotted")+
  #geom_vline(xintercept = -log2(1.5),linetype="dotted")+
  geom_hline(yintercept = -log10(0.05),linetype="dotted")+
  ggtitle("fisher test statistics per shRNA in each cluster")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2FC compared to CTR",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_fisher_stats_shRNA.pdf"),
       width = 6, height = 4)
```


######fisher stats by clone

```{r}
meta=as.data.frame(cds@colData@listData)
meta <- replace(meta, is.na(meta), "empty")
#####count clone to select the most meaninful ones (>40)

clone_count=meta%>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones 305
clone_count=spread(clone_count, key = KD, value = count)%>%dplyr::rename(freq_KD=KD,freq_CTR=CTR)
clone_count <- replace(clone_count, is.na(clone_count), 0)

clone_count$clone_20_counts <- apply(clone_count[, -1] >= 20, 1, any)
sub=clone_count%>%dplyr::select(clone,clone_20_counts)
clone_selected=clone_count%>%filter(clone_20_counts==T)

stat_filename = paste0(fname_prefix_stat, "_clone_count.csv")
write.csv(clone_count, stat_filename, quote = F, row.names = T, sep=",")

#####add info to meta and then on cds
meta=meta%>%left_join(sub,by="clone")
meta <- replace(meta, is.na(meta), FALSE)
colData(cds)$clone_20_counts_clone=meta$clone_20_counts

colData(cds)$clone_KD=paste0(cds$clone,"_",cds$KD)
meta=as.data.frame(cds@colData@listData)
meta <- replace(meta, is.na(meta), "empty")


####calculate % composition of each shRNA in clusters
tot=table(meta$clone_KD) #total of the cluster
df=table(meta$clone_KD,meta$monocle_clusters, meta$clone ,meta$shRNA_gene,meta$KD)
df = as.data.frame(df)
colnames(df)=c("clone_KD","monocle_cluster","clone","gene","KD","freq")

mdf2=df#11664 rows
total=as.data.frame(tot)
colnames(total)=c("clone_KD","tot")
mdf2=mdf2%>%left_join(total, by="clone_KD")
mdf2$tot = as.numeric(mdf2$tot)
mdf2$frxn = (mdf2$freq/mdf2$tot)*100
m=subset(mdf2,mdf2$freq!=0)

by_clone_KD=m
#add cluster info
by_clone_KD=by_clone_KD%>%left_join(clust_df, by="monocle_cluster")

# Get unique genes
genes <- unique(by_clone_KD$gene)

# Loop through each gene
for (g in genes) {
  # Create a plot for the current gene
  p <- ggplot(subset(by_clone_KD, by_clone_KD$gene == g), 
              aes(x = KD, y = frxn, fill = factor(cell_types, levels = cell_order))) + 
    geom_bar(stat = "identity") +
    scale_fill_okabeito(reverse = T) +
    labs(fill = "Cell type") +
    facet_wrap(~clone, ncol = 4) +
    labs(x = "treatment", y = "% clone by treatment")

  # Save the plot
  output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_clone_", g, "_100_percent_cell_type.pdf")
  ggsave(filename = output_file, plot = p, width = 8, height = 4)
}

#######polished graphs for clones

polished=by_clone_KD%>%filter(clone%in%clone_selected$clone)

# Get unique genes
genes <- unique(polished$gene)

# Loop through each gene
for (g in genes) {
  # Create a plot for the current gene
  p <- ggplot(subset(polished, polished$gene == g), 
              aes(x = KD, y = frxn, fill = factor(cell_types, levels = cell_order))) + 
    geom_bar(stat = "identity") +
    scale_fill_okabeito(reverse = T) +
    labs(fill = "Cell type") +
    facet_wrap(~clone, ncol = 4) +
    labs(x = "treatment", y = "% clone by treatment")

  # Save the plot
  output_file <- paste0(fname_prefix_mosaic, "_", "percent_by_clone_polished_", g, "_100_percent_cell_type.pdf")
  ggsave(filename = output_file, plot = p, width = 8, height = 4)
}


########select clusters with more than 70 cells and shRNA with more than 40 cells and clones with more than 20 cells
meta=subset(meta,meta$clone_20_counts_clone==T)

####calculate fisher test

meta_cl=meta%>%select(clone,KD,clust1,clust2,clust3,clust4,clust5,clust6,clust7,clust8,clust9)
cl_list=unique(meta_cl$clone)

meta_fisher_CTR=data.frame(clone=cl_list,KD="CTR",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T)

meta_fisher_KD=data.frame(clone=cl_list,KD="KD",clust1=T,clust2=T,clust3=T,clust4=T,clust5=T,clust6=T,clust7=T,clust8=T,clust9=T)

meta_fisher_CTR_F=data.frame(clone=cl_list,KD="CTR",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F)

meta_fisher_KD_F=data.frame(clone=cl_list,KD="KD",clust1=F,clust2=F,clust3=F,clust4=F,clust5=F,clust6=F,clust7=F,clust8=F,clust9=F)

meta_cl=rbind(meta_cl,meta_fisher_CTR,meta_fisher_KD,meta_fisher_CTR_F,meta_fisher_KD_F)

result_list <- list()

for (cluster in paste0("clust", 1:6)) {
  p_values <- numeric()

  for (clone in cl_list) {
    subset_data <- meta_cl[meta_cl$clone == clone, c("KD", cluster)]
    
    if (nrow(subset_data) > 0) {
      # Perform Fisher's Exact Test
      fisher_result <- fisher.test(subset_data$KD, subset_data[[cluster]], simulate.p.value = FALSE)
      
      # Extract p-value
      p_value <- fisher_result$p.value
      
      p_values <- c(p_values, p_value)
    } else {
      # If no data for the clone in the current cluster, set p-value to NA
      p_values <- c(p_values, NA)
    }
  }

  result_df <- data.frame(clone = cl_list, cluster = cluster, p_value = p_values, stringsAsFactors = FALSE)
  
  # Add a q-value column using p.adjust
  result_df$q_value <- p.adjust(result_df$p_value, method = "BH")
  
  # Add a column indicating significance at a 5% level
  result_df$sig <- result_df$q_value <= 0.05

  result_list[[cluster]] <- result_df
}

combined_result_cl_df <- do.call(rbind, result_list)#168

m=by_clone_KD
CTR=subset(m,m$KD=="CTR")%>%dplyr::rename(n_cells_cl_CTR=freq,tot_cl_CTR=tot,percent_cl_CTR=frxn)%>%select(-KD,-clone_KD)
KD=subset(m,m$KD=="KD")%>%dplyr::rename(n_cells_cl_KD=freq,tot_cl_KD=tot,percent_cl_KD=frxn)%>%select(-KD,-clone_KD)

FC=CTR%>%dplyr::left_join(KD,by=c("gene","clone","monocle_cluster","cell_types","cell_types_abbreviations"))
FC <- replace(FC, is.na(FC), 0.3)

FC=FC%>%mutate(FC=percent_cl_KD/percent_cl_CTR,
               cluster=paste0("clust",monocle_cluster))#273

combined_result_cl_df=combined_result_cl_df%>%left_join(FC, by=c("clone","cluster"))
info=meta%>%select(shRNA_gene,clone,cl_ID,clone_short,shRNA_label,assigned_cell_type,assigned_cell_type_short,monocle_clusters)%>%distinct(shRNA_gene,clone,cl_ID,clone_short,shRNA_label,assigned_cell_type,assigned_cell_type_short,monocle_clusters)%>%mutate(cluster=paste0("clust",monocle_clusters))

combined_result_cl_df=combined_result_cl_df%>%dplyr::select(-cell_types_abbreviations,-cell_types,-gene,-monocle_cluster)
combined_result_cl_df <- replace(combined_result_cl_df, is.na(combined_result_cl_df), 0)

combined_result_cl_df=combined_result_cl_df %>%left_join(info,by=c("clone","cluster"))
combined_result_cl_df <- subset(combined_result_cl_df, combined_result_cl_df$FC!=0)

Rdata_filename = paste0(fname_prefix_stat, "_fisher_stats.RData")
save(combined_result_cl_df,combined_result_sh_df,combined_result_df,meta,meta_sh,meta_cl,cds,
     file = Rdata_filename)

ggplot(data=subset(combined_result_cl_df,!is.na(combined_result_cl_df$shRNA_gene)), aes(x=log2(as.numeric(FC)), y=log10(q_value)*-1)) +
  geom_point(aes(colour = factor(assigned_cell_type_short, levels = cell_order_short)), size=1.5) +scale_colour_manual(values=color_df_select$cl_col)+
  ggrepel::geom_text_repel(
    data = subset(combined_result_cl_df, sig==T),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  geom_hline(yintercept = -log10(0.05),linetype="dotted")+
  #geom_vline(xintercept = log2(1.5),linetype="dotted")+
  #geom_vline(xintercept = -log2(1.5),linetype="dotted")+
  ggtitle("fisher test statistics per clone in each cluster")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2FC compared to CTR",
    y = "-Log10 adj p-value",
    color = "cell type"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_volcano_fisher_stats_clone.pdf"),
       width = 6, height = 4)

```


####save data

```{r}
meta=as.data.frame(cds@colData@listData)
meta_cardio=as.data.frame(cds_cardio@colData@listData)

Rdata_filename = paste0(fname_prefix_R, "_calculations_nb_percent_clones_shRNA_side2.RData")
save(by_gene_KD,by_gene_KD_ratio,by_shRNA_KD,by_clone_KD,clone_count,shRNA_count,
     file = Rdata_filename)

Rdata_filename = paste0(fname_prefix_R, "_cds_updated_side2.RData")
save(meta,cds,clones_AB03,clust_df,color_df,color_df_l,color_df_s,color_df_select,color_g,ID_key,info,meta_cardio,cds_cardio,
     file = Rdata_filename)
```


