---
title: "#A1 n5.2 3D E/D analysis (1) D0 2D vs CTR cardioidi; (2) D0 2D vs TET cardioidi; (3) CTR cardioidi vs TET cardioidi"
author: "Elisa Balmas PhD"
date: "2024-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up environment

```{r}
library(monocle3)
library(tidyr)
library(stringr)
library(grid)
library(viridis)
library(viridisLite)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(ggsignif)
library(umap)
library(heatmap3)
library(plyr)
library(edgeR)
library(compareGroups)
library(dbscan)
library(MAST)
library(geosphere)
library(RColorBrewer)
library(Seurat)
library(htmlwidgets)
library(Matrix)
library(tidyseurat)
library(SeuratWrappers)
library(scMAGeCK)
library(readxl)
library(rCASC)
library(deepToolsUtils)
library(R.utils)


library(ggplot2); library(reshape2); theme_set(theme_bw(26) + theme(panel.grid.major = element_blank(), 
                                                                    panel.grid.minor = element_blank()) +
                                                 theme(legend.key = element_blank()))
update_geom_defaults("point", aes(size = 4))

#setwd("/home/ebalmas/scpHUB/projects/H001AS7_iKDseq_10X_4tp_monocle/analysis_EB")

# Session options
options(stringsAsFactors = FALSE)
set.seed(12345)

# Set up the ggplot default params
theme_set(theme_bw(12) + 
          theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(),
                plot.title = element_text(size=15, face="bold", margin = margin(10,0,10,0)),
                axis.text.x = element_text(angle=45, hjust = 1)))


# Set up filename prefixes and output folders with the data
dir.create("Output")
dir.create(file.path("Output","scratch"))

dir.create(file.path("Output","revision"))
dir.create(file.path("Output","revision","E_D"))#create forders E_D (enrichment and depletion)
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d")))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "R_objects"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "dotplots"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "dotplots","single"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "stat"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "csv"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "UMAP"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "mosaic"))
dir.create(file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "heatmap"))


fname_scratch <- file.path("Output","scratch")

#set up forder path E_D
fname_prefix_R <- file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "R_objects", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot <- file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "dotplots", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_dotplot_single <- file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "dotplots", "single", 
                                                 format(Sys.Date(), "%y%m%d"))
fname_prefix_csv<-file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "csv", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_stat<-file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "stat", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_UMAP<-file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "UMAP", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_mosaic<-file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "mosaic", 
                               format(Sys.Date(), "%y%m%d"))
fname_prefix_heatmap<-file.path("Output","revision","E_D", format(Sys.Date(), "%y%m%d"), "heatmap", 
                               format(Sys.Date(), "%y%m%d"))


okabe_tab=read_excel("Output/scratch/color_scale.xlsx")
okabe_pal=okabe_tab$hex

okabe_pal=c("#E69F00","#56B4E9","#009E73","#f0E442","#0072B2","#D55E00","#CC79A7","#000000")

pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
         "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
         "#920000","#924900","#db6d00","#24ff24","#ffff6d")
```
####create dataset

```{r}
####data AB03
#load("Output/monocle/240411/R_objects/240411_cds_updated.RData")
#ID_key=read_excel("shRNA_IDs.xlsx")
#Rdata_filename = paste0(fname_prefix_R, "_AB03_starting_material_E_D.RData")
#save(clones_AB03,meta_AB03,
 #    file = Rdata_filename)

####data 2D
#load(paste0("/Users/elisabalmaselisabalmas/Library/CloudStorage/GoogleDrive-elisa.balmas@unito.it/Shared drives/MBC-HEDGe_EB/EB003_CHD/H001A_screening/H001AS7_H001AS5_monocle_analysis/ANALYSIS_PAPER_NEW/Output/monocle/240313/R_objects/240313_monocle_cds_clone_selected.RData"))
#Rdata_filename = paste0(fname_prefix_R, "_2D_starting_material_E_D.RData")
#save(meta_2D,
 #    file = Rdata_filename)
#load("/Output/revision/E_D/240819/R_objects/240819_2D_starting_material_E_D.RData")
#load("/Output/revision/E_D/240819/R_objects/240819_AB03_starting_material_E_D.RData")
#Rdata_filename = paste0(fname_prefix_R, "_starting_material_E_D.RData")
#save(clones_AB03,meta_AB03,meta_2D,
 #    file = Rdata_filename)

#load(paste0(getwd(),"/Output/revision/E_D/240819/R_objects/240819_starting_material_E_D.RData"))
#ID_key=read_excel("shRNA_IDs.xlsx")

#clones_info=clones_AB03

#meta_AB03$sample_ID=ifelse(meta_AB03$sample_id=="n3_no_TET","Cardioid n3 CT",
                     #               ifelse(meta_AB03$sample_id=="n3_TET","Cardioid n3 KD",
                      #                     ifelse(meta_AB03$sample_id=="n4_no_TET","Cardioid n4 CT",
                       #                           ifelse(meta_AB03$sample_id=="n4_TET","Cardioid n4 KD","else"))))

#meta_2D$knockdown="TET"
#meta_2D$treatment="TET"
#meta_2D$KD="KD"
#meta_AB03$DD="cardioid"
#meta_2D$ID="2D"
#meta_AB03$ID="3D"
#meta_2D$libID=paste0(meta_2D$lib_ID,"-",meta_2D$orig.ident,"_",meta_2D$ID)
#meta_AB03$libID=paste0(meta_AB03$lib_ID,"-",meta_AB03$orig.ident,"_",meta_AB03$ID)
#rownames(meta_AB03)=meta_AB03$libID
#rownames(meta_2D)=meta_2D$libID

#meta_2D=meta_2D%>%dplyr::select(lib_ID,libID,info, sample, sample_id,sample_ID, experiment, KD, ID, DD, treatment, batch, clone, side,shRNA_BC, shRNA_gene, cell_BC,shRNA_BC_FWD,shRNA_BC_RVS,shRNA_Sequence)

#meta_AB03=meta_AB03%>%dplyr::select(lib_ID,libID,info, sample, sample_id,sample_ID, experiment, KD, ID, DD, treatment, batch, clone, side,shRNA_BC, shRNA_gene, cell_BC,shRNA_BC_FWD,shRNA_BC_RVS,shRNA_Sequence)

#meta=rbind(meta_2D,meta_AB03) #41877 rows

#meta=meta%>%dplyr::right_join(clones_info, by=c("clone","shRNA_gene","cell_BC"))

#meta$set_ED=ifelse(meta$sample_id=="PSC","hiPSC",
                                 #   ifelse(meta$sample_id=="D0","hiPSC",
                                  #         ifelse(meta$sample_id=="n3_no_TET","cardioids_CT",
                                   #               ifelse(meta$sample_id=="n4_no_TET","cardioids_CT",
                                    #                     ifelse(meta$sample_id=="n3_TET","cardioids",                                                       #ifelse(meta$sample_id=="n4_TET","cardioids","other"))))))


#meta$KD_B2M= ifelse(meta$shRNA_gene == "B2M", "control", "gene")
#meta$KD_empty= ifelse(meta$shRNA_gene == "NA", "control", "gene")
#meta$KD_SCR= ifelse(meta$shRNA_gene == "SCR", "control", "gene")

#meta_ipsc=subset(meta, meta$set_ED!="other")

#meta_ipsc_TET=subset(meta_ipsc, meta_ipsc$set_ED!="cardioids_CT")

#meta_CT=subset(meta_ipsc, meta_ipsc$set_ED!="cardioids")

#S_filename = paste0(fname_scratch, "/E_D_starting_meta_all_sides")
#save(clones_info,meta,meta_AB03,meta_ipsc,meta_ipsc_TET,meta_CT,
 #    file = S_filename)

#meta_s2=subset(meta,meta$side=="Side2")

#meta_AB03_s2=subset(meta_AB03,meta_AB03$side=="Side2")
#meta_ipsc_s2=subset(meta_ipsc,meta_ipsc$side=="Side2")
#meta_ipsc_TET_s2=subset(meta_ipsc_TET,meta_ipsc_TET$side=="Side2")
#meta_CT_s2=subset(meta_CT,meta_CT$side=="Side2")

#S_filename = paste0(fname_scratch, "/E_D_starting_meta_s2")
#save(clones_info,meta,meta_s2,meta_AB03_s2,meta_ipsc_s2,meta_ipsc_TET_s2,meta_CT_s2,
 #    file = S_filename)

```

########load data

```{r}


load(paste0(getwd(),"/",fname_scratch,"/E_D_starting_meta_all_sides"))
load(paste0(getwd(),"/",fname_scratch,"/E_D_starting_meta_s2"))

```

####### fisher test enrichment depletion AB03
###ER between cardioids with and without TET

```{r}
#####total Fisher test
m_cl=meta_AB03%>%dplyr::select(clone,KD)
clone_list=unique(m_cl$clone)#52


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$KD, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_AB03 %>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = KD, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_AB03%>%
  dplyr::group_by(KD)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_CTR=totals[1,2]
counts$tot_KD=totals[2,2]
counts$TOT=counts$tot_CTR+counts$tot_KD
counts$TOT_clone=counts$CTR+counts$KD

counts$tot_other_CTR=counts$tot_CTR-counts$CTR
counts$tot_other_KD=counts$tot_KD-counts$KD
counts=counts%>%mutate(FC=(KD/tot_KD)/(CTR/tot_CTR),
                       percent_CTR=CTR/tot_CTR*100,
                       percent_KD=KD/tot_KD*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_AB03 %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_AB03_all.csv"), row.names=FALSE)
ED_AB03_all_results=result_ED_TOTAL

ED_AB03_all_results=subset(ED_AB03_all_results,ED_AB03_all_results$clone!="empty_NA_empty")

```

####volcano plot

```{r}
ggplot(data=ED_AB03_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_AB03_all_results,ED_AB03_all_results$qval<=0.05&log2(ED_AB03_all_results$FC)>fc_level|ED_AB03_all_results$qval<=0.05&log2(ED_AB03_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_AB03_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)

ggplot(data=ED_AB03_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = side, size=1.5)) +
  scale_colour_manual(values = okabe_pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_AB03_all_results,ED_AB03_all_results$qval<=0.05&log2(ED_AB03_all_results$FC)>fc_level|ED_AB03_all_results$qval<=0.05&log2(ED_AB03_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "side"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_AB03_clone_enrichment_depletion_side.pdf"),
       width = 8, height = 5)
```

####### fisher test enrichment depletion AB03 in SIDE 2 only
###ER between cardioids with and without TET

```{r}
#####total Fisher test
m_cl=meta_AB03_s2%>%dplyr::select(clone,KD)
clone_list=unique(m_cl$clone)#52


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$KD, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_AB03_s2 %>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = KD, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_AB03_s2%>%
  dplyr::group_by(KD)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_CTR=totals[1,2]
counts$tot_KD=totals[2,2]
counts$TOT=counts$tot_CTR+counts$tot_KD
counts$TOT_clone=counts$CTR+counts$KD

counts$tot_other_CTR=counts$tot_CTR-counts$CTR
counts$tot_other_KD=counts$tot_KD-counts$KD
counts=counts%>%mutate(FC=(KD/tot_KD)/(CTR/tot_CTR),
                       percent_CTR=CTR/tot_CTR*100,
                       percent_KD=KD/tot_KD*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_AB03_s2 %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_AB03_s2.csv"), row.names=FALSE)
ED_AB03_s2_results=result_ED_TOTAL
ED_AB03_s2_results=subset(ED_AB03_s2_results,ED_AB03_s2_results$clone!="empty_NA_empty")

```

####volcano plot

```{r}
ggplot(data=ED_AB03_s2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_AB03_s2_results,ED_AB03_s2_results$qval<=0.05&log2(ED_AB03_s2_results$FC)>fc_level|ED_AB03_s2_results$qval<=0.05&log2(ED_AB03_s2_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_AB03_clone_enrichment_depletion_SIDE2.pdf"),
       width = 8, height = 5)


```

####### fisher test enrichment depletion AB03 and hPSC
###ER between cardioids without TET and hPSC (controls) all sides

```{r}
#####total Fisher test
m_cl=meta_CT%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#50


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_CT %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_CT%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[2,2]
counts$tot_cardio=totals[1,2]

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$tot_cardio
counts$TOT_clone=counts$hiPSC+counts$cardioids_CT

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
counts$tot_other_cardioids=counts$tot_cardio-counts$cardioids_CT
counts=counts%>%mutate(FC=(cardioids_CT/tot_cardio)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_cardioids=cardioids_CT/tot_cardio*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_CT %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_CT_all.csv"), row.names=FALSE)
ED_CT_all_results=result_ED_TOTAL
ED_CT_all_results=subset(ED_CT_all_results,ED_CT_all_results$clone!="empty_NA_empty")


```

####volcano plot

```{r}
ggplot(data=ED_CT_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_CT_all_results,ED_CT_all_results$qval<=0.05&log2(ED_CT_all_results$FC)>fc_level|ED_CT_all_results$qval<=0.05&log2(ED_CT_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)

ggplot(data=ED_CT_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = side, size=1.5)) +
  scale_colour_manual(values = okabe_pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_CT_all_results,ED_CT_all_results$qval<=0.05&log2(ED_CT_all_results$FC)>fc_level|ED_CT_all_results$qval<=0.05&log2(ED_CT_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "side"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_clone_enrichment_depletion_side.pdf"),
       width = 8, height = 5)
```

####### fisher test enrichment depletion AB03 and hPSC
###ER between cardioids without TET and hPSC (controls)

```{r}
#####total Fisher test
m_cl=meta_CT_s2%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#22


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_CT_s2 %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_CT_s2%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[2,2]
counts$tot_cardio=totals[1,2]

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$tot_cardio
counts$TOT_clone=counts$hiPSC+counts$cardioids_CT

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
counts$tot_other_cardioids=counts$tot_cardio-counts$cardioids_CT
counts=counts%>%mutate(FC=(cardioids_CT/tot_cardio)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_cardioids=cardioids_CT/tot_cardio*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_CT_s2 %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_CT_side2.csv"), row.names=FALSE)
ED_CT_s2_results=result_ED_TOTAL
ED_CT_s2_results=subset(ED_CT_s2_results,ED_CT_s2_results$clone!="empty_NA_empty")


```

####volcano plot

```{r}
ggplot(data=ED_CT_s2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_CT_s2_results,ED_CT_s2_results$qval<=0.05&log2(ED_CT_s2_results$FC)>fc_level|ED_CT_s2_results$qval<=0.05&log2(ED_CT_s2_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_CT_s2_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)

```

####### fisher test enrichment depletion AB03 and hPSC
###ER between cardioids with TET and hPSC (controls) all sides

```{r}
#####total Fisher test
m_cl=meta_ipsc_TET%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#50


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_ipsc_TET %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_ipsc_TET%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[2,2]
counts$tot_cardio=totals[1,2]

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$tot_cardio
counts$TOT_clone=counts$hiPSC+counts$cardioids

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
counts$tot_other_cardioids=counts$tot_cardio-counts$cardioids
counts=counts%>%mutate(FC=(cardioids/tot_cardio)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_cardioids=cardioids/tot_cardio*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_ipsc_TET %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_ipsc_cardio_all.csv"), row.names=FALSE)
ED_ipsc_cardio_all_results=result_ED_TOTAL
ED_ipsc_cardio_all_results=subset(ED_CT_all_results,ED_CT_all_results$clone!="empty_NA_empty")


```

####volcano plot

```{r}
ggplot(data=ED_ipsc_cardio_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_ipsc_cardio_all_results,ED_ipsc_cardio_all_results$qval<=0.05&log2(ED_ipsc_cardio_all_results$FC)>fc_level|ED_ipsc_cardio_all_results$qval<=0.05&log2(ED_ipsc_cardio_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_ipsc_cardio_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)

ggplot(data=ED_ipsc_cardio_all_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = side, size=1.5)) +
  scale_colour_manual(values = okabe_pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_ipsc_cardio_all_results,ED_ipsc_cardio_all_results$qval<=0.05&log2(ED_ipsc_cardio_all_results$FC)>fc_level|ED_ipsc_cardio_all_results$qval<=0.05&log2(ED_ipsc_cardio_all_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "side"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_ipsc_cardio_clone_enrichment_depletion_side.pdf"),
       width = 8, height = 5)
```

####### fisher test enrichment depletion AB03 and hPSC
###ER between cardioids with TET and hPSC (controls) side 2

```{r}
#####total Fisher test
m_cl=meta_ipsc_TET_s2%>%dplyr::select(clone,set_ED)
clone_list=unique(m_cl$clone)#22


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$set_ED, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(2)
fc_level2 <- (log2(2))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta_ipsc_TET_s2 %>%
  dplyr::group_by(clone,set_ED) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = set_ED, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta_ipsc_TET_s2%>%
  dplyr::group_by(set_ED)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_hiPSC=totals[2,2]
counts$tot_cardio=totals[1,2]

#counts$hiPSC=counts$hiPSC+1
#counts$tot_hiPSC=counts$tot_hiPSC+1

counts$TOT=counts$tot_hiPSC+counts$tot_cardio
counts$TOT_clone=counts$hiPSC+counts$cardioids

counts$tot_other_CTR=counts$tot_hiPSC-counts$hiPSC
counts$tot_other_cardioids=counts$tot_cardio-counts$cardioids
counts=counts%>%mutate(FC=(cardioids/tot_cardio)/(hiPSC/tot_hiPSC),
                       percent_hiPSC=hiPSC/tot_hiPSC*100,
                       percent_cardioids=cardioids/tot_cardio*100,
                       percent_cl=TOT_clone/TOT*100)
side=meta_ipsc_TET_s2 %>%
     dplyr::group_by(clone,side)%>%
     dplyr::summarise(count_clone = n())


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_info,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(side,by="clone")

result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)

write.csv(result_ED_TOTAL, file= paste0(fname_prefix_csv, "_", "result_ED_ipsc_cardio_s2.csv"), row.names=FALSE)
ED_ipsc_cardio_s2_results=result_ED_TOTAL
ED_ipsc_cardio_s2_results=subset(ED_ipsc_cardio_s2_results,ED_ipsc_cardio_s2_results$clone!="empty_NA_empty")


```

####volcano plot

```{r}
ggplot(data=ED_ipsc_cardio_s2_results, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(ED_ipsc_cardio_s2_results,ED_ipsc_cardio_s2_results$qval<=0.05&log2(ED_ipsc_cardio_s2_results$FC)>fc_level|ED_ipsc_cardio_s2_results$qval<=0.05&log2(ED_ipsc_cardio_s2_results$FC)<fc_level2),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion Cardioid compared to hiPSC")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in Cardioids vs hiPSC",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_ipsc_cardio_side2_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)


```

##########################
######OLD E/D codes#######
##########################


####### fisher test enrichment depletion

```{r}
#####total Fisher test
m_cl=meta%>%dplyr::select(clone,KD)
clone_list=unique(m_cl$clone)#52


for (clone in clone_list) {
  # Create a new column named after the clone and assign boolean values
  m_cl[[clone]] <- m_cl$clone == clone
}

# Initialize empty vectors for p-values and clone names
p_values <- vector(mode="numeric", length = length(clone_list))
clone_names <- vector(mode="character", length = length(clone_list))

# Loop through each clone
for (i in 1:length(clone_list)) {
  # Filter data based on current clone
  # Perform Fisher's test
  fisher_result <- fisher.test(m_cl[[clone_list[i]]], m_cl$KD, simulate.p.value=TRUE)
  
  # Store p-value and clone name
  p_values[i] <- fisher_result$p.value
  clone_names[i] <- clone_list[i]
}

result_ED_TOTAL <- data.frame(clone = clone_names, P_value_TOTAL = p_values)
result_ED_TOTAL$qval=p.adjust(result_ED_TOTAL$P_value_TOTAL)

```

####

```{r}
sig_level <- 0.05
#Set a fold-change cutoff
fc_level <- log2(1.5)
fc_level2 <- (log2(1.5))*-1

result_ED_TOTAL$sig=result_ED_TOTAL$qval<=sig_level

counts=meta %>%
  dplyr::group_by(clone,KD) %>%
  dplyr::summarise(count = n())###tot clones
counts=spread(counts, key = KD, value = count)
counts <- counts %>% replace(is.na(.), 0)

totals=meta%>%
  dplyr::group_by(KD)%>%
  dplyr::summarise(total = n())
totals=as.data.frame(totals)

counts$tot_CTR=totals[1,2]
counts$tot_KD=totals[2,2]
counts$TOT=counts$tot_CTR+counts$tot_KD
counts$TOT_clone=counts$CTR+counts$KD

counts$tot_other_CTR=counts$tot_CTR-counts$CTR
counts$tot_other_KD=counts$tot_KD-counts$KD
counts=counts%>%mutate(FC=(KD/tot_KD)/(CTR/tot_CTR),
                       percent_CTR=CTR/tot_CTR*100,
                       percent_KD=KD/tot_KD*100,
                       percent_cl=TOT_clone/TOT*100)


result_ED_TOTAL=result_ED_TOTAL%>%left_join(counts,by="clone")
result_ED_TOTAL=result_ED_TOTAL%>%left_join(clones_AB03,by="clone")
result_ED_TOTAL=result_ED_TOTAL %>% 
  filter_all(all_vars(!is.infinite(.)))
result_ED_TOTAL=subset(result_ED_TOTAL,result_ED_TOTAL$FC>0)
```

####volcano plot

```{r}
ggplot(data=result_ED_TOTAL, aes(x=log2(FC), y=(log2(percent_cl)))) +
  geom_point(aes(colour = log10(qval)*-1, size=1.5)) +
  scale_color_viridis(name = "-log10(adj p-value)",
                         option="E", discrete=F)+
  #scale_colour_manual(values = pal)+
  #geom_text(aes(label=to_show_2vsother), size=4, hjust=1.2)+
  ggrepel::geom_text_repel(
    data = subset(result_ED_TOTAL,result_ED_TOTAL$qval<=0.05&log2(result_ED_TOTAL$percent_cl)>1.5|result_ED_TOTAL$qval<=0.05&log2(result_ED_TOTAL$FC)*-1.5>1.5),
    aes(label = clone_short),
    size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  )+
  #theme(legend.position = "none") +
  geom_vline(xintercept = fc_level, linetype="dotted")+
  geom_vline(xintercept = fc_level2, linetype="dotted")+
  #ylim(-5,1.5)+
  #xlim(-6,1)+
 # geom_vline(xintercept = c(fc_level, fc_level2))
  ggtitle("Clone enrichment depletion KD compared to CTR")+
  theme_minimal() +  # You can change the theme as needed
  labs(
    x = "Log2 of clone fraction in KD vs CTR",
    y = "Log2 of clone fraction",
    color = "-log10(q value)"  # Legend title
  ) 

ggsave(filename = paste0(fname_prefix_dotplot, "_clone_enrichment_depletion.pdf"),
       width = 8, height = 5)
```


